---
title: "Plots of metabolite data Arabidopsis field trials"
author: "Heike Sprenger"
date: "Monday, July 18, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
#getwd()
#setwd("D:/work/repos/ara_field_trials/")
#setwd("~/work/repos/ara_field_trials/")
```


# Load workspace, packages and scripts
```{r load workspace}
# load packages
library(knitr)
library(pander)
library(ggplot2)
library(lattice)
library(gplots)
library(pcaMethods)
library(plyr)

trellis.device() # zuruecksetzen auf default
show.settings() # Farbeinstellungen usw. anzeigen

# png default size: 480x480 pixel

# set options for knitr
opts_chunk$set(fig.width=6, fig.height=5, tidy=TRUE, tidy.opts=list(width.cutoff=70))
opts_knit$set(root.dir = '../')

# load workspace
#load("metabolite_arafield.RData")
```


# Source functions
```{r source functions}
source("../functions/func_bw_plot.R", encoding = 'UTF-8')
source("../functions/func_boxplot.R", encoding = 'UTF-8')
source("../functions/func_cv.R", encoding = 'UTF-8')
```


# Load data
Use changed metabolite data, from ``metabolite_arafield4_overview4.Rmd``
```{r load data}
metabolite_arafield2 <- read.table("output/metabolite_arafield2_modified.txt", header = TRUE, sep = "\t")
metabolite_arafield3 <- read.table("output/metabolite_arafield3_modified.txt", header = TRUE, sep = "\t")
metabolite_arafield4 <- read.table("output/metabolite_arafield4_modified.txt", header = TRUE, sep = "\t")

lt50 <- read.table("data/LT50_64acc.txt", header=TRUE, sep="\t")
dim(lt50) # 64 accessions
pander(head(lt50))

lt50$Capacity <- lt50$LT50ACC - lt50$LT50NAC
```

## set factor levels correctly
```{r set factor levels correctly}
# antho score (en)
metabolite_arafield2$antho_value_en <- factor(metabolite_arafield2$antho_value_en, 
                                              levels = c("very low", "low", "intermediate", "high", "very high"))
metabolite_arafield3$antho_value_en <- factor(metabolite_arafield3$antho_value_en, 
                                              levels = c("low", "intermediate", "high", "very high"))
metabolite_arafield4$antho_value_en <- factor(metabolite_arafield4$antho_value_en, 
                                              levels = c("very low", "low", "intermediate", "high", "very high"))

# bbch categories
metabolite_arafield2$bbch_categories <- factor(metabolite_arafield2$bbch_categories, 
                                               levels = c("vegetative", "generative buds", "flowering"))
metabolite_arafield3$bbch_categories <- factor(metabolite_arafield3$bbch_categories, 
                                               levels = c("vegetative", "generative buds", "flowering"))
metabolite_arafield4$bbch_categories <- factor(metabolite_arafield4$bbch_categories, 
                                               levels = c("vegetative", "generative buds", "flowering"))

# flowering state
metabolite_arafield2$flowering_state <- factor(metabolite_arafield2$flowering_state, 
                                               levels = c("non-flowering", "flowering"))
metabolite_arafield3$flowering_state <- factor(metabolite_arafield3$flowering_state, 
                                               levels = c("non-flowering", "flowering"))
metabolite_arafield4$flowering_state <- factor(metabolite_arafield4$flowering_state, 
                                               levels = c("non-flowering", "flowering"))

# vegetative state
metabolite_arafield2$vegetative_state <- factor(metabolite_arafield2$vegetative_state, 
                                                levels = c("vegetative", "non-vegetative"))
metabolite_arafield3$vegetative_state <- factor(metabolite_arafield3$vegetative_state, 
                                                levels = c("vegetative", "non-vegetative"))
metabolite_arafield4$vegetative_state <- factor(metabolite_arafield4$vegetative_state, 
                                                levels = c("vegetative", "non-vegetative"))
```


# Boxplots (from lattice package)
## Boxplot of metabolites and LT50 for all samples
### Boxplots for LT50
```{r boxplots from lattice package LT50}
pdf("figures/Metabolite_AraField_LT50_bwplots.pdf")   
func_bw_plot(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants", 
             xlab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants", 
             xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot(metabolite_arafield4$capacity, title = "Acclimation capacity", 
             xlab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots for Field 2
```{r boxplots from lattice package field 2}
pdf("figures/Metabolite_AraField2_bwplots.pdf")   
func_bw_plot(metabolite_arafield2$glucoseT, title = "Glucose")
func_bw_plot(metabolite_arafield2$fructoseT, title = "Fructose")
func_bw_plot(metabolite_arafield2$sucroseT, title = "Sucrose")
func_bw_plot(metabolite_arafield2$raffinoseT, title = "Raffinose")
func_bw_plot(metabolite_arafield2$amino_acidsT, title = "Amino acids")
func_bw_plot(metabolite_arafield2$malateT, title = "Malate")
func_bw_plot(metabolite_arafield2$fumarateT, title = "Fumarate")
func_bw_plot(metabolite_arafield2$prolineT, title = "Proline")
dev.off()
```

### Boxplots for Field 3
```{r boxplots from lattice package field 3}
pdf("figures/Metabolite_AraField3_bwplots.pdf")   
func_bw_plot(metabolite_arafield3$glucoseT, title = "Glucose")
func_bw_plot(metabolite_arafield3$fructoseT, title = "Fructose")
func_bw_plot(metabolite_arafield3$sucroseT, title = "Sucrose")
func_bw_plot(metabolite_arafield3$raffinoseT, title = "Raffinose")
func_bw_plot(metabolite_arafield3$amino_acidsT, title = "Amino acids")
func_bw_plot(metabolite_arafield3$malateT, title = "Malate")
func_bw_plot(metabolite_arafield3$fumarateT, title = "Fumarate")
func_bw_plot(metabolite_arafield3$prolineT, title = "Proline")
dev.off()
```

### Boxplots for Field 4
```{r boxplots from lattice package field 4}
pdf("figures/Metabolite_AraField4_bwplots.pdf")   
func_bw_plot(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot(metabolite_arafield4$malateT, title = "Malate")
func_bw_plot(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot(metabolite_arafield4$prolineT, title = "Proline")
dev.off()
```


## Boxplots of metabolites and LT50 per BBCH category
### Boxplots of metabolites and LT50 per BBCH category: Field 2
```{r boxplots of metabolites and LT50 per BBCH category field 2}
pdf("figures/Metabolite_AraField2_bwplots_per_bbch.pdf")   
cat1_value <- metabolite_arafield2$bbch_categories
func_bw_plot_cat1(metabolite_arafield2$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield2$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield2$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield2$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield2$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield2$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield2$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield2$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield2$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield2$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield2$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots of metabolites and LT50 per BBCH category: Field 3
```{r boxplots of metabolites and LT50 per BBCH category field 3}
pdf("figures/Metabolite_AraField3_bwplots_per_bbch.pdf")   
cat1_value <- metabolite_arafield3$bbch_categories
func_bw_plot_cat1(metabolite_arafield3$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield3$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield3$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield3$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield3$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield3$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield3$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield3$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield3$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield3$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield3$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots of metabolites and LT50 per BBCH category: Field 4
```{r boxplots of metabolites and LT50 per BBCH category field 4}
pdf("figures/Metabolite_AraField4_bwplots_per_bbch.pdf")   
cat1_value <- metabolite_arafield4$bbch_categories
func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield4$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield4$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


## Boxplots of metabolites and LT50 per flowering state
**here: is equal to BBCH category because there are no plants in vegetative stage**
### Boxplots of metabolites and LT50 per flowering state: Field 2
```{r boxplots of metabolites and LT50 per flowering state: Field 2}
pdf("figures/Metabolite_AraField2_bwplots_per_flowering_state.pdf")  
trellis.par.set(new_theme)
cat1_value <- metabolite_arafield2$flowering_state

func_bw_plot_cat1(metabolite_arafield2$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield2$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield2$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield2$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield2$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield2$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield2$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield2$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield2$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield2$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield2$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots of metabolites and LT50 per flowering state: Field 3
```{r boxplots of metabolites and LT50 per flowering state: Field 3}
pdf("figures/Metabolite_AraField3_bwplots_per_flowering_state.pdf")  
trellis.par.set(new_theme)
cat1_value <- metabolite_arafield3$flowering_state

func_bw_plot_cat1(metabolite_arafield3$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield3$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield3$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield3$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield3$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield3$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield3$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield3$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield3$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield3$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield3$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots of metabolites and LT50 per flowering state: Field 4
```{r boxplots of metabolites and LT50 per flowering state: Field 4}
pdf("figures/Metabolite_AraField4_bwplots_per_flowering_state.pdf")  
trellis.par.set(new_theme)
cat1_value <- metabolite_arafield4$flowering_state

func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield4$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield4$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


## Boxplots of metabolites and LT50 per vegetative state
### Boxplots of metabolites and LT50 per vegetative state: Field 2
```{r boxplots of metabolites and LT50 per vegetative state: Field 2}
pdf("figures/Metabolite_AraField2_bwplots_per_vegetative_state.pdf")
cat1_value <- metabolite_arafield2$vegetative_state
func_bw_plot_cat1(metabolite_arafield2$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield2$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield2$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield2$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield2$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield2$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield2$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield2$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield2$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield2$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield2$capacity, title = "Acclimation capacity",
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots of metabolites and LT50 per vegetative state: Field 3
```{r boxplots of metabolites and LT50 per vegetative state: Field 3}
pdf("figures/Metabolite_AraField3_bwplots_per_vegetative_state.pdf")
cat1_value <- metabolite_arafield3$vegetative_state
func_bw_plot_cat1(metabolite_arafield3$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield3$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield3$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield3$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield3$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield3$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield3$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield3$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield3$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield3$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield3$capacity, title = "Acclimation capacity",
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


### Boxplots of metabolites and LT50 per vegetative state: Field 4
**doesn't make sense here: there are no plants in vegetative stage!**
```{r boxplots of metabolites and LT50 per vegetative state: Field 4}
# pdf("figures/Metabolite_AraField4_bwplots_per_vegetative_state.pdf")  
# cat1_value <- metabolite_arafield4$vegetative_state
# func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
# func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
# func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
# func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
# func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
# func_bw_plot_cat1(metabolite_arafield4$malateT, title = "Malate")
# func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
# func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
# func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
#                   ylab_text = expression("LT"[50]*"NAC in °C"))
# func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
#                   xlab_text = expression("LT"[50]*"ACC in °C"))
# func_bw_plot_cat1(metabolite_arafield4$capacity, title = "Acclimation capacity", 
#                   ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
# dev.off()
```


## Boxplots of metabolites and LT50 per anthocyanin score
### Boxplots of metabolites and LT50 per anthocyanin score: Field 2
```{r boxplots of metabolites and LT50 per anthocyanin score: Field 2}
pdf("figures/Metabolite_AraField2_bwplots_per_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield2$antho_value_en
func_bw_plot_cat1(metabolite_arafield2$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield2$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield2$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield2$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield2$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield2$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield2$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield2$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield2$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield2$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield2$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots of metabolites and LT50 per anthocyanin score: Field 3
```{r boxplots of metabolites and LT50 per anthocyanin score: Field 3}
pdf("figures/Metabolite_AraField3_bwplots_per_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield3$antho_value_en
func_bw_plot_cat1(metabolite_arafield3$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield3$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield3$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield3$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield3$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield3$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield3$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield3$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield3$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield3$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield3$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

### Boxplots of metabolites and LT50 per anthocyanin score: Field 4
```{r boxplots of metabolites and LT50 per anthocyanin score: Field 4}
pdf("figures/Metabolite_AraField4_bwplots_per_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield4$antho_value_en
func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield4$malateT, title = "Malate")
func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(metabolite_arafield4$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


## Boxplots of metabolites and LT50 per BBCH categories and anthocyanin score
**Attention: Chunck is not evaluated because it resulted in error messages during knitting HTML document, but not during normal code running!**
### Boxplots of metabolites and LT50 per BBCH categories and anthocyanin score: Field 2
```{r boxplots of metabolites and LT50 per BBCH categories and anthocyanin score: Field 2, eval=FALSE}
pdf("figures/Metabolite_AraField2_bwplots_per_bbch_and_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield2$antho_value_en
cat2_value <- metabolite_arafield2$bbch_categories
func_bw_plot_cat2(metabolite_arafield2$glucoseT, title = "Glucose")
func_bw_plot_cat2(metabolite_arafield2$fructoseT, title = "Fructose")
func_bw_plot_cat2(metabolite_arafield2$sucroseT, title = "Sucrose")
func_bw_plot_cat2(metabolite_arafield2$raffinoseT, title = "Raffinose")
func_bw_plot_cat2(metabolite_arafield2$amino_acidsT, title = "Amino acids")
func_bw_plot_cat2(metabolite_arafield2$malateT, title = "Malate")
func_bw_plot_cat2(metabolite_arafield2$fumarateT, title = "Fumarate")
func_bw_plot_cat2(metabolite_arafield2$prolineT, title = "Proline")
func_bw_plot_cat2(metabolite_arafield2$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat2(metabolite_arafield2$LT50ACC, title = "Frost tolerance of acclimated plants",
                  ylab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat2(metabolite_arafield2$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


### Boxplots of metabolites and LT50 per BBCH categories and anthocyanin score: Field 3
```{r boxplots of metabolites and LT50 per BBCH categories and anthocyanin score: Field 3, eval=FALSE}
pdf("figures/Metabolite_AraField3_bwplots_per_bbch_and_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield3$antho_value_en
cat2_value <- metabolite_arafield3$bbch_categories
func_bw_plot_cat2(metabolite_arafield3$glucoseT, title = "Glucose")
func_bw_plot_cat2(metabolite_arafield3$fructoseT, title = "Fructose")
func_bw_plot_cat2(metabolite_arafield3$sucroseT, title = "Sucrose")
func_bw_plot_cat2(metabolite_arafield3$raffinoseT, title = "Raffinose")
func_bw_plot_cat2(metabolite_arafield3$amino_acidsT, title = "Amino acids")
func_bw_plot_cat2(metabolite_arafield3$malateT, title = "Malate")
func_bw_plot_cat2(metabolite_arafield3$fumarateT, title = "Fumarate")
func_bw_plot_cat2(metabolite_arafield3$prolineT, title = "Proline")
func_bw_plot_cat2(metabolite_arafield3$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat2(metabolite_arafield3$LT50ACC, title = "Frost tolerance of acclimated plants",
                  ylab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat2(metabolite_arafield3$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


### Boxplots of metabolites and LT50 per BBCH categories and anthocyanin score: Field 4
```{r boxplots of metabolites and LT50 per BBCH categories and anthocyanin score: Field 4, eval=FALSE}
pdf("figures/Metabolite_AraField4_bwplots_per_bbch_and_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield4$antho_value_en
cat2_value <- metabolite_arafield4$bbch_categories
func_bw_plot_cat2(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat2(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat2(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat2(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat2(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat2(metabolite_arafield4$malateT, title = "Malate")
func_bw_plot_cat2(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat2(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat2(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat2(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  ylab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat2(metabolite_arafield4$capacity, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


# Boxplots (from graphics package)
## Boxplots to get overview of measurements
## Boxplots to get overview of measurements: LT50
```{r boxplots from graphics package: LT50}
pdf("figures/Metabolite_AraField_LT50_Boxplots.pdf")
par(mar=c(5.1, 5.1, 4.1, 2.1))
func_boxplot_simple(lt50$LT50NAC, title = "Frost tolerance of non-acclimated plants", 
             ylab_text = expression("LT"[50]*"NAC in °C"))
func_boxplot_simple(lt50$LT50ACC, title = "Frost tolerance of acclimated plants", 
             ylab_text = expression("LT"[50]*"ACC in °C"))
func_boxplot_simple(lt50$Capacity, title = "Acclimation capacity", 
             ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```

## Boxplots to get overview of measurements: Field 2
```{r boxplots from graphics package: Field 2}
pdf("figures/Metabolite_AraField2_Boxplots.pdf")
par(mar=c(5.1, 5.1, 4.1, 2.1))
func_boxplot_simple(metabolite_arafield2$glucoseT, title = "Glucose")
func_boxplot_simple(metabolite_arafield2$fructoseT, title = "Fructose")
func_boxplot_simple(metabolite_arafield2$sucroseT, title = "Sucrose")
func_boxplot_simple(metabolite_arafield2$raffinoseT, title = "Raffinose")
func_boxplot_simple(metabolite_arafield2$amino_acidsT, title = "Amino acids")
func_boxplot_simple(metabolite_arafield2$malateT, title = "Malate")
func_boxplot_simple(metabolite_arafield2$fumarateT, title = "Fumarate")
func_boxplot_simple(metabolite_arafield2$prolineT, title = "Proline")
dev.off()
```

## Boxplots to get overview of measurements: Field 3
```{r boxplots from graphics package: Field 3}
pdf("figures/Metabolite_AraField3_Boxplots.pdf")
par(mar=c(5.1, 5.1, 4.1, 2.1))
func_boxplot_simple(metabolite_arafield3$glucoseT, title = "Glucose")
func_boxplot_simple(metabolite_arafield3$fructoseT, title = "Fructose")
func_boxplot_simple(metabolite_arafield3$sucroseT, title = "Sucrose")
func_boxplot_simple(metabolite_arafield3$raffinoseT, title = "Raffinose")
func_boxplot_simple(metabolite_arafield3$amino_acidsT, title = "Amino acids")
func_boxplot_simple(metabolite_arafield3$malateT, title = "Malate")
func_boxplot_simple(metabolite_arafield3$fumarateT, title = "Fumarate")
func_boxplot_simple(metabolite_arafield3$prolineT, title = "Proline")
dev.off()
```

## Boxplots to get overview of measurements: Field 4
```{r boxplots from graphics package: Field 4}
pdf("figures/Metabolite_AraField4_Boxplots.pdf")
par(mar=c(5.1, 5.1, 4.1, 2.1))
func_boxplot_simple(metabolite_arafield4$glucoseT, title = "Glucose")
func_boxplot_simple(metabolite_arafield4$fructoseT, title = "Fructose")
func_boxplot_simple(metabolite_arafield4$sucroseT, title = "Sucrose")
func_boxplot_simple(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_boxplot_simple(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_boxplot_simple(metabolite_arafield4$malateT, title = "Malate")
func_boxplot_simple(metabolite_arafield4$fumarateT, title = "Fumarate")
func_boxplot_simple(metabolite_arafield4$prolineT, title = "Proline")
dev.off()
```


## Boxplots to get overview of coefficient of variation (CV)
### Boxplots to get overview of coefficient of variation (CV): Field 2
```{r boxplots to get overview of coefficient of variation: Field 2}
# calculate CV per metabolite and accession
metabolite_arafield2_sub <- subset(metabolite_arafield2, 
                                 select= c(accession, glucose_positive, fructose_positive, sucrose_positive, 
                                           raffinose_positive, amino_acids_positive, malate_positive, 
                                           fumarate_positive, proline_positive))
metabolite_arafield2_CV <- aggregate(. ~ accession, metabolite_arafield2_sub, func_CV)

ylab_text_cv <- "Coefficient of variation (CV)"

pdf("figures/Metabolite_AraField2_CV_Boxplots.pdf")
par(mar=c(9, 5, 4, 2))
# all metabolites
boxplot(metabolite_arafield2_CV[,-1], main = "all metabolites", ylab = ylab_text_cv, las=2)
# single metabolites
par(mar=c(5, 5, 4, 2))
# func_boxplot_simple(metabolite_arafield2_CV$glucose_positive, title = "Glucose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$fructose_positive, title = "Fructose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$sucrose_positive, title = "Sucrose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$raffinose_positive, title = "Raffinose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$amino_acids_positive, title = "Amino acids", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$malate_positive, title = "Malate", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$fumarate_positive, title = "Fumarate", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$proline_positive, title = "Proline", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield2_CV$anthocyanin_positive, title = "Anthocyanins", ylab_text = ylab_text_cv)
dev.off()
```

### Boxplots to get overview of coefficient of variation (CV): Field 3
```{r boxplots to get overview of coefficient of variation: Field 3}
# calculate CV per metabolite and accession
metabolite_arafield3_sub <- subset(metabolite_arafield3, 
                                 select= c(accession, glucose_positive, fructose_positive, sucrose_positive, 
                                           raffinose_positive, amino_acids_positive, malate_positive, 
                                           fumarate_positive, proline_positive))
metabolite_arafield3_CV <- aggregate(. ~ accession, metabolite_arafield3_sub, func_CV)

pdf("figures/Metabolite_AraField3_CV_Boxplots.pdf")
par(mar=c(9, 5, 4, 2))
# all metabolites
boxplot(metabolite_arafield3_CV[,-1], main = "all metabolites", ylab = ylab_text_cv, las=2)
# single metabolites
par(mar=c(5, 5, 4, 2))
# func_boxplot_simple(metabolite_arafield3_CV$glucose_positive, title = "Glucose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield3_CV$fructose_positive, title = "Fructose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield3_CV$sucrose_positive, title = "Sucrose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield3_CV$raffinose_positive, title = "Raffinose", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield3_CV$amino_acids_positive, title = "Amino acids", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield3_CV$malate_positive, title = "Malate", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield3_CV$fumarate_positive, title = "Fumarate", ylab_text = ylab_text_cv)
# func_boxplot_simple(metabolite_arafield3_CV$proline_positive, title = "Proline", ylab_text = ylab_text_cv)
dev.off()
```

### Boxplots to get overview of coefficient of variation (CV): Field 4
```{r boxplots to get overview of coefficient of variation: Field 4}
# calculate CV per metabolite and accession
metabolite_arafield4_sub <- subset(metabolite_arafield4, 
                                 select= c(accession, glucose_positive, fructose_positive, sucrose_positive, 
                                           raffinose_positive, amino_acids_positive, malate_positive, 
                                           fumarate_positive, proline_positive))
head(metabolite_arafield4_sub)
metabolite_arafield4_CV <- aggregate(. ~ accession, metabolite_arafield4_sub, func_CV)
head(metabolite_arafield4_CV)

pdf("figures/Metabolite_AraField4_CV_Boxplots.pdf")
par(mar=c(9, 5, 4, 2))
# all metabolites
boxplot(metabolite_arafield4_CV[,-1], main = "all metabolites", ylab = ylab_text_cv, las=2)
# single metabolites
par(mar=c(5, 5, 4, 2))
func_boxplot_simple(metabolite_arafield4_CV$glucose_positive, title = "Glucose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$fructose_positive, title = "Fructose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$sucrose_positive, title = "Sucrose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$raffinose_positive, title = "Raffinose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$amino_acids_positive, title = "Amino acids", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$malate_positive, title = "Malate", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$fumarate_positive, title = "Fumarate", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$proline_positive, title = "Proline", ylab_text = ylab_text_cv)
dev.off()
```


# Pairs plot
```{r pairs plot}
nonNeg_colnames <- c("glucose_positive", "fructose_positive", "sucrose_positive",
                    "raffinose_positive", "amino_acids_positive", "malate_positive", 
                    "fumarate_positive", "proline_positive")

pairs_colnames <- c("LT50NAC", "LT50ACC", "capacity","norm_bbch_id", "anthoscore", nonNeg_colnames)
pairs_labels <- c(expression("LT"[50]*"NAC"),expression("LT"[50]*"ACC"),"Capa- \n city",
                  "norm BBCH", "Antho- \n Score","Glc","Frc","Suc","Raf","AA","Mal","Fum","Pro")

pdf("figures/Metabolite_AraField2_Pairs.pdf", width=15, height=15)
pairs(metabolite_arafield2[, pairs_colnames], panel = panel.smooth, 
      labels = pairs_labels, cex.labels = 2, font.labels = 2, cex.main = 2, 
      cex.axis = 1.5, main = "complete dataset field trial 2", gap = 0.1)
dev.off()

pdf("figures/Metabolite_AraField3_Pairs.pdf", width=15, height=15)
pairs(metabolite_arafield3[, pairs_colnames], panel = panel.smooth, 
      labels = pairs_labels, cex.labels = 2, font.labels = 2, cex.main = 2, 
      cex.axis = 1.5, main = "complete dataset field trial 3", gap = 0.1)
dev.off()

pdf("figures/Metabolite_AraField4_Pairs.pdf", width=15, height=15)
pairs(metabolite_arafield4[, pairs_colnames], panel = panel.smooth, 
      labels = pairs_labels, cex.labels = 2, font.labels = 2, cex.main = 2, 
      cex.axis = 1.5, main = "complete dataset field trial 4", gap = 0.1)
dev.off()
```


# Combine data of all field trials
```{r combine data of all field trials}
combined_colnames <- c("versuch_id","accession", nonNeg_colnames)
combined_colnamesT <- c("versuch_id","accession","glucoseT","fructoseT","sucroseT", 
                        "raffinoseT","amino_acidsT","malateT","fumarateT","prolineT")

# original values
metabolite_arafield_all <- rbind(metabolite_arafield2[,combined_colnames],
                                 metabolite_arafield3[,combined_colnames],
                                 metabolite_arafield4[,combined_colnames]) 
dim(metabolite_arafield_all)

# transformed values
metabolite_arafield_allT <- rbind(metabolite_arafield2[,combined_colnamesT],
                                  metabolite_arafield3[,combined_colnamesT],
                                  metabolite_arafield4[,combined_colnamesT]) 


# modify factor for field trial
table(metabolite_arafield_all$versuch_id)
table(metabolite_arafield_allT$versuch_id)

metabolite_arafield_all$versuch_id <- factor(metabolite_arafield_all$versuch_id, 
                                             levels = c("43", "105", "119"))
metabolite_arafield_allT$versuch_id <- factor(metabolite_arafield_allT$versuch_id, 
                                             levels = c("43", "105", "119"))

levels(metabolite_arafield_all$versuch_id) <- c("field2", "field3", "field4")
levels(metabolite_arafield_allT$versuch_id) <- c("field2", "field3", "field4")
```

## Aggregate and scale all data
+ scaling means: columns (= metabolites) are scaled to mean of zero and sd of 1 (over all samples/rows)
```{r aggregate and scale}
# scale (per column = metabolite)
metabolite_arafield_all_scaled <- scale(metabolite_arafield_all[,3:10])
dim(metabolite_arafield_all)
dim(metabolite_arafield_all_scaled)
# 416 8
rownames(metabolite_arafield_all_scaled) <- metabolite_arafield_all$accession

# aggregate
metabolite_arafield_agg <- aggregate(metabolite_arafield_all[,3:10], 
                                     by = list(metabolite_arafield_all$accession,
                                               metabolite_arafield_all$versuch_id), mean)

# scale aggregated data
metabolite_arafield_agg_scaled <- scale(metabolite_arafield_agg[,3:10])
dim(metabolite_arafield_agg_scaled)
# 203 8
rownames(metabolite_arafield_agg_scaled) <- metabolite_arafield_agg$Group.1 # accessions
```

## Heatmap
```{r heatmap, fig.width=10, fig.height=8}
colors = c(seq(0,5,length=100), seq(5.1,10,length=100), seq(10.1,50,length=100))
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)

heatmap.2(t(as.matrix(metabolite_arafield_all[,3:10])), 
          trace = "none", col = my_palette, breaks = colors,
          main = "original data")

f <- metabolite_arafield_all$versuch_id
f_agg <- metabolite_arafield_agg$Group.2
cols_trials <- c("green", "orange", "black")

colors = c(seq(-6,-1.1,length=100), seq(-1,1,length=100), seq(1.1,6,length=100))
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)

heatmap.2(t(as.matrix(metabolite_arafield_all_scaled)), 
          trace = "none", col = my_palette, breaks = colors,
          symkey = F, Colv = FALSE, dendrogram = "row",
          main = "scaled data (no clustering)", 
          ColSideColors = cols_trials[f], na.color = "grey")

heatmap.2(t(as.matrix(metabolite_arafield_all_scaled)), 
          trace = "none", col = my_palette, breaks = colors,
          main = "scaled data", ColSideColors = cols_trials[f], 
          symkey = F, na.color = "grey")

heatmap.2(t(as.matrix(metabolite_arafield_agg_scaled)), 
          trace = "none", col = my_palette, breaks = colors, 
          symkey = F, Colv = FALSE, dendrogram = "row",
          main = "aggregated & scaled data (no clustering)",
          na.color = "grey")

heatmap.2(t(as.matrix(metabolite_arafield_agg_scaled)), 
          trace = "none", col = my_palette, breaks = colors,
          symkey = F, main = "aggregated & scaled data",
          ColSideColors = cols_trials[f_agg], na.color = "grey")

pdf("figures/Metabolite_AraField_Heatmap_agg_scaled.pdf", width=7, height=12)
heatmap.2(as.matrix(metabolite_arafield_agg_scaled), 
          trace = "none", col = my_palette, breaks = colors,
          symkey = F, main = "aggregated & scaled data", cexRow = 0.5,
          RowSideColors = cols_trials[f_agg], na.color = "grey")
dev.off()
```


## PCA
```{r PCA}
# aggregated scaled
pca_data_agg_scaled <- prep(metabolite_arafield_agg_scaled, scale = "none")
dim(pca_data_agg_scaled)
pca_res_agg_scaled <- pca(pca_data_agg_scaled, method = "rnipals", nPcs = 5)

plot(pca_res_agg_scaled@scores[,1], pca_res_agg_scaled@scores[,2], col = cols_trials[f_agg], pch = 19)
legend("topleft", fill = cols_trials, c("field2", "field3", "field4"))

metabolite_arafield_agg_scaled_lt50 <- merge(metabolite_arafield_agg_scaled, lt50, 
                                             by.x = "row.names", by.y = "Accession", all = TRUE)

# Colour points in a plot differently depending on a vector of values
# http://stackoverflow.com/questions/9946630/colour-points-in-a-plot-differently-depending-on-a-vector-of-values
rbPal <- colorRampPalette(c('red','blue'))
metabolite_arafield_agg_scaled_lt50$Color_ACC <- rbPal(10)[as.numeric(cut(metabolite_arafield_agg_scaled_lt50$LT50ACC,breaks = 10))]

plot(pca_res_agg_scaled@scores[,1], pca_res_agg_scaled@scores[,2], col = metabolite_arafield_agg_scaled_lt50$Color_ACC, pch = 19)



# All scaled
pca_data_all_scaled <- prep(metabolite_arafield_all_scaled, scale = "none")
dim(pca_data_all_scaled)
pca_res_all_scaled <- pca(pca_data_all_scaled, method = "rnipals", nPcs = 5)

plot(pca_res_all_scaled@scores[,1], pca_res_all_scaled@scores[,2], col = cols_trials[f], pch = 19)
legend("topleft", fill = cols_trials, c("field2", "field3", "field4"))

metabolite_arafield_all_scaled_lt50 <- merge(metabolite_arafield_all_scaled, lt50, 
                                             by.x = "row.names", by.y = "Accession", all = TRUE)

# Colour points in a plot differently depending on a vector of values
# http://stackoverflow.com/questions/9946630/colour-points-in-a-plot-differently-depending-on-a-vector-of-values
rbPal <- colorRampPalette(c('red','blue'))
metabolite_arafield_all_scaled_lt50$Color_ACC <- rbPal(10)[as.numeric(cut(metabolite_arafield_all_scaled_lt50$LT50ACC,breaks = 10))]

plot(pca_res_all_scaled@scores[,1], pca_res_all_scaled@scores[,2], col = metabolite_arafield_all_scaled_lt50$Color_ACC, pch = 19)

# original data
pca_data_all <- prep(metabolite_arafield_all[,3:10], scale = "none")
dim(pca_data_all)
pca_res_all <- pca(pca_data_all, method = "rnipals", nPcs = 5)

plot(pca_res_all@scores[,1], pca_res_all@scores[,2], col = cols_trials[f], pch = 19)
legend("topleft", fill = cols_trials, c("field2", "field3", "field4"))

plot(pca_res_all@scores[,1], pca_res_all@scores[,2], col = metabolite_arafield_all_scaled_lt50$Color_ACC, pch = 19)
```


# Save workspace
```{r save workspace}
save.image("metabolite_arafield.RData")
```

