---
title: "Modelling of metabolite data Arabidopsis field trials"
author: "Heike Sprenger"
date: "Monday, July 18, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

# Set working directory  
```{r set working directory}
#getwd()
# setwd("D:/work/repos/ara_field_trials/")
# setwd("X:/mpimp/repos/ara_field_trials")
setwd("E:/uschi-home-sync/uschi/Seafile/uschi-work/repos/ara_field_trials")
```

# Load workspace, packages and scripts
```{r load workspace, include=TRUE, message=FALSE}
# load packages
library(knitr)
library(pander)
library(ggplot2)
library(DAAG)
library(leaps)
library(faraway)
library(MASS)
library(glmnet)
library(caret)
library(tidyverse)
library(ggpubr)
library(lemon) # for coord_capped_cart
library(RColorBrewer)

# set options for knitr
opts_chunk$set(fig.width=6, fig.height=5, tidy=TRUE, tidy.opts=list(width.cutoff=70))
opts_knit$set(root.dir = '../')

# load workspace
#load("metabolite_arafield_model.RData")

pal_grey <- c("black", "grey50", "grey75")
```


# Source functions
```{r source functions}
source("../functions/func_calc_loocv.R", encoding = 'UTF-8')
```


# Load data
Use changed metabolite data, from ``metabolite_arafield4_overview4.Rmd``
```{r load data}
field2 <- read.table("output/metabolite_arafield2_modified.txt", header = TRUE, sep = "\t")
field3 <- read.table("output/metabolite_arafield3_modified.txt", header = TRUE, sep = "\t")
field4 <- read.table("output/metabolite_arafield4_modified.txt", header = TRUE, sep = "\t")

# OLD VERSION
# lt50 <- read.table("data/LT50.txt", header = TRUE, sep = "\t")
# dim(lt50) 
# 58 accessions (= 52 accessions for training + 6 for testing)

# NEW VERSION with 64 accessions
lt50 <- read.table("data/lt50_64acc.txt", header = TRUE, sep = "\t")
dim(lt50)
# 64 accessions (= 52 accessions for training + 12 for testing)
table(lt50$Model)

# capacity
yCapacity4full <- field4$LT50ACC - field4$LT50NAC 
yCapacity <- lt50$LT50ACC - lt50$LT50NAC

setdiff(lt50$Accession, field4$accession) # Ws missing in Field 4
setdiff(lt50$Accession, field3$accession)
setdiff(lt50$Accession, field2$accession)
setdiff(field4$accession, lt50$Accession)
setdiff(field3$accession, lt50$Accession)
setdiff(field2$accession, lt50$Accession)

length(unique(field2$accession))
length(unique(field3$accession))
length(unique(field4$accession))

setdiff(field3$accession, field4$accession)

# Number accessions with known LT50
length(intersect(field2$accession, lt50$Accession))
length(intersect(field3$accession, lt50$Accession))
length(intersect(field4$accession, lt50$Accession))
```


## Set factor levels correctly
```{r set factor levels correctly}
# antho score (en)
field2$antho_value_en <- factor(field2$antho_value_en, 
                                              levels = c("very low", "low", "intermediate", "high", "very high"))
field3$antho_value_en <- factor(field3$antho_value_en, 
                                              levels = c("low", "intermediate", "high", "very high")) # "very low" is not present
field4$antho_value_en <- factor(field4$antho_value_en, 
                                              levels = c("very low", "low", "intermediate", "high", "very high"))

# bbch_id as factor
class(field4$bbch_id)
field2$bbch_factor <- factor(field2$bbch_id)
field3$bbch_factor <- factor(field3$bbch_id)
field4$bbch_factor <- factor(field4$bbch_id)

# bbch categories
field2$bbch_categories <- factor(field2$bbch_categories, levels = c("vegetative", "generative buds", "flowering"))
field3$bbch_categories <- factor(field3$bbch_categories, levels = c("vegetative", "generative buds", "flowering"))
field4$bbch_categories <- factor(field4$bbch_categories, levels = c("vegetative", "generative buds", "flowering"))

# flowering state
field2$flowering_state <- factor(field2$flowering_state, levels = c("non-flowering", "flowering"))
field3$flowering_state <- factor(field3$flowering_state, levels = c("non-flowering", "flowering"))
field4$flowering_state <- factor(field4$flowering_state, levels = c("non-flowering", "flowering"))

# vegetative state
field2$vegetative_state <- factor(field2$vegetative_state, levels = c("vegetative", "non-vegetative"))
field3$vegetative_state <- factor(field3$vegetative_state, levels = c("vegetative", "non-vegetative"))
field4$vegetative_state <- factor(field4$vegetative_state, levels = c("vegetative", "non-vegetative"))
```

## Remove missing values (NAs)
```{r remove missing values (NAs)}
dim(field2)
dim(field3)
dim(field4)

field2 <- field2[which(!is.na(field2$sucrose_positive) ), ]
field3 <- field3[which(!is.na(field3$fumarate_positive) ), ]
# Field4-Tsu, SampleID: 837060, Plant: [At.Tsu-0.n].2.2-4 --> missing values for BBCH and antho-Score! --> remove?
field4 <- field4[which(!is.na(field4$anthoscore) ), ]

dim(field2)
dim(field3)
dim(field4)
```


## Standardize metabolite levels
```{r standardize metabolite levels}
hist(field4$proline_positive)
hist(field4$prolineT)
hist(scale(field4$prolineT))

field2_scaled <- data.frame(field2[,c("accession", "plant_name", "LT50ACC", "LT50NAC", 
                                      "capacity", "bbch_id", "anthoscore", "bbch_factor", "antho_value_en")],
                            scale(field2[,c("glucoseT","fructoseT","sucroseT","raffinoseT",
                                            "amino_acidsT","malateT","fumarateT","prolineT")]))

field3_scaled <- data.frame(field3[,c("accession", "plant_name", "LT50ACC", "LT50NAC", 
                                      "capacity", "bbch_id", "anthoscore", "bbch_factor", "antho_value_en")],
                            scale(field3[,c("glucoseT","fructoseT","sucroseT","raffinoseT",
                                            "amino_acidsT","malateT","fumarateT","prolineT")]))

field4_scaled <- data.frame(field4[,c("accession", "plant_name", "LT50ACC", "LT50NAC", 
                                      "capacity", "bbch_id", "anthoscore", "bbch_factor", "antho_value_en")],
                            scale(field4[,c("glucoseT","fructoseT","sucroseT","raffinoseT",
                                            "amino_acidsT","malateT","fumarateT","prolineT")]))
```



# Define data for modeling
## Define training and test accessions
```{r define training and test accessions}
table(lt50$Model)
accessions_training <- as.character(lt50$Accession[which(lt50$Model == "train")])
accessions_testing <- as.character(lt50$Accession[which(lt50$Model == "test")])
accessions_known <- as.character(lt50$Accession)
accessions_unknown <- as.character(setdiff(field3$accession, lt50$Accession))
sum(length(accessions_known), length(accessions_unknown))

# indices for accessions for training/testing
idx_train2 <- which(field2$accession %in% accessions_training)
idx_train3 <- which(field3$accession %in% accessions_training)
idx_train4 <- which(field4$accession %in% accessions_training)
idx_test2 <- which(field2$accession %in% accessions_testing)
idx_test3 <- which(field3$accession %in% accessions_testing)
idx_test4 <- which(field4$accession %in% accessions_testing)
idx_all2 <- which(field2$accession %in% accessions_known)
idx_all3 <- which(field3$accession %in% accessions_known)
idx_all4 <- which(field4$accession %in% accessions_known)

idx_unknown2 <- which(field2$accession %in% accessions_unknown)
idx_unknown3 <- which(field3$accession %in% accessions_unknown)
idx_unknown4 <- which(field4$accession %in% accessions_unknown)
```

## Define colnames
```{r define colnames}
# with original BBCH
model_colnames <- c("accession", "plant_name", "LT50ACC", "LT50NAC", "capacity",
                    "glucose_positive", "fructose_positive", "sucrose_positive",
                    "raffinose_positive", "amino_acids_positive", "malate_positive", 
                    "fumarate_positive", "proline_positive", "bbch_id", "anthoscore")

# BBCH and antho_value as factors
model_colnames_factor <- c("accession", "plant_name", "LT50ACC", "LT50NAC", "capacity",
                    "glucose_positive", "fructose_positive", "sucrose_positive",
                    "raffinose_positive", "amino_acids_positive", "malate_positive", 
                    "fumarate_positive", "proline_positive", "bbch_factor", "antho_value_en")

# with NORMALIZED BBCH
model_colnames_norm <- c("accession", "plant_name", "LT50ACC", "LT50NAC", "capacity",
                    "glucose_positive", "fructose_positive", "sucrose_positive",
                    "raffinose_positive", "amino_acids_positive", "malate_positive", 
                    "fumarate_positive", "proline_positive", "norm_bbch_id", "anthoscore")

model_colnames_short <- c("accession", "plant_name", "LT50ACC", "LT50NAC", "cap",
                          "glc","frc","suc","raf","aa","mal","fum","pro","bbch","anthoscore")

# transformed and scaled variables
model_colnamesT <- c("accession", "plant_name", "LT50ACC", "LT50NAC", "capacity",
                    "glucoseT", "fructoseT", "sucroseT","raffinoseT", "amino_acidsT", 
                    "malateT", "fumarateT", "prolineT", "bbch_id", "anthoscore")

model_colnamesT_factor <- model_colnamesT
model_colnamesT_factor[14:15] <- c("bbch_factor", "antho_value_en")
```


### Subset for model training/testing
```{r subset for model training/testing}
# training set
field2_train <- field2[idx_train2, model_colnames]
field3_train <- field3[idx_train3, model_colnames]
field4_train <- field4[idx_train4, model_colnames]
field2_trainT <- field2_scaled[idx_train2, model_colnamesT]
field3_trainT <- field3_scaled[idx_train3, model_colnamesT]
field4_trainT <- field4_scaled[idx_train4, model_colnamesT]
field2_trainT_fac <- field2_scaled[idx_train2, model_colnamesT_factor]
field3_trainT_fac <- field3_scaled[idx_train3, model_colnamesT_factor]
field4_trainT_fac <- field4_scaled[idx_train4, model_colnamesT_factor]
field2_train_norm <- field2[idx_train2, model_colnames_norm]
field3_train_norm <- field3[idx_train3, model_colnames_norm]
field4_train_norm <- field4[idx_train4, model_colnames_norm]

# testing set
field2_test <- field2[idx_test2, model_colnames]
field3_test <- field3[idx_test3, model_colnames]
field4_test <- field4[idx_test4, model_colnames]
field2_testT <- field2_scaled[idx_test2, model_colnamesT]
field3_testT <- field3_scaled[idx_test3, model_colnamesT]
field4_testT <- field4_scaled[idx_test4, model_colnamesT]
field2_testT_fac <- field2_scaled[idx_test2, model_colnamesT_factor]
field3_testT_fac <- field3_scaled[idx_test3, model_colnamesT_factor]
field4_testT_fac <- field4_scaled[idx_test4, model_colnamesT_factor]
field2_test_norm <- field2[idx_test2, model_colnames_norm]
field3_test_norm <- field3[idx_test3, model_colnames_norm]
field4_test_norm <- field4[idx_test4, model_colnames_norm]

# all accessions
field2_all <- field2[idx_all2, model_colnames]
field3_all <- field3[idx_all3, model_colnames]
field4_all <- field4[idx_all4, model_colnames]
field2_all_fac <- field2[idx_all2, model_colnames_factor]
field3_all_fac <- field3[idx_all3, model_colnames_factor]
field4_all_fac <- field4[idx_all4, model_colnames_factor]
field2_allT_fac <- field2_scaled[idx_all2, model_colnamesT_factor]
field3_allT_fac <- field3_scaled[idx_all3, model_colnamesT_factor]
field4_allT_fac <- field4_scaled[idx_all4, model_colnamesT_factor]


# only accessions with unknown LT50
field2_unknown <- field2[idx_unknown2, model_colnames]
field3_unknown <- field3[idx_unknown3, model_colnames]
field4_unknown <- field4[idx_unknown4, model_colnames]
field2_unknown_fac <- field2[idx_unknown2, model_colnames_factor]
field3_unknown_fac <- field3[idx_unknown3, model_colnames_factor]
field4_unknown_fac <- field4[idx_unknown4, model_colnames_factor]

# remove NA samples
field2_train <- field2_train[which(!is.na(field2_train$sucrose_positive) ), ]
field3_train <- field3_train[which(!is.na(field3_train$fumarate_positive) ), ]
field3_test <- field3_test[which(!is.na(field3_test$fumarate_positive) ), ]

# Field4-Tsu, SampleID: 837060, Plant: [At.Tsu-0.n].2.2-4 --> missing values for BBCH and antho-Score! --> remove?
field4_train <- field4_train[which(!is.na(field4_train$anthoscore) ), ]

length(unique(field2_train$accession)) # 42 
length(unique(field3_train$accession)) # 52/48 (with/without NAs for fumarate!)
length(unique(field4_train$accession)) # 51

length(unique(field2_test$accession)) # 6
length(unique(field3_test$accession)) # 12/9 (with/without NAs for fumarate)
length(unique(field4_test$accession)) # 12
```


## Data splitting NEW
```{r data splitting NEW}
n_obs <- nrow(lt50) # 64
source("../functions/func_get_splits.R")
set.seed(1234)
folds <- func_get_splits(seq = sample(1:n_obs), num = 8)

folds_acc <- folds

for(i in 1:8){
  for(j in 1:8){
    folds_acc[i,j] <- as.character(lt50$Accession)[folds[i,j]]
  }
}

# trainsets will be a list of 8 dataframes
x_f4_train <- x_f4_test <- y_f4_train <- y_f4_test <- list()

for(i in 1:8){
  idx_train4 <- which(field4$accession %in% c(folds_acc[-i,]))
  idx_test4 <- which(field4$accession %in% folds_acc[i,])
  
  x_f4_train[[i]] <- field4[idx_train4, model_colnames[6:15]]
  x_f4_test[[i]] <- field4[idx_test4, model_colnames[6:15]]
  y_f4_train[[i]] <- field4[idx_train4, "LT50ACC"]
  y_f4_test[[i]] <- field4[idx_test4, "LT50ACC"]
}

dim(f4_train[[1]])
# 119 rows = observations, 15 columns

dim(f4_test[[1]])
# 16 rows = observations, 15 columns
```


## Define freezing tolerance per trial
```{r define freezing tolerance per trial}
# field2_train_LT50ACC <- field2_train$LT50ACC
# field3_train_LT50ACC <- field3_train$LT50ACC
# field4_train_LT50ACC <- field4_train$LT50ACC
# 
# yLT50NAC2 <- field2_train$LT50NAC
# yLT50NAC3 <- field3_train$LT50NAC
# yLT50NAC4 <- field4_train$LT50NAC
# 
# yCapacity2 <- field2_train$capacity
# yCapacity3 <- field3_train$capacity
# yCapacity4 <- field4_train$capacity
```


## Change colnames and export data
- using ORIGINAL/NORMALIZED BBCH!
```{r change colnames and export data}
# training data
colnames(field2_train) <- colnames(field3_train) <- colnames(field4_train) <- model_colnames_short
# using NORMALIZED BBCH
colnames(field2_train_norm) <- colnames(field3_train_norm) <- colnames(field4_train_norm) <- model_colnames_short
# transformed and scaled data
colnames(field2_trainT) <- colnames(field3_trainT) <- colnames(field4_trainT) <- model_colnames_short
colnames(field2_trainT_fac) <- colnames(field3_trainT_fac) <- colnames(field4_trainT_fac) <- model_colnames_short

# test data
colnames(field2_test) <- colnames(field3_test) <- colnames(field4_test) <- model_colnames_short
# using NORMALIZED BBCH
colnames(field2_test_norm) <- colnames(field3_test_norm) <- colnames(field4_test_norm) <- model_colnames_short
# transformed and scaled data
colnames(field2_testT) <- colnames(field3_testT) <- colnames(field4_testT) <- model_colnames_short
colnames(field2_testT_fac) <- colnames(field3_testT_fac) <- colnames(field4_testT_fac) <- model_colnames_short

# all accessions
colnames(field2_all) <- colnames(field3_all) <- colnames(field4_all) <- model_colnames_short
colnames(field2_all_fac) <- colnames(field3_all_fac) <- colnames(field4_all_fac) <- model_colnames_short
colnames(field2_allT_fac) <- colnames(field3_allT_fac) <- colnames(field4_allT_fac) <- model_colnames_short

# data for accessions of unknown LT50
colnames(field2_unknown) <- colnames(field3_unknown) <- colnames(field4_unknown) <- model_colnames_short
colnames(field2_unknown_fac) <- colnames(field3_unknown_fac) <- colnames(field4_unknown_fac) <- model_colnames_short

# training data
write.table(field2_train, "output/metabolite_arafield2_train.txt", sep="\t")
write.table(field3_train, "output/metabolite_arafield3_train.txt", sep="\t")
write.table(field4_train, "output/metabolite_arafield4_train.txt", sep="\t")
write.table(field2_train_norm, "output/metabolite_arafield2_train_norm.txt", sep="\t")
write.table(field3_train_norm, "output/metabolite_arafield3_train_norm.txt", sep="\t")
write.table(field4_train_norm, "output/metabolite_arafield4_train_norm.txt", sep="\t")
# test data
write.table(field2_test, "output/metabolite_arafield2_test.txt", sep="\t")
write.table(field3_test, "output/metabolite_arafield3_test.txt", sep="\t")
write.table(field4_test, "output/metabolite_arafield4_test.txt", sep="\t")
write.table(field2_test_norm, "output/metabolite_arafield2_test_norm.txt", sep="\t")
write.table(field3_test_norm, "output/metabolite_arafield3_test_norm.txt", sep="\t")
write.table(field4_test_norm, "output/metabolite_arafield4_test_norm.txt", sep="\t")
```


# Linear Regression of LT50ACC (Training)
## Use Normalized BBCH
* Unterschied zwischen anova- und summary-Signifikanzen: erstere geben an, ob ein Effekt signifikant zur Erklaerung der Gesamtvarianz beitraegt.
* zweitere ist "nur" ein Test, ob die Steigung signifikant unterschiedlich von null ist
```{r regACC norm BBCH}
field2_regressionACC_norm <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field2_train_norm)
field3_regressionACC_norm <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field3_train_norm)
field4_regressionACC_norm <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field4_train_norm)
summary(field2_regressionACC_norm)$r.squared # Multiple R-squared: 0.356
summary(field3_regressionACC_norm)$r.squared # Multiple R-squared: 0.298
summary(field4_regressionACC_norm)$r.squared # Multiple R-squared: 0.361   

# use FACTOR for antho_score and norm_bbch
field2_regressionACC_norm_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                       factor(bbch) + factor(anthoscore), data = field2_train_norm)
field3_regressionACC_norm_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                       factor(bbch) + factor(anthoscore), data = field3_train_norm)
field4_regressionACC_norm_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                       factor(bbch) + factor(anthoscore), data = field4_train_norm)
summary(field2_regressionACC_norm_factor)$r.squared # Multiple R-squared: 0.449
summary(field3_regressionACC_norm_factor)$r.squared # Multiple R-squared: 0.358
summary(field4_regressionACC_norm_factor)$r.squared # Multiple R-squared: 0.431 
```


## Use original BBCH - without factor
```{r regACC original BBCH - without factor}
field2_regressionACC <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                               bbch + anthoscore, data = field2_train)
field3_regressionACC <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field3_train)
field4_regressionACC <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field4_train)

summary(field2_regressionACC)$r.squared # 0.350
summary(field3_regressionACC)$r.squared # 0.286
summary(field4_regressionACC)$r.squared # 0.405 
```


## Use original BBCH - with factor
```{r regACC original BBCH - with factor}
field2_regressionACC_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    factor(bbch) + factor(anthoscore), data = field2_train)
field3_regressionACC_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    factor(bbch) + factor(anthoscore), data = field3_train)
field4_regressionACC_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    factor(bbch) + factor(anthoscore), data = field4_train)

summary(field2_regressionACC_factor)$r.squared # 0.449
summary(field3_regressionACC_factor)$r.squared # 0.358
summary(field4_regressionACC_factor)$r.squared # 0.431
summary(field4_regressionACC_factor)$adj.r.squared # Adjusted R-squared: 0.301

## --> identical to model with factor of normalized scores
```

## Use transformed and scaled data (without factor)
```{r regACC transformed and scaled data}
field2T_regressionACC <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                               bbch + anthoscore, data = field2_trainT)
field3T_regressionACC <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field3_trainT)
field4T_regressionACC <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field4_trainT)

summary(field2T_regressionACC)$r.squared # 0.377
summary(field3T_regressionACC)$r.squared # 0.273
summary(field4T_regressionACC)$r.squared # 0.399
```


## Use transformed and scaled data (with factor)
```{r regACC transformed and scaled data - with factor}
field2T_regressionACC_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                     bbch + anthoscore, data = field2_trainT_fac)
field3T_regressionACC_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                     bbch + anthoscore, data = field3_trainT_fac)
field4T_regressionACC_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                     bbch + anthoscore, data = field4_trainT_fac)

summary(field2T_regressionACC_factor)$r.squared # 0.481
summary(field3T_regressionACC_factor)$r.squared # 0.350
summary(field4T_regressionACC_factor)$r.squared # 0.428

cor(field4T_regressionACC_factor$fitted.values, field4_trainT$LT50ACC) # R = 0.654
#predict(field4T_regressionACC_factor, field4_trainT[,6:15]) # equivalent to fitted.values
```

## Combine predictions across trials
```{r combine training predictions across trials}
pred_df2_train <- data.frame(data = "trial2",
                       accession = field2_trainT_fac$accession,
                       measured = field2_trainT_fac$LT50ACC,
                       predicted = field2T_regressionACC_factor$fitted.values)

pred_df3_train <- data.frame(data = "trial3",
                       accession = field3_trainT_fac$accession,
                       measured = field3_trainT_fac$LT50ACC,
                       predicted = field3T_regressionACC_factor$fitted.values)

pred_df4_train <- data.frame(data = "trial4",
                       accession = field4_trainT_fac$accession,
                       measured = field4_trainT_fac$LT50ACC,
                       predicted = field4T_regressionACC_factor$fitted.values)

pred_df_train <- rbind(pred_df2_train, pred_df3_train, pred_df4_train)
```

## Correlation scatter plot
```{r correlation training scatter plot}
range(pred_df_train$predicted)
range(pred_df_train$measured)
lim_values <- c(-12.5, -5)

sp_train <- ggscatter(pred_df_train, x = "measured", y = "predicted",
                add = "reg.line", 
                add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                conf.int = TRUE)

sp_train_lim <- ggpar(sp_train, ylim = lim_values, xlim = lim_values)
sp_train_facet <- facet(sp_train_lim + theme_bw(), facet.by = "data")
sp_train_final <- sp_train_facet + stat_cor(method = "pearson", label.x.npc = 'center', label.y.npc = 'top')

ggsave(filename = "figures/predictions_train_corr_plot.png", 
       plot = sp_train_final, width = 10, height = 4, dpi = 600)
```


# Linear Regression of LT50ACC (All Data)
## Use original BBCH - without factor
```{r regACC_all original BBCH - without factor}
field2_regressionACC_all <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    bbch + anthoscore, data = field2_all)
field3_regressionACC_all <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    bbch + anthoscore, data = field3_all)
field4_regressionACC_all <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    bbch + anthoscore, data = field4_all)

summary(field2_regressionACC_all)$r.squared # 0.367
summary(field3_regressionACC_all)$r.squared # 0.279
summary(field4_regressionACC_all)$r.squared # 0.384
summary(field4_regressionACC_all)$adj.r.squared # Adjusted R-squared: 0.334
```

## Use original BBCH - with factor
```{r regACC_all original BBCH - with factor}
field2_regressionACC_all_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    bbch + anthoscore, data = field2_all_fac)
field3_regressionACC_all_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    bbch + anthoscore, data = field3_all_fac)
field4_regressionACC_all_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                    bbch + anthoscore, data = field4_all_fac)

summary(field2_regressionACC_all_factor)$r.squared # 0.479
summary(field3_regressionACC_all_factor)$r.squared # 0.394
summary(field4_regressionACC_all_factor)$r.squared # 0.412
summary(field4_regressionACC_all_factor)$adj.r.squared # Adjusted R-squared: 0.309
```

## Use transformed and scaled data (with factor)
```{r regACC transformed and scaled data - with factor}
field2T_regressionACC_all_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                     bbch + anthoscore, data = field2_allT_fac)
field3T_regressionACC_all_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                     bbch + anthoscore, data = field3_allT_fac)
field4T_regressionACC_all_factor <- lm(LT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                     bbch + anthoscore, data = field4_allT_fac)

summary(field2T_regressionACC_all_factor)$r.squared # 0.500
summary(field3T_regressionACC_all_factor)$r.squared # 0.394
summary(field4T_regressionACC_all_factor)$r.squared # 0.410

cor(field4T_regressionACC_all_factor$fitted.values, field4_allT_fac$LT50ACC) # R = 0.640
#predict(field4T_regressionACC_all_factor, field4_allt_fac[,6:15]) # equivalent to fitted.values
```

# Apply models to different datasets (other trials)
```{r apply models to different datasets ALL}
field2T_regressionACC_all_factor$xlevels$bbch <- union(field2T_regressionACC_all_factor$xlevels$bbch,
                                                   union(levels(field3_allT_fac$bbch),
                                                         levels(field4_allT_fac$bbch)))
field3T_regressionACC_all_factor$xlevels$bbch <- union(field3T_regressionACC_all_factor$xlevels$bbch,
                                                   levels(field4_allT_fac$bbch))
field4T_regressionACC_all_factor$xlevels$bbch <- union(field4T_regressionACC_all_factor$xlevels$bbch,
                                                   levels(field3_allT_fac$bbch))


field3T_regressionACC_all_factor$xlevels$anthoscore <- union(field3T_regressionACC_all_factor$xlevels$anthoscore,
                                                             levels(field4_allT_fac$anthoscore))

field3T_regressionACC_all_factor_pred2 <- predict(field3T_regressionACC_all_factor, field2_allT_fac[,6:15]) # 3>2
field3T_regressionACC_all_factor_pred4 <- predict(field3T_regressionACC_all_factor, field4_allT_fac[,6:15]) # 3>4
field4T_regressionACC_all_factor_pred2 <- predict(field4T_regressionACC_all_factor, field2_allT_fac[,6:15]) # 4>2
field4T_regressionACC_all_factor_pred3 <- predict(field4T_regressionACC_all_factor, field3_allT_fac[,6:15]) # 4>3
field2T_regressionACC_all_factor_pred3 <- predict(field2T_regressionACC_all_factor, field3_allT_fac[,6:15]) # 2>3
field2T_regressionACC_all_factor_pred4 <- predict(field2T_regressionACC_all_factor, field4_allT_fac[,6:15]) # 2>4
```

```{r apply models to different datasets}
field2T_regressionACC_factor$xlevels$bbch <- union(field2T_regressionACC_factor$xlevels$bbch,
                                                   union(levels(field3_trainT_fac$bbch),
                                                         levels(field4_trainT_fac$bbch)))
field4T_regressionACC_factor$xlevels$bbch <- union(field4T_regressionACC_factor$xlevels$bbch,
                                                   levels(field3_trainT_fac$bbch))
field3T_regressionACC_factor$xlevels$bbch <- union(field3T_regressionACC_factor$xlevels$bbch,
                                                   levels(field4_trainT_fac$bbch))
field3T_regressionACC_factor$xlevels$anthoscore <- union(field3T_regressionACC_factor$xlevels$anthoscore,
                                                         levels(field4_trainT_fac$anthoscore))

field3T_regressionACC_factor_pred2 <- predict(field3T_regressionACC_factor, field2_trainT_fac[,6:15]) # 3>2
field3T_regressionACC_factor_pred4 <- predict(field3T_regressionACC_factor, field4_trainT_fac[,6:15]) # 3>4
field4T_regressionACC_factor_pred2 <- predict(field4T_regressionACC_factor, field2_trainT_fac[,6:15]) # 4>2
field4T_regressionACC_factor_pred3 <- predict(field4T_regressionACC_factor, field3_trainT_fac[,6:15]) # 4>3
field2T_regressionACC_factor_pred3 <- predict(field2T_regressionACC_factor, field3_trainT_fac[,6:15]) # 2>3
field2T_regressionACC_factor_pred4 <- predict(field2T_regressionACC_factor, field4_trainT_fac[,6:15]) # 2>4
```

## Plot predicted vs. measured LT50ACC (across trials)
```{r plot predicted vs. measured LT50ACC (across trials)}
plot(field2_trainT$LT50ACC, field3T_regressionACC_factor_pred2)
cor(field2_trainT$LT50ACC, field3T_regressionACC_factor_pred2) # 0.45
cor(field2_allT_fac$LT50ACC, field3T_regressionACC_all_factor_pred2) # 0.48

plot(field4_trainT$LT50ACC, field3T_regressionACC_factor_pred4)
cor(field4_trainT$LT50ACC, field3T_regressionACC_factor_pred4) # 0.485
cor(field4_allT_fac$LT50ACC, field3T_regressionACC_all_factor_pred4) # 0.48

plot(field3_trainT$LT50ACC, field4T_regressionACC_factor_pred3)
cor(field3_trainT$LT50ACC, field4T_regressionACC_factor_pred3) # 0.413
cor(field3_allT_fac$LT50ACC, field4T_regressionACC_all_factor_pred3) # 0.46

plot(field2_trainT$LT50ACC, field4T_regressionACC_factor_pred2)
cor(field2_trainT$LT50ACC, field4T_regressionACC_factor_pred2)

plot(field3_trainT$LT50ACC, field2T_regressionACC_factor_pred3)
cor(field3_trainT$LT50ACC, field2T_regressionACC_factor_pred3)

plot(field4_trainT$LT50ACC, field2T_regressionACC_factor_pred4)
cor(field4_trainT$LT50ACC, field2T_regressionACC_factor_pred4)
```

## Combine predictions across trials
```{r combine predictions across trials}
pred_df2 <- data.frame(data = "trial2",
                       accession = field2_allT_fac$accession,
                       measured = field2_allT_fac$LT50ACC,
                       model2 = field2T_regressionACC_all_factor$fitted.values,
                       model3 = field3T_regressionACC_all_factor_pred2,
                       model4 = field4T_regressionACC_all_factor_pred2)

pred_df3 <- data.frame(data = "trial3",
                       accession = field3_allT_fac$accession,
                       measured = field3_allT_fac$LT50ACC,
                       model2 = field2T_regressionACC_all_factor_pred3,
                       model3 = field3T_regressionACC_all_factor$fitted.values,
                       model4 = field4T_regressionACC_all_factor_pred3)

pred_df4 <- data.frame(data = "trial4",
                       accession = field4_allT_fac$accession,
                       measured = field4_allT_fac$LT50ACC,
                       model2 = field2T_regressionACC_all_factor_pred4,
                       model3 = field3T_regressionACC_all_factor_pred4,
                       model4 = field4T_regressionACC_all_factor$fitted.values)

pred_df_all <- rbind(pred_df2, pred_df3, pred_df4)

pred_df_all_ <- 
  pred_df_all %>%
  gather(key = "model", value = "predicted", 4:6)
```

## Correlation scatter plot
```{r correlation scatter plot}
range(pred_df_all_$predicted)
lim_values <- c(-12.5, -4)

sp <- ggscatter(pred_df_all_, x = "measured", y = "predicted",
                add = "reg.line", 
                add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                conf.int = TRUE)

sp_lim <- ggpar(sp, ylim = lim_values, xlim = lim_values)

sp_facet <- facet(sp_lim + theme_bw(), facet.by = c("model","data"))

sp_final <- sp_facet + stat_cor(method = "pearson", label.x.npc = 'center', label.y.npc = 'top')

ggsave(filename = "figures/predictions_all_corr_plot.png", plot = sp_final, width = 7, height = 7)
```


# Cross-validation
## 5-fold CV
```{r cross-validation}
field2_DAAGdata <- field2_train[,c(3:15)]
field3_DAAGdata <- field3_train[,c(3:15)]
field4_DAAGdata <- field4_train[,c(3:15)]

field2_DAAGdata_all <- field2_all[,c(3:15)]
field3_DAAGdata_all <- field3_all[,c(3:15)]
field4_DAAGdata_all <- field4_all[,c(3:15)]

sum(is.na(field2_DAAGdata)) # 0
sum(is.na(field3_DAAGdata)) # 0
sum(is.na(field4_DAAGdata)) # 0

field2_DAAGdataT_fac <- field2_trainT_fac[,c(3:15)]
field3_DAAGdataT_fac <- field3_trainT_fac[,c(3:15)]
field4_DAAGdataT_fac <- field4_trainT_fac[,c(3:15)]

# 5-fold cross-validation
field2_5cv <- CVlm(data = field2_DAAGdata, form.lm = field2_regressionACC, m = 5) # overall ms: 3.55
field3_5cv <- CVlm(data = field3_DAAGdata, form.lm = field3_regressionACC, m = 5) # overall ms: 2.77
field4_5cv <- CVlm(data = field4_DAAGdata, form.lm = field4_regressionACC, m = 5) # Overall ms: 2.4

# field2_5cvT <- CVlm(data = field2_DAAGdataT_fac, form.lm = field2T_regressionACC_factor, m = 5) 
# field3_5cvT <- CVlm(data = field3_DAAGdataT_fac, form.lm = field3T_regressionACC_factor, m = 5) 
# field4_5cvT <- CVlm(data = field4_DAAGdataT_fac, form.lm = field4T_regressionACC_factor, m = 5) 
# not possible with factors!
```

## Combine 5-fold CV results across trials
```{r combine 5-fold CV results across trials}
# field2_5cv_df <- data.frame(data = "trial2",
#                        accession = field2_train$accession,
#                        measured = field2_5cv$LT50ACC,
#                        predicted = )
# 
# pred_df3 <- data.frame(data = "trial3",
#                        accession = field3_train$accession,
#                        measured = field3_5cv$LT50ACC,
#                        predicted = )
# 
# pred_df4 <- data.frame(data = "trial4",
#                        accession = field4_train$accession,
#                        measured = field4_5cv$LT50ACC,
#                        predicted = field4_5cv$Predicted)
```

## Correlation scatter plot
```{r correlation scatter plot}
# range(pred_df_all_$predicted)
# lim_values <- c(-12.5, -4)
# 
# sp <- ggscatter(pred_df_all_, x = "measured", y = "predicted",
#                 add = "reg.line", 
#                 add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
#                 conf.int = TRUE)
# 
# sp_lim <- ggpar(sp, ylim = lim_values, xlim = lim_values)
# 
# sp_facet <- facet(sp_lim + theme_bw(), facet.by = c("model","data"))
# 
# sp_final <- sp_facet + stat_cor(method = "pearson", label.x.npc = 'center', label.y.npc = 'top')
# 
# ggsave(filename = "figures/predictions_all_corr_plot.png", plot = sp_final, width = 7, height = 7)
```


## Leave-One-Out-cross-validation (LOO-CV)
```{r Leave-One-Out-cross-validation (LOO-CV), warning=FALSE, message=FALSE}
# training data
field2_loocv <- CVlm(data = field2_DAAGdata, form.lm = field2_regressionACC, m = 76, printit = F)
attr(field2_loocv, "ms") # Overall ms: 3.19
plot(field2_loocv$LT50ACC, field2_loocv$cvpred)


field3_loocv <- CVlm(data = field3_DAAGdata, form.lm = field3_regressionACC, m = 97, printit = F)
attr(field3_loocv, "ms") # Overall ms: 2.64
plot(field3_loocv$LT50ACC, field3_loocv$cvpred)


field4_loocv <- CVlm(data = field4_DAAGdata, form.lm = field4_regressionACC, m = 109, printit = F)
attr(field4_loocv, "ms") # Overall ms: 2.31
cor.test(field4_loocv$LT50ACC, field4_loocv$cvpred)
plot(field4_loocv$LT50ACC, field4_loocv$Predicted)
#field4_loocv$Predicted equals field4_regressionACC$fitted.values


# ALL data
field2_all_loocv <- CVlm(data = field2_DAAGdata_all, form.lm = field2_regressionACC_all, m = 87, printit = F)
attr(field2_all_loocv, "ms") # Overall ms: 2.89
plot(field2_all_loocv$LT50ACC, field2_all_loocv$cvpred)

field3_all_loocv <- CVlm(data = field3_DAAGdata_all, form.lm = field3_regressionACC_all, m = 114, printit = F)
attr(field3_all_loocv, "ms") # Overall ms: 2.55
plot(field3_all_loocv$LT50ACC, field3_all_loocv$cvpred)
abline(0,1)

field4_all_loocv <- CVlm(data = field4_DAAGdata_all, form.lm = field4_regressionACC_all, m = 135, printit = F)
attr(field4_all_loocv, "ms") # Overall ms: 2.19
plot(field4_all_loocv$LT50ACC, field4_all_loocv$cvpred)
abline(0,1)
```

```{r}
field2_loocvT <- CVlm(data = field2_DAAGdataT_fac, form.lm = field2T_regressionACC_factor, m = 76, printit = F)
field3_loocvT <- CVlm(data = field3_DAAGdataT_fac, form.lm = field3T_regressionACC_factor, m = 108, printit = F)
field4_loocvT <- CVlm(data = field4_DAAGdataT_fac, form.lm = field4T_regressionACC_factor, m = 109, printit = F)
```


# Prediction of LT50ACC
## Prediction of LT50ACC with model (for accessions that are included in the model = training data)
```{r prediction of LT50ACC using factor model with ORIGINAL BBCH}
cor.test(field2_regressionACC_factor$fitted.values, field2_train$LT50ACC) # R = 0.67
cor.test(field3_regressionACC_factor$fitted.values, field3_train$LT50ACC) # R = 0.598
cor.test(field4_regressionACC_factor$fitted.values, field4_train$LT50ACC) # R = 0.656 / R2 = 0.43

plot(field2_train$LT50ACC, field2_regressionACC_factor$fitted.values, 
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), 
     pch = 19)
abline(lm(field2_regressionACC_factor$fitted.values ~ field2_train$LT50ACC))

plot(field3_train$LT50ACC, field3_regressionACC_factor$fitted.values, 
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), 
     pch = 19)
abline(lm(field3_regressionACC_factor$fitted.values ~ field3_train$LT50ACC))

plot(field4_train$LT50ACC, field4_regressionACC_factor$fitted.values, 
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), pch = 19)
abline(lm(field4_regressionACC_factor$fitted.values ~ field4_train$LT50ACC))
```


## Prediction of LT50ACC with model (for accessions that are included in the model = training data)
```{r prediction of LT50ACC using model with ORIGINAL BBCH}
cor.test(field2_regressionACC$fitted.values, field2_train$LT50ACC) # R = 0.592
cor.test(field3_regressionACC$fitted.values, field3_train$LT50ACC) # R = 0.535
cor.test(field4_regressionACC$fitted.values, field4_train$LT50ACC) # R = 0.636

plot(field2_train$LT50ACC, field2_regressionACC$fitted.values, 
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), 
     pch = 19)
abline(lm(field2_regressionACC$fitted.values ~ field2_train$LT50ACC))

plot(field3_train$LT50ACC, field3_regressionACC_factor$fitted.values, 
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), 
     pch = 19)
abline(lm(field3_regressionACC$fitted.values ~ field3_train$LT50ACC))

plot(field4_train$LT50ACC, field4_regressionACC$fitted.values, 
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), pch = 19)
abline(lm(field4_regressionACC$fitted.values ~ field4_train$LT50ACC))
```


## Prediction of LT50ACC test data
```{r prediction of LT50ACC for test data}
# Vorhersage der LT50-Werte fuer ausgewaehlte Akzessionen mit Hilfe des berechneten Regressionsmodells
field2_predictionACC <- predict(field2_regressionACC, field2_test[,6:15])
field3_predictionACC <- predict(field3_regressionACC, field3_test[,6:15])
field4_predictionACC <- predict(field4_regressionACC, field4_test[,6:15])

field2T_predictionACC <- predict(field2T_regressionACC, field2_testT[,6:15])
field3T_predictionACC <- predict(field3T_regressionACC, field3_testT[,6:15])
field4T_predictionACC <- predict(field4T_regressionACC, field4_testT[,6:15])

field2T_predictionACC_factor <- predict(field2T_regressionACC_factor, field2_testT[,6:15])
field3T_predictionACC_factor <- predict(field3T_regressionACC_factor, field3_testT[,6:15])
field4T_predictionACC_factor <- predict(field4T_regressionACC_factor, field4_testT[,6:15])

field2_predictionACC_factor <- predict(field2_regressionACC_factor, field2_test[,6:15])
field3_predictionACC_factor <- predict(field3_regressionACC_factor, field3_test[,6:15])
field4_predictionACC_factor <- predict(field4_regressionACC_factor, field4_test[,6:15])

field2_predictionACC_norm <- predict(field2_regressionACC_norm, field2_test_norm[,6:15])
field3_predictionACC_norm <- predict(field3_regressionACC_norm, field3_test_norm[,6:15])
field4_predictionACC_norm <- predict(field4_regressionACC_norm, field4_test_norm[,6:15])

field2_predictionACC_norm_factor <- predict(field2_regressionACC_norm_factor, field2_test_norm[,6:15])
field3_predictionACC_norm_factor <- predict(field3_regressionACC_norm_factor, field3_test_norm[,6:15])
field4_predictionACC_norm_factor <- predict(field4_regressionACC_norm_factor, field4_test_norm[,6:15])
```

### Tables of predicted LT50ACC for test data
```{r tables of predicted LT50ACC for test data}
field2_predictionACC_table <- data.frame(accession = field2_test$accession,
                                         measured = field2_test$LT50ACC,
                                         predicted = field2_predictionACC,
                                         predicted_factor = field2_predictionACC_factor,
                                         predicted_norm = field2_predictionACC_norm)

field3_predictionACC_table <- data.frame(accession = field3_test$accession,
                                         measured = field3_test$LT50ACC,
                                         predicted = field3_predictionACC,
                                         predicted_factor = field3_predictionACC_factor,
                                         predicted_norm = field3_predictionACC_norm)

field4_predictionACC_table <- data.frame(accession = field4_test$accession,
                                         measured = field4_test$LT50ACC,
                                         predicted = field4_predictionACC,
                                         predicted_factor = field4_predictionACC_factor,
                                         predicted_norm = field4_predictionACC_norm)

head(field2_predictionACC_table)
head(field3_predictionACC_table)
head(field4_predictionACC_table)

write.table(field2_predictionACC_table, "output/field2_predictionACC.txt", sep="\t")
write.table(field3_predictionACC_table, "output/field3_predictionACC.txt", sep="\t")
write.table(field4_predictionACC_table, "output/field4_predictionACC.txt", sep="\t")
```


### Correlation of predicted and measured LT50ACC for test data
```{r correlation of predicted and measured LT50ACC for test data}
# field 2
cor(field2_test$LT50ACC, field2_predictionACC)
cor(field2_test$LT50ACC, field2T_predictionACC)
cor(field2_test$LT50ACC, field2T_predictionACC_factor)
cor(field2_test$LT50ACC, field2_predictionACC_factor)
cor(field2_test$LT50ACC, field2_predictionACC_norm)
cor(field2_test$LT50ACC, field2_predictionACC_norm_factor)

# field 3
cor(field3_test$LT50ACC, field3_predictionACC)
cor(field3_test$LT50ACC, field3T_predictionACC_factor)
cor(field3_test$LT50ACC, field3_predictionACC_factor)
cor(field3_test$LT50ACC, field3_predictionACC_norm)
cor(field3_test$LT50ACC, field3_predictionACC_norm_factor)

# field 4
cor(field4_test$LT50ACC, field4_predictionACC)
cor(field4_test$LT50ACC, field4T_predictionACC_factor)
cor(field4_test$LT50ACC, field4_predictionACC_factor)
cor(field4_test$LT50ACC, field4_predictionACC_norm)
cor(field4_test$LT50ACC, field4_predictionACC_norm_factor)
```


### Difference between predicted and measured LT50 for test data
```{r difference between predicted and measured LT50 for test data}
sum(((field4_predictionACC)-(field4_test$LT50ACC))^2)
# 52.6
sum(((field4_predictionACC_norm)-(field4_test$LT50ACC))^2)
# 64.9
sum(((field4_predictionACC_factor)-(field4_test$LT50ACC))^2)
# 55.3

sum(((field4_predictionACC)-(field4_test$LT50ACC))^2) / sum((field4_test$LT50ACC)^2)
# 0.0232
sum(((field4_predictionACC_norm)-(field4_test$LT50ACC))^2) / sum((field4_test$LT50ACC)^2)
# 0.0286
```


### Barplots of predicted LT50ACC for test data
```{r barplots of predicted LT50ACC for test data}
plot(field4_test$LT50ACC, field4_predictionACC)
plot(field4_test$LT50ACC, field4_predictionACC_factor)

barplot(t(as.matrix(field2_predictionACC_table[,c(2,4)])), 
        beside=T, names.arg = field2_predictionACC_table$accession, las=2)

barplot(t(as.matrix(field3_predictionACC_table[,c(2,4)])), 
        beside=T, names.arg = field3_predictionACC_table$accession, las=2)

barplot(t(as.matrix(field4_predictionACC_table[,2:3])), 
        beside=T, names.arg = field4_predictionACC_table$accession, las=2)

barplot(t(as.matrix(field4_predictionACC_table[,c(2,4)])), 
        beside=T, names.arg = field4_predictionACC_table$accession, las=2)
```


### Confidence interval
```{r confidence interval}
# mehr zum Thema Vorhersage und Konfidenz-Intervalle im Kapitel 3 im Buch Regression, Seite 123
# Vorhersage mit 95%-Vorhersageintervall
head(predict(field4_regressionACC, field4_test, interval="prediction"))
# Vorhersage mit 95%-Konfidenzintervall
head(predict(field4_regressionACC, field4_test, interval="confidence"))
```


# Accessions of unknown LT50
```{r accessions of unknown LT50}
accessions_unknown

field3_regressionACC_all_factor$xlevels$`bbch` <- union(field3_regressionACC_all_factor$xlevels$`bbch`,
                                                        levels(field3_unknown$bbch))

field4_regressionACC_all_factor$xlevels$`bbch` <- union(field4_regressionACC_all_factor$xlevels$`bbch`,
                                                        levels(field4_unknown$bbch))

field2_unknown_predictionACC_factor <- predict(field2_regressionACC_all_factor, field2_unknown_fac[,6:15])
field3_unknown_predictionACC_factor <- predict(field3_regressionACC_all_factor, field3_unknown_fac[,6:15])
field4_unknown_predictionACC_factor <- predict(field4_regressionACC_all_factor, field4_unknown_fac[,6:15])
names(field2_unknown_predictionACC_factor) <- field2_unknown$accession
names(field3_unknown_predictionACC_factor) <- field3_unknown$accession
names(field4_unknown_predictionACC_factor) <- field4_unknown$accession

field2_unknown_predictionACC <- predict(field2_regressionACC, field2_unknown[,6:15])
field3_unknown_predictionACC <- predict(field3_regressionACC, field3_unknown[,6:15])
field4_unknown_predictionACC <- predict(field4_regressionACC, field4_unknown[,6:15])
names(field2_unknown_predictionACC) <- field2_unknown$accession
names(field3_unknown_predictionACC) <- field3_unknown$accession
names(field4_unknown_predictionACC) <- field4_unknown$accession
```


# ANOVA per model/trial
```{r ANOVA per model}
summary(field2_regressionACC_factor)
summary(field3_regressionACC_factor)
summary(field4_regressionACC_factor)

anova(field2_regressionACC_factor)
anova(field3_regressionACC_factor)
anova(field4_regressionACC_factor)

# scaled and transformed data
summary(field2T_regressionACC_factor)
summary(field3T_regressionACC_factor)
summary(field4T_regressionACC_factor)

# training data
anova(field2T_regressionACC_factor)
anova(field3T_regressionACC_factor)
anova(field4T_regressionACC_factor)

# all data
anova(field2T_regressionACC_all_factor)
anova(field3T_regressionACC_all_factor)
anova(field4T_regressionACC_all_factor)
```

## Combine ANOVA p-values of all trials
```{r Combine ANOVA of all trials}
field2T_regressionACC_all_factor_anova <- 
  # anova(field2T_regressionACC_factor)[1:10,] %>% 
  anova(field2T_regressionACC_all_factor)[1:10,] %>% 
  rownames_to_column("predictor") %>%
  mutate(trial = "Field 2007/08")

field3T_regressionACC_all_factor_anova <- 
  # anova(field3T_regressionACC_factor)[1:10,] %>% 
  anova(field3T_regressionACC_all_factor)[1:10,] %>% 
  rownames_to_column("predictor") %>%
  mutate(trial = "Field 2008/09")

field4T_regressionACC_all_factor_anova <- 
  # anova(field4T_regressionACC_factor)[1:10,] %>% 
  anova(field4T_regressionACC_all_factor)[1:10,] %>% 
  rownames_to_column("predictor") %>%
  mutate(trial = "Field 2009/10")

# fieldT_regressionACC_factor_anova <-
#   bind_rows(field2T_regressionACC_factor_anova,
#             field3T_regressionACC_factor_anova,
#             field4T_regressionACC_factor_anova)

fieldT_regressionACC_all_factor_anova <-
  bind_rows(field2T_regressionACC_all_factor_anova,
            field3T_regressionACC_all_factor_anova,
            field4T_regressionACC_all_factor_anova)

colnames(fieldT_regressionACC_factor_anova)[6] <- "pvalue"
colnames(fieldT_regressionACC_all_factor_anova)[6] <- "pvalue"

fieldT_regressionACC_factor_anova %>%
  ggplot(.) + geom_tile(aes(x = trial, y = predictor, fill = -log10(pvalue)))

fieldT_regressionACC_all_factor_anova %>%
  ggplot(.) + geom_tile(aes(x = trial, y = predictor, fill = -log10(pvalue)))

plot(-log10(fieldT_regressionACC_factor_anova$pvalue),-log10(fieldT_regressionACC_all_factor_anova$pvalue) )

# for table in manuscript?
# fieldT_regressionACC_factor_anova %>%
#   dplyr::select(predictor, trial, pvalue) %>%
#   pivot_wider(names_from = trial, values_from = pvalue, id_cols = "predictor") %>% View
```


# Reduced model
## Cp-plot
```{r Cp-plot}
################### Cp-Plot ####################
x <- as.matrix(field4_train[,c(6:15)]) # Create design matrix x
y <- field4_train$LT50ACC # Create response vector y
leapsModels <- leaps(x, y, method="Cp", nbest=3, names=colnames(x))

pdf(file="figures/Metabolite_AraField4_CpPlot.pdf")
Cpplot(leapsModels)
dev.off()

adjr <- leaps(x, y, method="adjr2") # Check which model the adjusted R2 selects
maxadjr(adjr, 8)

# y <- arafield4_regressiondata$yLT50ACC[cooks.distance(arafield4_regressionACC)<0.2]
# x <- cbind(arafield4_regressiondata[,4:14])
# x_part <- x[cooks.distance(arafield4_regressionACC)<0.2,]
# a <- leaps(x_part, y)
# b <- leaps(x, y, method="adjr2")
```


# LASSO model
## LASSO model field 3
```{r LASSO model field 3}
x <- as.matrix(field3_train[,6:15])
y <- field3_train$LT50ACC

test_x <- as.matrix(field3_test[,6:15])
test_y <- field3_test$LT50ACC

set.seed(1)
field3_lasso_ACC <- glmnet(x,y)

# Cross-validation
set.seed(1)
field3_lasso_cv_ACC <- cv.glmnet(x, y)

# compare different alpha values
set.seed(1)
field3_lasso_cv_ACC1 <- cv.glmnet(x, y, alpha=1)
set.seed(1)
field3_lasso_cv_ACC.5 <- cv.glmnet(x, y, alpha=0.5)
set.seed(1)
field3_lasso_cv_ACC0 <- cv.glmnet(x, y, alpha=0)
plot(field3_lasso_cv_ACC1)
plot(field3_lasso_cv_ACC.5)
plot(field3_lasso_cv_ACC0)
plot(log(field3_lasso_cv_ACC1$lambda),field3_lasso_cv_ACC1$cvm,pch=19,
     col="red",xlab="log(Lambda)",ylab=field3_lasso_cv_ACC1$name, 
     xlim = c(-6,6), ylim = c(2.5,3.1))
points(log(field3_lasso_cv_ACC.5$lambda),field3_lasso_cv_ACC.5$cvm,pch=19,col="grey")
points(log(field3_lasso_cv_ACC0$lambda),field3_lasso_cv_ACC0$cvm,pch=19,col="blue")
legend("topleft",legend=c("alpha= 1","alpha= .5","alpha 0"),pch=19,col=c("red","grey","blue"))

# use no specific lambda
field3_lasso_ACC_pred_test <- predict(field3_lasso_ACC, new = test_x)

# calculate mean error
field3_lasso_ACC_pred_test_err <- apply( (field3_lasso_ACC_pred_test - test_y)^2, 2, mean )

# use optimal lamba value
field3_lasso_ACC_pred_all_lambda <- predict(field3_lasso_cv_ACC, new = x, s="lambda.min")
field3_lasso_ACC_pred_test_lambda <- predict(field3_lasso_cv_ACC, new = test_x, s="lambda.min")

# number of variables in model (for lambda.min)
which(field3_lasso_cv_ACC$lambda == field3_lasso_cv_ACC$lambda.min)
# 38
field3_lasso_cv_ACC$nzero[38]
# 10

# coefficients
field3_lasso_cv_ACC$glmnet.fit$beta[,"s38"]

cor.test(y, field3_lasso_ACC_pred_all_lambda)
# 0.534
cor.test(test_y, field3_lasso_ACC_pred_test_lambda)
# 0.560

field3_lasso_ACC_pred_test_lambda_table <- data.frame(accession = field3_test$accession,
                                                      measured = field3_test$LT50ACC,
                                                      lasso = field3_lasso_ACC_pred_test_lambda, 
                                                      regression_fac = field3_predictionACC_factor)

sum(((field3_lasso_ACC_pred_test_lambda)-(field3_test$LT50ACC))^2)
# 47.8

sum(((field3_lasso_ACC_pred_test_lambda)-(field3_test$LT50ACC))^2) / sum((field3_test$LT50ACC)^2)
# 0.0211
```


## LASSO model field 4
```{r LASSO model field 4}
x <- as.matrix(field4_train[,6:15])
y <- field4_train$LT50ACC

test_x <- as.matrix(field4_test[,6:15])
test_y <- field4_test$LT50ACC

set.seed(1)
field4_lasso_ACC <- glmnet(x,y)

# Cross-validation
set.seed(1)
field4_lasso_cv_ACC <- cv.glmnet(x, y)
# plot(field4_lasso_cv_ACC)

# compare different alpha values
# compare different alpha values
field4_lasso_cv_ACC1 <- cv.glmnet(x, y, alpha=1)
field4_lasso_cv_ACC.95 <- cv.glmnet(x, y, alpha=0.95)
field4_lasso_cv_ACC.5 <- cv.glmnet(x, y, alpha=0.5)
field4_lasso_cv_ACC0 <- cv.glmnet(x, y, alpha=0)
plot(field4_lasso_cv_ACC1)
plot(field4_lasso_cv_ACC.95)
plot(field4_lasso_cv_ACC.5)
plot(field4_lasso_cv_ACC0)
plot(log(field4_lasso_cv_ACC1$lambda),field4_lasso_cv_ACC1$cvm,pch=19,
     col="red",xlab="log(Lambda)",ylab=field4_lasso_cv_ACC1$name, 
     xlim = c(-6,6), ylim = c(2.1,3.4))
points(log(field4_lasso_cv_ACC.95$lambda),field4_lasso_cv_ACC.95$cvm,pch=19,col="green")
points(log(field4_lasso_cv_ACC.5$lambda),field4_lasso_cv_ACC.5$cvm,pch=19,col="grey")
points(log(field4_lasso_cv_ACC0$lambda),field4_lasso_cv_ACC0$cvm,pch=19,col="blue")
legend("topleft",legend=c("alpha= 1","alpha= .95","alpha= .5","alpha 0"),pch=19,col=c("red","green","grey","blue"))

# use no specific lambda
field4_lasso_ACC_pred_test <- predict(field4_lasso_ACC, new = test_x)

# calculate mean error
field4_lasso_ACC_pred_test_err <- apply( (field4_lasso_ACC_pred_test - test_y)^2, 2, mean )
# points(log(field4_lasso_ACC$lambda), field4_lasso_ACC_pred_test_err, col="blue", pch="*")

# use optimal lamba value
field4_lasso_ACC_pred_all_lambda <- predict(field4_lasso_cv_ACC, new = x, s="lambda.min")
field4_lasso_ACC_pred_test_lambda <- predict(field4_lasso_cv_ACC, new = test_x, s="lambda.min")

# number of variables in model (for lambda.min)
which(field4_lasso_cv_ACC$lambda == field4_lasso_cv_ACC$lambda.min)
# 23
field4_lasso_cv_ACC$nzero[23]
# 4

# coefficients
field4_lasso_cv_ACC$glmnet.fit$beta[,"s22"]


# plot observed vs. predicted DRYM
# plot(y, lasso_all_pred_all_lambda)
# abline(0,1)
cor.test(y, field4_lasso_ACC_pred_all_lambda) # 0.609
cor.test(test_y, field4_lasso_ACC_pred_test_lambda) # 0.592

field4_lasso_ACC_pred_test_lambda_table <- data.frame(accession = field4_test$accession,
                                                      measured = field4_test$LT50ACC,
                                                      lasso = field4_lasso_ACC_pred_test_lambda, 
                                                      regression = field4_predictionACC,
                                                      regression_fac = field4_predictionACC_factor)


# barplot(t(as.matrix(lasso_all_pred_test_lambda_table[,2:3])), 
#         beside=T, names.arg = lasso_all_pred_test_lambda_table$acc, las=2)
# 
# barplot(t(as.matrix(lasso_all_pred_test_lambda_table[,2:4])), 
#         beside=T, names.arg = lasso_all_pred_test_lambda_table$acc, las=2)

sum(((field4_lasso_ACC_pred_test_lambda)-(field4_test$LT50ACC))^2)
# 38.3

sum(((field4_lasso_ACC_pred_test_lambda)-(field4_test$LT50ACC))^2) / sum((field4_test$LT50ACC)^2)
# 0.0169
```


## Prediction of LT50ACC with model (for accessions that are included in the model = training data)
```{r prediction of LT50ACC using model with ORIGINAL BBCH and LASSO}
cor.test(field4_regressionACC$fitted.values, field4_train$LT50ACC)
# R = 0.636
cor.test(field4_lasso_ACC_pred_all_lambda, field4_train$LT50ACC)
# R = 0.609

par(mfrow=c(1,2))

plot(field4_train$LT50ACC, field4_regressionACC$fitted.values, main = "regression",
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), pch = 19)
abline(lm(field4_regressionACC$fitted.values ~ field4_train$LT50ACC))
#lm(field4_regressionACC$fitted.values ~ field4_train$LT50ACC)

plot(field4_train$LT50ACC, field4_lasso_ACC_pred_all_lambda, main = "lasso",
     ylab="LT50ACC fitted values", 
     xlab="LT50ACC measured values", 
     xlim=c(-12.5,-5), ylim=c(-12.5,-5), pch = 19)
abline(lm(field4_lasso_ACC_pred_all_lambda ~ field4_train$LT50ACC))
#lm(field4_lasso_ACC_pred_all_lambda ~ field4_train$LT50ACC)
```



# Variable selection

**step is a slightly simplified version of stepAIC in package MASS (Venables & Ripley, 2002 and earlier editions).**

## Forward selection
```{r forward selection}
# Create initial NULL model
field2_regressionACC_null <- lm(field2_train$LT50ACC ~ 1, data = field2_train)
field3_regressionACC_null <- lm(field3_train$LT50ACC ~ 1, data = field3_train)
field4_regressionACC_null <- lm(field4_train$LT50ACC ~ 1, data = field4_train)

# field2T_regressionACC_null <- lm(field2_trainT$LT50ACC ~ 1, data = field2_trainT)
# field3T_regressionACC_null <- lm(field3_trainT$LT50ACC ~ 1, data = field3_trainT)
# field4T_regressionACC_null <- lm(field4_trainT$LT50ACC ~ 1, data = field4_trainT)

field2T_regressionACC_fac_null <- lm(field2_trainT_fac$LT50ACC ~ 1, data = field2_trainT_fac)
field3T_regressionACC_fac_null <- lm(field3_trainT_fac$LT50ACC ~ 1, data = field3_trainT_fac)
field4T_regressionACC_fac_null <- lm(field4_trainT_fac$LT50ACC ~ 1, data = field4_trainT_fac)

# Choose a model by AIC in a Stepwise Algorithm
# Forward
field4_regressionACC_fw <- step(field4_regressionACC_null, 
                                scope = formula(field4_regressionACC), 
                                direction = "forward")

# Step:  AIC=92.9
# yLT50ACC ~ norm_bbch + raf + fum + suc
#              Df Sum of Sq RSS  AIC
# <none>                    233 92.9
# + glc         1     2.340 231 93.8
# + pro         1     2.281 231 93.8
# + frc         1     2.266 231 93.8
# + anthoscore  1     1.144 232 94.4
# + antho       1     0.510 233 94.7
# + aa          1     0.101 233 94.9
# + mal         1     0.006 233 94.9

summary(field4_regressionACC_fw)
# R = 0.3912

# Vorhersage-Werte fuer Oekotypen, die nicht im Modell sind
predict(field4_regressionACC_fw, field4_test)

# Summe der quadratischen Differenzen zwischen Mess- und Vorhersagewert 
sum(((predict(field4_regressionACC_fw, field4_test)) - (field4_test$LT50ACC))^2)
# 51.4
sum(((predict(field4_regressionACC_fw, field4_test)) - (field4_test$LT50ACC))^2)/ sum((field4_test$LT50ACC)^2)
# 0.0227
```


## stepAIC
### default direction (if missing): backward!!!
```{r selection in both directions}
############  SELECTION in BOTH DIRECTIONS ###################
# Without FACTOR!
step(field4_regressionACC_null, 
     scope = formula(field4_regressionACC_factor), 
     direction = "both")
# Step:  AIC=92.9
# yLT50ACC ~ norm_bbch + raf + fum + suc
# Call:
# lm(formula = yLT50ACC ~ norm_bbch + raf + fum + suc, data = arafield4_regressiondata)
# Coefficients:
# (Intercept)    norm_bbch          raf          fum          suc  
#     -7.2765       1.0280      -0.9556      -0.0720       0.0542  

# das selbe Modell wie bei forward selection

# Both directions, starting from Null model (FACTOR!)
field2_regressionACC_factorAIC <- step(field2_regressionACC_null, 
                                       scope = formula(field2_regressionACC_factor), 
                                       direction = "both")
# raf, fum, mal

field3_regressionACC_factorAIC <- step(field3_regressionACC_null, 
                                       scope = formula(field3_regressionACC_factor), 
                                       direction = "both")
summary(field3_regressionACC_factorAIC)
# factor(bbch), pro, raf, glc

field4_regressionACC_factorAIC <- step(field4_regressionACC_null, 
                                       scope = formula(field4_regressionACC_factor), 
                                       direction = "both")
summary(field4_regressionACC_factorAIC)
# factor(bbch), raf, suc, fum


# Summe der quadratischen Differenzen zwischen Mess- und Vorhersagewert
sum(((predict(field4_regressionACC_factorAIC, field4_test)) - (field4_test$LT50ACC))^2) / sum((field4_test$LT50ACC)^2)
# 0.0227
sum(((predict(field4_regressionACC_factorAIC, field4_test)) - (field4_test$LT50ACC))^2)
# 51.6

# bind predicted and measured LT50ACC for accessions of test set
head(data.frame(accession = field3_test$accession, 
                measured = field3_test$LT50ACC,
                predicted_reduced = predict(field3_regressionACC_factorAIC, field3_test),
                predicted_full = predict(field3_regressionACC_factor, field3_test)))

head(data.frame(accession = field4_test$accession, 
                measured = field4_test$LT50ACC,
                predicted_reduced = predict(field4_regressionACC_factorAIC, field4_test),
                predicted_full = predict(field4_regressionACC_factor, field4_test)))
```

### transformed and scaled data: default direction (if missing): backward!!!
```{r transformed and scaled data: selection in both directions}
############  SELECTION in BOTH DIRECTIONS ###################
# Both directions, starting from Null model (WITH FACTOR!)
field2T_regressionACC_factorAIC <- step(field2T_regressionACC_fac_null, 
                                       scope = formula(field2T_regressionACC_factor), 
                                       direction = "both")
# raf, fum, mal (AIC=72.58)

field3T_regressionACC_factorAIC <- step(field3T_regressionACC_fac_null, 
                                       scope = formula(field3T_regressionACC_factor), 
                                       direction = "both")
summary(field3T_regressionACC_factorAIC)
# factor(bbch), aa, suc (AIC=99.49)

# raf, glc, pro, fum (AIC=99.08)

field4T_regressionACC_factorAIC <- step(field4T_regressionACC_fac_null, 
                                       scope = formula(field4T_regressionACC_factor), 
                                       direction = "both")
summary(field4T_regressionACC_factorAIC)
# factor(bbch), raf, suc (AIC=94.17)

# factor(bbch), raf, suc, fum (AIC=83.17)


# Summe der quadratischen Differenzen zwischen Mess- und Vorhersagewert
sum(((predict(field4T_regressionACC_factorAIC, field4_testT)) - (field4_test$LT50ACC))^2) / sum((field4_test$LT50ACC)^2)
# 0.0209
sum(((predict(field4T_regressionACC_factorAIC, field4_testT)) - (field4_test$LT50ACC))^2)
# 47.6

# bind predicted and measured LT50ACC for accessions of test set
head(data.frame(accession = field3_test$accession, 
                measured = field3_test$LT50ACC,
                predicted_reduced = predict(field3T_regressionACC_factorAIC, field3_testT),
                predicted_full = predict(field3T_regressionACC_factor, field3_testT)))

head(data.frame(accession = field4_test$accession, 
                measured = field4_test$LT50ACC,
                predicted_reduced = predict(field4T_regressionACC_factorAIC, field4_testT),
                predicted_full = predict(field4T_regressionACC_factor, field4_testT)))
```


### stepAIC default
#### default direction (if missing): backward!!!
```{r stepAIC default}
field4_stepAIC_ACC <- stepAIC(field4_regressionACC)
field4_stepAIC_ACC$coefficients

field2T_stepAIC_ACC_factor <- stepAIC(field2T_regressionACC_factor) # mal fum raf (AIC=72.58)
field3T_stepAIC_ACC_factor <- stepAIC(field3T_regressionACC_factor) # fum pro glc raf (AIC=99.08)
field4T_stepAIC_ACC_factor <- stepAIC(field4T_regressionACC_factor) # suc raf factor(bbch) (AIC=94.17)

field2T_stepAIC_ACC_factor$call
# lm(formula = LT50ACC ~ raf + mal + fum, data = field2_trainT)
field3T_stepAIC_ACC_factor$call
# lm(formula = LT50ACC ~ glc + raf + fum + pro, data = field3_trainT)
field4T_stepAIC_ACC_factor$call
# lm(formula = LT50ACC ~ suc + raf + factor(bbch), data = field4_trainT)

# predited values for accessions that are included in the null model (training data)
# field4_stepAIC_ACC$fitted.values

# predited values for accessions that are NOT included in the null model (test data)
field4_prediction_stepAIC_ACC <- predict(field4_stepAIC_ACC, field4_test)

field2T_prediction_stepAIC_ACC_factor <- predict(field2T_stepAIC_ACC_factor, field2_testT)
field3T_prediction_stepAIC_ACC_factor <- predict(field3T_stepAIC_ACC_factor, field3_testT)
field4T_prediction_stepAIC_ACC_factor <- predict(field4T_stepAIC_ACC_factor, field4_testT)
 
head(data.frame(accession = field4_test$accession, 
                measured = field4_test$LT50ACC, 
                predicted_reduced = field4_prediction_stepAIC_ACC,
                predicted_full = predict(field4_regressionACC, field4_test)))

# Summe der quadratischen Differenzen zwischen Mess- und Vorhersagewert
sum(((field4_prediction_stepAIC_ACC) - (field4_test$LT50ACC))^2) # 51.4
sum(((field4_prediction_stepAIC_ACC) - (field4_test$LT50ACC))^2)/ sum((field4_test$LT50ACC)^2) # 0.0227

sum(((field4T_prediction_stepAIC_ACC_factor) - (field4_test$LT50ACC))^2) # 47.6
sum(((field4T_prediction_stepAIC_ACC_factor) - (field4_test$LT50ACC))^2)/ sum((field4_test$LT50ACC)^2) # 0.0209
sum(((predict(field4T_regressionACC_factor, field4_testT)) - (field4_test$LT50ACC))^2) # 54.1 (full model)
```


### comparison of two models --> ANOVA
```{r comparison of two models}
# anova(arafield4_regressionACC, arafield4_stepAIC_ACC)
# Analysis of Variance Table
# 
# Model 1: yLT50ACC ~ glc + frc + suc + raf + aa + mal + fum + pro + antho + 
#     norm_bbch + anthoscore
# Model 2: yLT50ACC ~ frc + raf + fum + pro + norm_bbch
#   Res.Df RSS Df Sum of Sq    F Pr(>F)
# 1     97 223                         
# 2    103 230 -6     -6.28 0.45   0.84

# p-value > 0.05, also Annahme von H0, dass beide Modelle aehnlich
```


### backward selection
```{r backward selection}
step(object = field2T_regressionACC_factor, scope=formula(field2T_regressionACC_factor), direction="backward")
field2T_regressionACC_factor_red <- step(object = field2T_regressionACC_factor, 
                                         scope = formula(field2T_regressionACC_factor), 
                                         direction = "both")
# AIC=72.58, mal, raf, fum


step(object = field3T_regressionACC_factor, scope=formula(field3T_regressionACC_factor), direction="backward")
# AIC=99.08, glc, raf, fum, pro --> minimal AIC (but small difference)
field3T_regressionACC_factor_red <- step(object = field3T_regressionACC_null, 
                                         scope = formula(field3T_regressionACC_factor), 
                                         direction = "both")
# AIC=99.49, suc, aa, factor(bbch)
field3T_prediction_stepAIC_ACC_factor2 <- predict(field3T_regressionACC_factorAIC, field3_testT)

stepAIC(field4T_regressionACC_factor, direction = "both") # suc raf factor(bbch) (AIC=94.17)



```

## field2
```{r variable selection field2}
stepAIC(object = field2T_regressionACC_factor, scope = formula(field2T_regressionACC_factor), direction = "backward") # 72.58
stepAIC(object = field2T_regressionACC_factor, scope = formula(field2T_regressionACC_factor), direction = "both") # 72.58
stepAIC(object = field2T_regressionACC_fac_null, scope = formula(field2T_regressionACC_factor), direction = "both") # 72.58
stepAIC(object = field2T_regressionACC_fac_null, scope = formula(field2T_regressionACC_factor), direction = "forward") # 72.58

step(object = field2T_regressionACC_factor, scope = formula(field2T_regressionACC_factor), direction = "backward") # 72.58
step(object = field2T_regressionACC_factor, scope = formula(field2T_regressionACC_factor), direction = "both") # 72.58
step(object = field2T_regressionACC_fac_null, scope = formula(field2T_regressionACC_factor), direction = "both") # 72.58
step(object = field2T_regressionACC_fac_null, scope = formula(field2T_regressionACC_factor), direction = "forward") # 72.58


field2T_regressionACC_factor_red <- step(object = field2T_regressionACC_factor, 
                                         scope = formula(field2T_regressionACC_factor), 
                                         direction = "both")
# AIC=72.58, mal, raf, fum
```

## field3
```{r variable selection field3}
stepAIC(object = field3T_regressionACC_factor, scope = formula(field3T_regressionACC_factor), direction = "backward") # 99.08
stepAIC(object = field3T_regressionACC_factor, scope = formula(field3T_regressionACC_factor), direction = "both") # 99.08
stepAIC(object = field3T_regressionACC_fac_null, scope = formula(field3T_regressionACC_factor), direction = "both") 
# AIC=99.49, bbch + aa + suc
stepAIC(object = field3T_regressionACC_fac_null, scope = formula(field3T_regressionACC_factor), direction = "forward") # 100.85

step(object = field3T_regressionACC_factor, scope = formula(field3T_regressionACC_factor), direction = "backward") # 99.08
step(object = field3T_regressionACC_factor, scope = formula(field3T_regressionACC_factor), direction = "both") # 99.08
step(object = field3T_regressionACC_fac_null, scope = formula(field3T_regressionACC_factor), direction = "both") # 99.49
step(object = field3T_regressionACC_fac_null, scope = formula(field3T_regressionACC_factor), direction = "forward") # 100.85

field3T_regressionACC_factor_red <- step(object = field3T_regressionACC_factor, 
                                         scope = formula(field3T_regressionACC_factor), 
                                         direction = "both")
# AIC=99.08, glc + raf + fum + pro
```

## field4
```{r variable selection field4}
stepAIC(object = field4T_regressionACC_factor, scope = formula(field4T_regressionACC_factor), direction = "backward") # 94.17
stepAIC(object = field4T_regressionACC_factor, scope = formula(field4T_regressionACC_factor), direction = "both") # 94.17
stepAIC(object = field4T_regressionACC_fac_null, scope = formula(field4T_regressionACC_factor), direction = "both") # 94.17
stepAIC(object = field4T_regressionACC_fac_null, scope = formula(field4T_regressionACC_factor), direction = "forward") # 94.17

step(object = field4T_regressionACC_factor, scope = formula(field4T_regressionACC_factor), direction = "backward") # 94.17
step(object = field4T_regressionACC_factor, scope = formula(field4T_regressionACC_factor), direction = "both") # 94.17
step(object = field4T_regressionACC_fac_null, scope = formula(field4T_regressionACC_factor), direction = "both") # 94.17
step(object = field4T_regressionACC_fac_null, scope = formula(field4T_regressionACC_factor), direction = "forward") # 94.17

field4T_regressionACC_factor_red <- step(object = field4T_regressionACC_factor, 
                                         scope = formula(field4T_regressionACC_factor), 
                                         direction = "both")
# AIC=94.17, suc, raf, factor(bbch)
```


## Combine predictions across trials REDUCED
```{r combine training predictions across trials REDUCED}
pred_df2_train_red <- data.frame(data = "trial2",
                                 accession = field2_trainT_fac$accession,
                                 measured = field2_trainT_fac$LT50ACC,
                                 # predicted = field2T_regressionACC_factorAIC$fitted.values)
                                 predicted = field2T_regressionACC_factor_red$fitted.values)

pred_df3_train_red <- data.frame(data = "trial3",
                                 accession = field3_trainT_fac$accession,
                                 measured = field3_trainT_fac$LT50ACC,
                                 # predicted = field3T_regressionACC_factorAIC$fitted.values)
                                 predicted = field3T_regressionACC_factor_red$fitted.values)

pred_df4_train_red <- data.frame(data = "trial4",
                                 accession = field4_trainT_fac$accession,
                                 measured = field4_trainT_fac$LT50ACC,
                                 # predicted = field4T_regressionACC_factorAIC$fitted.values)
                                 predicted = field4T_regressionACC_factor_red$fitted.values)

pred_df_train_red <- rbind(pred_df2_train_red, 
                           pred_df3_train_red, 
                           pred_df4_train_red)
```


## Correlation scatter plot
```{r correlation training scatter plot REDUCED}
range(pred_df_train_red$predicted)
range(pred_df_train_red$measured)
lim_values <- c(-12.5, -5)

sp_red <- ggscatter(pred_df_train_red, x = "measured", y = "predicted",
                add = "reg.line", 
                add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                conf.int = TRUE)

sp_red_lim <- ggpar(sp_red, ylim = lim_values, xlim = lim_values)
sp_red_facet <- facet(sp_red_lim + theme_bw(), facet.by = "data")
sp_red_final <- sp_red_facet + stat_cor(method = "pearson", label.x = -11, label.y.npc = 'top')

ggsave(filename = "figures/predictions_train_reduced_corr_plot.png", 
       plot = sp_red_final, width = 10, height = 4, dpi = 600)
```


### add quadratic term for norm_bbch
#### increases R2 from 0.34 to 0.37
```{r add quadratic term for norm_bbch}
summary(lm(LT50ACC ~ frc + raf + fum + poly(bbch,2), data = field4_train_norm))
# Adjusted R-squared:  0.345 

testlm2 <- lm(I(1/LT50ACC) ~ frc + raf + fum + poly(bbch,2), data = field4_train_norm)
summary(lm(I(1/LT50ACC) ~ frc + raf + fum + poly(bbch,2), data = field4_train_norm))
# Adjusted R-squared: 0.39
```


### regression model for LT50NAC
```{r regression model for LT50NAC}
############## Regressionsmodell fuer LT50NAC #######################
field4_regressionNAC <- lm(LT50NAC ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field4_train)

summary(field4_regressionNAC)
# Adjusted R-squared:  0.121

# Vorhersage der LT50-Werte mit Hilfe des Modells (fuer Oekotypen, die im Modell enthalten sind)
# field4_regressionNAC$fitted.values

cor(field4_train$LT50NAC, field4_regressionNAC$fitted.values)
# 0.45
```


### prediction of LT50NAC for test data
```{r prediction of LT50NAC for test data}
field4_predictionNAC <- predict(field4_regressionNAC, field4_test)

head(data.frame(accession = field4_test$accession,
                measured = field4_test$LT50NAC,
                predicted = field4_predictionNAC))
```


### stepAIC for LT50NAC
```{r stepAIC for LT50NAC}
field4_stepAIC_NAC <- stepAIC(field4_regressionNAC)

summary(field4_stepAIC_NAC)
# Adjusted R-squared:  0.164 

# fuer Oekotypen, die im Modell sind
field4_stepAIC_NAC$fitted.values

# fuer Oekotypen, die nicht im Modell sind
field4_prediction_stepAIC_NAC <- predict(field4_stepAIC_NAC, field4_test)

head(data.frame(accession = field4_test$accession,
                measured = field4_test$LT50NAC,
                predicted = field4_prediction_stepAIC_NAC))
```


### regression model for acclimation capacity
```{r regression model for acclimation capacity}
# yCapacity4

field4_regressionCap <- lm(cap ~ glc + frc + suc + raf + aa + mal + fum + pro + 
                                bbch + anthoscore, data = field4_train_norm)

summary(field4_regressionCap)
# Adjusted R-squared:  0.278 

# Vorhersage der Kapazitaet mit Hilfe des Modells (fuer Oekotypen, die im Modell enthalten sind)
# field4_regressionCap$fitted.values

cor(field4_train_norm$cap, field4_regressionCap$fitted.values)
# 0.587
```


### prediction of capacity for test data
```{r prediction of capacity for test data}
# Vorhersage der LT50-Werte fuer ausgewaehlte Akzessionen mit Hilfe des berechneten Regressionsmodells
field4_predictionCap <- predict(field4_regressionCap, field4_test_norm)

head(data.frame(accession = field4_test$accession,
                measured = field4_test$cap,
                predicted = field4_predictionCap))
```


### stepAIC for Capacity
```{r stepAIC for Capacity}
field4_stepAIC_Cap <- stepAIC(field4_regressionCap)

summary(field4_stepAIC_Cap)
# Adjusted R-squared:  0.305

# fuer Oekotypen, die im Modell sind
# field4_stepAIC_Cap$fitted.values

# fuer Oekotypen, die nicht im Modell sind
field4_prediction_stepAIC_Cap <- predict(field4_stepAIC_Cap, field4_test_norm)

head(data.frame(accession = field4_test$accession,
                measured = field4_test$cap,
                predicted = field4_prediction_stepAIC_Cap))
```


# Model comparison regarding test set
```{r Model comparison regarding test set}
pred_df_test2 <- data.frame(trial = "trial2",
                            accession = field2_test$accession, 
                            measured = field2_test$LT50ACC,
                            predicted_full = predict(field2T_regressionACC_factor, field2_testT_fac),
                            # predicted_reduced = predict(field2T_regressionACC_factorAIC, field2_testT_fac))
                            predicted_reduced = predict(field2T_regressionACC_factor_red, field2_testT_fac))

pred_df_test3 <- data.frame(trial = "trial3",
                            accession = field3_test$accession, 
                            measured = field3_test$LT50ACC,
                            predicted_full = predict(field3T_regressionACC_factor, field3_testT_fac),
                            # predicted_reduced = predict(field3T_regressionACC_factorAIC, field3_testT_fac))
                            predicted_reduced = predict(field3T_regressionACC_factor_red, field3_testT_fac))

pred_df_test4 <- data.frame(trial = "trial4",
                            accession = field4_test$accession, 
                            measured = field4_test$LT50ACC,
                            predicted_full = predict(field4T_regressionACC_factor, field4_testT_fac),
                            # predicted_reduced = predict(field4T_regressionACC_factorAIC, field4_testT_fac))
                            predicted_reduced = predict(field4T_regressionACC_factor_red, field4_testT_fac))

pred_df_test <- rbind(pred_df_test2, pred_df_test3, pred_df_test4)

pred_df_test_ <- 
  pred_df_test %>%
  gather(key = "model", value = "predicted", 4:5)

pred_df_test_2 <- 
  pred_df_test %>%
  gather(key = "model", value = "value", 3:5) %>%
  droplevels()
```

## Correlation scatter plot
```{r correlation test scatter plot}
range(pred_df_test_$predicted)
range(pred_df_test_$measured)
lim_values <- c(-12.5, -5)

sp_test <- ggscatter(pred_df_test_, x = "measured", y = "predicted",
                add = "reg.line", 
                add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                conf.int = TRUE)

sp_test_lim <- ggpar(sp_test, ylim = lim_values, xlim = lim_values)
sp_test_facet <- facet(sp_test_lim + theme_bw(), facet.by = c("model","trial"))
sp_test_final <- sp_test_facet + stat_cor(method = "pearson", label.x = -12, label.y.npc = 'top')

ggsave(filename = "figures/predictions_test_corr_plot.png", 
       plot = sp_test_final, width = 7, height = 5, dpi = 600)
```

## Barplot of predicted LT50ACC
```{r barplot of predicted LT50ACC}
lt50_test <- lt50 %>% 
  filter(Accession %in% accessions_testing) %>% 
  droplevels() %>%
  arrange(-LT50ACC) # decreasing order by LT50ACC

pred_df_test_2$accession <- factor(pred_df_test_2$accession, 
                                   levels = as.character(lt50_test$Accession))

levels(pred_df_test_2$trial) <- c("Field 2007/08", "Field 2008/09", "Field 2009/10")

p_bar_test <- 
  ggplot(pred_df_test_2) + 
  geom_bar(aes(x = accession, y = value, fill = model),
           position = "dodge", stat = "summary",  fun.y = "mean") +
  scale_fill_manual(name = "", values = pal_grey,
                    breaks = c("measured", "predicted_full", "predicted_reduced"),
                    labels = c("measured", "predicted full model", "predicted reduced model")) +
  theme_bw() + ylab(expression("LT"[50]*"ACC (C")) +
  facet_wrap(~ trial) + 
  # theme_pubr() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "bottom")

ggsave(filename = "figures/predictions_test_barplot.png", 
       plot = p_bar_test, width = 8, height = 5, dpi = 600)
```

# LOO-CV
## Full model
```{r loo-cv field, warning=FALSE}
# full model
field2_error_cv <- func_calc_loocv(field2)
field3_error_cv <- func_calc_loocv(field3)
field4_error_cv <- func_calc_loocv(field4)

field2_pred_cv <- func_calc_loocv(field2, return_val = "pred", data_name = "field2")
field3_pred_cv <- func_calc_loocv(field3, return_val = "pred", data_name = "field3")
field4_pred_cv <- func_calc_loocv(field4, return_val = "pred", data_name = "field4")

pred_loocv_df <- rbind(field2_pred_cv, field3_pred_cv, field4_pred_cv)

range(pred_loocv_df$predicted)
range(pred_loocv_df$measured)
lim_values <- c(-12.5, -5)

sp_loocv <- ggscatter(pred_loocv_df, x = "measured", y = "predicted",
                add = "reg.line", 
                add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                conf.int = TRUE)

sp_loocv_lim <- ggpar(sp_loocv, ylim = lim_values, xlim = lim_values)
sp_loocv_facet <- facet(sp_loocv_lim + theme_bw(), facet.by = "data")
sp_loocv_final <- sp_loocv_facet + stat_cor(method = "pearson", label.x.npc = 'center', label.y.npc = 'top')

ggsave(filename = "figures/predictions_loocv_corr_plot.png", 
       plot = sp_loocv_final, width = 10, height = 4, dpi = 600)
```

## Reduced model
### field2
```{r LOO-CV reduced model field2}
field2_sel_vars <- variable.names(field2T_stepAIC_ACC_factor)[-1]
field2_sel_varsT <- c(model_colnamesT_factor[1:5], "raffinoseT", "malateT", "fumarateT")
field2_sel_vars_short <- c(model_colnames_short[1:5], field2_sel_vars)
field2_formula <- as.formula(paste("LT50ACC", paste(field2_sel_vars, collapse = " + "), sep = " ~ "))
print(field2_formula) # LT50ACC ~ raf + mal + fum

field2_red_pred_cv <- func_calc_loocv(field_data = field2, return_val = "pred", 
                                      data_name = "field2", 
                                      colnames_short = field2_sel_vars_short, 
                                      colnamesT_factor = field2_sel_varsT, 
                                      formula = field2_formula)
```

### field3
```{r LOO-CV reduced model field3}
# field3T_regressionACC_factor_red
# glc + raf + fum + pro
field3_sel_vars <- c("glc", "raf", "fum", "pro")
field3_sel_varsT <- c(model_colnamesT_factor[1:5], "glucoseT", "raffinoseT", "fumarateT", "prolineT")

# alternative:
# field3T_regressionACC_factorAIC
# factor(bbch) + aa + suc
# field3_sel_vars <- c("bbch", "aa", "suc")
# field3_sel_varsT <- c(model_colnamesT_factor[1:5], "bbch_factor", "amino_acidsT", "sucroseT")
field3_sel_vars_short <- c(model_colnames_short[1:5], field3_sel_vars)

# field3_formula <- as.formula(paste("LT50ACC ~ bbch + aa + suc"))
field3_formula <- as.formula(paste("LT50ACC ~ glc + raf + fum + pro"))
print(field3_formula) # LT50ACC ~ glc + raf + fum + pro

field3_red_pred_cv <- func_calc_loocv(field_data = field3, return_val = "pred", 
                                      data_name = "field3", 
                                      colnames_short = field3_sel_vars_short, 
                                      colnamesT_factor = field3_sel_varsT, 
                                      formula = field3_formula)
```

### field4
```{r LOO-CV reduced model field4}
# field4T_regressionACC_factor_red
field4_sel_vars <- c("bbch", "raf", "suc")
field4_sel_varsT <- c(model_colnamesT_factor[1:5], "bbch_factor", "raffinoseT", "sucroseT")
field4_sel_vars_short <- c(model_colnames_short[1:5], field4_sel_vars)
field4_formula <- as.formula(paste("LT50ACC", paste(field4_sel_vars, collapse = " + "), sep = " ~ "))
print(field4_formula) # LT50ACC ~ bbch + raf + suc

field4_red_pred_cv <- func_calc_loocv(field_data = field4, return_val = "pred", 
                                      data_name = "field4", 
                                      colnames_short = field4_sel_vars_short, 
                                      colnamesT_factor = field4_sel_varsT, 
                                      formula = field4_formula)
```

### Mean per accession (reduced model)
```{r Mean per accession (reduced model)}
pred_red_loocv_df <- rbind(field2_red_pred_cv, field3_red_pred_cv, field4_red_pred_cv)

pred_red_loocv_mean_df <- 
  pred_red_loocv_df %>%
  group_by(data,accession) %>%
  summarise(predicted_mean = mean(predicted),
            predicted_se = sd(predicted)/sqrt(length(predicted)),
            predicted_n = length(predicted)) %>%
  left_join(lt50, by = c("accession" = "Accession")) %>%
  dplyr::select(-LT50NAC, -SE_LT50NAC, -Model, -Capacity)

colnames(pred_red_loocv_mean_df)[3] <- "predicted"
colnames(pred_red_loocv_mean_df)[6] <- "measured"

pred_red_loocv_mean_df$abs_diff <- abs(pred_red_loocv_mean_df$predicted - pred_red_loocv_mean_df$measured)

pred_red_loocv_df$model <- "prediction (reduced model)"
pred_loocv_df$model <- "prediction (full model)"
```

### Combine full and reduced model
```{r Mean combine full and reduced model}
pred_all_loocv_df <- rbind(pred_loocv_df, pred_red_loocv_df)

pred_all_loocv_mean_df <- 
  pred_all_loocv_df %>%
  group_by(data,model,accession) %>%
  summarise(predicted_mean = mean(predicted),
            predicted_se = sd(predicted)/sqrt(length(predicted)),
            predicted_n = length(predicted)) %>%
  left_join(lt50, by = c("accession" = "Accession")) %>%
  dplyr::select(-LT50NAC, -SE_LT50NAC, -Model, -Capacity)

colnames(pred_all_loocv_mean_df)[4] <- "predicted"
colnames(pred_all_loocv_mean_df)[7] <- "measured"

pred_all_loocv_mean_df$abs_diff <- abs(pred_all_loocv_mean_df$predicted - pred_all_loocv_mean_df$measured)

range(pred_all_loocv_mean_df$predicted)
range(pred_all_loocv_mean_df$measured)
lim_values <- c(-12.5, -5)

sp_all_loocv_mean <- ggscatter(pred_all_loocv_mean_df, x = "measured", y = "predicted",
                           add = "reg.line", 
                           add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                           conf.int = TRUE)

sp_all_loocv_mean_lim <- ggpar(sp_all_loocv_mean, ylim = lim_values, xlim = lim_values)
sp_all_loocv_mean_facet <- facet(sp_all_loocv_mean_lim + theme_bw(), facet.by = c("model","data"))
sp_all_loocv_mean_final <- sp_all_loocv_mean_facet + stat_cor(method = "pearson", label.x = -11, label.y.npc = 'top')

ggsave(filename = "figures/predictions_loocv_corr_plot_mean_full_red.png", 
       plot = sp_all_loocv_mean_final, width = 10, height = 7, dpi = 600)
```


## Mean per accession for LT50 (pred vs measured)
```{r loo-cv field mean per accession, warning=FALSE}
pred_loocv_mean_df <- 
  pred_loocv_df %>%
  group_by(data,accession) %>%
  summarise(predicted_mean = mean(predicted),
            predicted_se = sd(predicted)/sqrt(length(predicted)),
            predicted_n = length(predicted)) %>%
  left_join(lt50, by = c("accession" = "Accession")) %>%
  dplyr::select(-LT50NAC, -SE_LT50NAC, -Model, -Capacity)

colnames(pred_loocv_mean_df)[3] <- "predicted"
colnames(pred_loocv_mean_df)[6] <- "measured"

pred_loocv_mean_df$abs_diff <- abs(pred_loocv_mean_df$predicted - pred_loocv_mean_df$measured)

range(pred_loocv_mean_df$predicted)
range(pred_loocv_mean_df$measured)
lim_values <- c(-12.5, -5)

sp_loocv_mean <- ggscatter(pred_loocv_mean_df, x = "measured", y = "predicted",
                           add = "reg.line", 
                           add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                           conf.int = TRUE)

sp_loocv_mean_lim <- ggpar(sp_loocv_mean, ylim = lim_values, xlim = lim_values)
sp_loocv_mean_facet <- facet(sp_loocv_mean_lim + theme_bw(), facet.by = "data")
sp_loocv_mean_final <- sp_loocv_mean_facet + stat_cor(method = "pearson", label.x = -11, label.y.npc = 'top')

ggsave(filename = "figures/predictions_loocv_corr_plot_mean.png", 
       plot = sp_loocv_mean_final, width = 10, height = 4, dpi = 600)
```

## Plot of absolute difference between pred and measured
```{r Plot of absolute difference between pred and measured}
plot_abs_diff <- 
  pred_loocv_mean_df %>%
  ggplot(., aes(x = measured, y = abs_diff)) + 
  geom_point() + facet_wrap(~data) +
  geom_smooth()

ggsave(filename = "figures/predictions_loocv_abs_diff.png", 
       plot = plot_abs_diff, width = 8, height = 5, dpi = 600)

# reduced + full model
pred_all_loocv_mean_df %>%
  ggplot(., aes(x = measured, y = abs_diff, color = model, group = model)) + 
  geom_point() + facet_wrap(~data) +
  geom_smooth()
```

## Barplot of mean per accession (full model)
```{r Barplot of mean per accession (full model)}
pred_loocv_mean_longer <-
  pred_loocv_mean_df %>%
  dplyr::select(-predicted_n, -abs_diff) %>%
  dplyr::rename("mean_predicted" = predicted) %>%
  dplyr::rename("se_predicted" = predicted_se) %>%
  dplyr::rename("mean_measured" = measured) %>%
  dplyr::rename("se_measured" = SE_LT50ACC) %>%
  pivot_longer(cols = -c(data, accession),
               names_to = c(".value", "set"), 
               names_pattern = "(.*)_(.*)") %>%
  left_join(lt50[,c(1,3)], by = c("accession" = "Accession"))

barplot_pred_loocv_mean <- 
  pred_loocv_mean_longer %>%
  ggplot(., aes(x = reorder(accession, LT50ACC), y = mean, fill = set)) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se),
                  width = 0.2,                    # Width of the error bars
                  position=position_dodge(0.9)) +
  facet_rep_wrap(~data, ncol=1, repeat.tick.labels = 'all') +
  scale_fill_manual(name  = "", values = pal_grey[c(1,3)]) +
  theme_bw() + xlab("accession") + 
  ylab(expression("LT"[50]*"ACC (C")) +
  theme(axis.text.x  = element_text(angle=90, hjust=1), legend.position = "bottom")

ggsave(filename = "figures/predictions_loocv_barplot_mean.png", 
       plot = barplot_pred_loocv_mean, width = 10, height = 12, dpi = 600)
```

## Barplot of mean per accession (reduced and full model)
```{r Barplot of mean per accession (reduced and full model)}
pred_all_loocv_mean_longer <-
  pred_red_loocv_mean_df %>%
  dplyr::select(-predicted_n, -abs_diff) %>%
  dplyr::rename("mean_predictedR" = predicted) %>%
  dplyr::rename("se_predictedR" = predicted_se) %>%
  dplyr::rename("mean_measured" = measured) %>%
  dplyr::rename("se_measured" = SE_LT50ACC) %>%
  bind_cols(pred_loocv_mean_df[,3:4]) %>%
  dplyr::rename("mean_predictedF" = predicted) %>%
  dplyr::rename("se_predictedF" = predicted_se) %>%
  pivot_longer(cols = -c(data, accession),
               names_to = c(".value", "set"), 
               names_pattern = "(.*)_(.*)") %>%
  left_join(lt50[,c(1,3)], by = c("accession" = "Accession"))

levels(pred_all_loocv_mean_longer$data) <- c("Field 2007/08", "Field 2008/09", "Field 2009/10")

barplot_all_pred_loocv_mean <- 
  pred_all_loocv_mean_longer %>%
  ggplot(., aes(x = reorder(accession, LT50ACC), y = mean, fill = set)) + 
  geom_bar(position = position_dodge(), stat = "identity") +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se),
                  width = 0.2,                    # Width of the error bars
                  position=position_dodge(0.9)) +
  facet_rep_wrap(~data, ncol=1, repeat.tick.labels = 'all') +
  scale_fill_manual(name  = "",
                    values = pal_grey, #brewer.pal(n = 4, "Paired")[2:4],
                    breaks = c("measured", "predictedF", "predictedR"),
                    labels = c("measured", "predicted full model", "predicted reduced model")) +
  theme_bw() + xlab("accession") + 
  ylab(expression("LT"[50]*"ACC (C")) +
  theme(axis.text.x  = element_text(angle=90, hjust=1), legend.position = "bottom")

ggsave(filename = "figures/predictions_loocv_barplot_mean_all.png", 
       plot = barplot_all_pred_loocv_mean, width = 9, height = 10, dpi = 600)
```

# compare different models
## plot function for comparison
```{r plot function for comparison}
func_pred_plot <- function(x, y, main_text, limits = lim_values, labels){
  plot(x, y, col = "white", main = main_text, 
       xlim = limits, ylim = limits)
  
  text(x, y, labels)
  text(-10, -6, paste("cor = ", round(cor(x, y), digits = 3)), col = "red")
  abline(0, 1, col = "red")
  
  print(cor.test(x, y))
}
```

## compare different models field 2
```{r compare different models field 2}
test_y <- field2_test$LT50ACC
range(test_y)
lim_values <- c(-11.5, -5.5)

pdf("figures/field2_compare_model_validation.pdf", 8, 8)
par(mfrow=c(2,2))

func_pred_plot(x = test_y,
               y = field2_lasso_ACC_pred_test_lambda,
               main_text = "lasso model",
               labels = field2_test$accession)

func_pred_plot(x = test_y,
               y = field2T_predictionACC_factor,
               main_text = "full reg model with scaled data",
               labels = field2_testT$accession)

func_pred_plot(x = test_y,
               y = field2T_prediction_stepAIC_ACC_factor,
               main_text = "reduced reg model with scaled data",
               labels = field2_testT$accession)

func_pred_plot(x = test_y,
               y = field2_predictionACC_factor,
               main_text = "full reg model (factor)",
               labels = field2_test$accession)
dev.off()
```


## compare different models field 3
```{r compare different models field 3}
test_y <- field3_test$LT50ACC
range(test_y)
lim_values <- c(-11.5, -5.5)

pdf("figures/field3_compare_model_validation.pdf", 8, 8)
par(mfrow=c(2,2))

func_pred_plot(x = test_y,
               y = field3_lasso_ACC_pred_test_lambda,
               main_text = "lasso model",
               labels = field3_test$accession)

# func_pred_plot(x = test_y,
#                y = field3_predictionACC,
#                main_text = "regression model",
#                labels = field3_test$accession)

func_pred_plot(x = test_y,
               y = field3T_predictionACC_factor,
               main_text = "full reg model with scaled data",
               labels = field3_testT$accession)

func_pred_plot(x = test_y,
               # y = field3T_prediction_stepAIC_ACC_factor,
               y = field3T_prediction_stepAIC_ACC_factor2,
               main_text = "reduced reg model with scaled data",
               labels = field3_testT$accession)

func_pred_plot(x = test_y,
               y = field3_predictionACC_factor,
               main_text = "full reg model (factor)",
               labels = field3_test$accession)

# func_pred_plot(x = test_y,
#                y = field3_predictionACC_norm,
#                main_text = "regression model (norm BBCH)",
#                labels = field3_test$accession)
dev.off()
```

## compare different models field 4
```{r compare different models field 4}
test_y <- field4_test$LT50ACC
range(test_y)
lim_values <- c(-11.5, -5.5)

pdf("figures/field4_compare_model_validation.pdf", 8, 8)
par(mfrow=c(2,2))
func_pred_plot(x = test_y,
               y = field4_lasso_ACC_pred_test_lambda,
               main_text = "lasso model",
               labels = field4_test$accession)

# func_pred_plot(x = test_y,
#                y = field4_predictionACC,
#                main_text = "regression model",
#                labels = field4_test$accession)

func_pred_plot(x = test_y,
               y = field4T_predictionACC_factor,
               main_text = "full reg model with scaled data",
               labels = field4_testT$accession)

func_pred_plot(x = test_y,
               y = field4T_prediction_stepAIC_ACC_factor,
               main_text = "reduced reg model with scaled data",
               labels = field4_testT$accession)


func_pred_plot(x = test_y,
               y = field4_predictionACC_factor,
               main_text = "full reg model (factor)",
               labels = field4_test$accession)

# func_pred_plot(x = test_y,
#                y = field4_predictionACC_norm,
#                main_text = "regression model (norm BBCH)",
#                labels = field4_test$accession)
dev.off()
```


# Caret
## Tuning parameters
```{r caret tuning parameters}
grid <- 10^ seq (10,-2, length = 100) # for lambda

# for glmnet (elastic net)
eGrid <- expand.grid(.alpha = c(0, .05, .1, .3, .5, .7, .9, .95, .99, 1), 
                     .lambda = 10^seq(from = -6,to = 0, length = 50)) # or grid
                     #.lambda = seq(1e-6, 1, length = 50)) # or grid

# for lasso
eGrid_lasso <- expand.grid(.fraction = 10^seq(from = -6,to = 0, length = 50)) # or grid


Control <- trainControl(#method = "LOOCV",
                        method = "repeatedcv", 
                        repeats = 5, number = 10,
                        verboseIter = TRUE)
```

### plot
```{r plot}
field4_pred_loocv <- func_calc_loocv_caret(field4, method_val = "lasso", eGrid_val = eGrid_lasso)
field4_known_LT50 <- field4$LT50ACC[which(field4$accession %in% accessions_known)]
plot(field4_known_LT50, field4_pred_loocv)

par(mfrow=c(1,1))
func_pred_plot(x = field4_known_LT50,
               y = field4_pred_loocv,
               main_text = "lasso model",
               labels = names(field4_pred_loocv))


```


## try out training with first fold of training data: field 3
```{r try out training with first fold of training data field 3}
method_val <- "glmnet"
# method_val <- "lasso"

# x_f3_train_caret <- field3_train[,c(6:15)]
# y_f3_train_caret <- field3_train$LT50ACC
# x_f3_test <- field3_test[,c(6:15)]
# y_f3_test <- field3_test$LT50ACC

x_f3_train_caret <- field3_trainT[,c(6:15)]
y_f3_train_caret <- field3_trainT$LT50ACC
x_f3_test <- field3_testT[,c(6:15)]
y_f3_test <- field3_testT$LT50ACC

set.seed(1234)
lassoFit_f3T <- train(x = x_f3_train_caret, 
                   y = y_f3_train_caret,
                   method = "lasso",
                   metric = "RMSE",
                   tuneGrid = eGrid_lasso,
                   trControl = Control)

set.seed(1234)
glmnetFit_f3T <- train(x = x_f3_train_caret, 
                   y = y_f3_train_caret,
                   method = "glmnet",
                   metric = "RMSE",
                   tuneGrid = eGrid,
                   trControl = Control)

lassoFit_f3$bestTune # alpha: 1
# 0.754
# R2: 0.2149733

glmnetFit_f3$bestTune
# original training data
# alpha: 0, lambda: 0.3237 --> ridge regression
# R2: 0.1481273, R: 0.385, RMSE: 1.61

# glmnet
table(coef(netFit_f3$finalModel, s = netFit_f3$bestTune$lambda)[,1] != 0)
# 4874 predictors (+ intercept) have coefficients that are not zero, the rest was removed

# table(coef(netFit_f3$finalModel, s = netFit_f3$bestTune$fraction)[,1] != 0)

plot(lassoFit_f3, metric = "Rsquared")
plot(lassoFit_f3T, metric = "Rsquared")
plot(lassoFit_f3, metric = "RMSE")
plot(glmnetFit_f3, metric = "Rsquared")
plot(glmnetFit_f3T, metric = "Rsquared")

plot(y_f3_test, predict(netFit_f3, newdata = x_f3_test))
cor.test(y_f3_test, predict(lassoFit_f3, newdata = x_f3_test)) #  0.558
cor.test(y_f3_test, predict(glmnetFit_f3, newdata = x_f3_test)) # 0.569
cor.test(y_f3_train_caret, predict(lassoFit_f3, newdata = x_f3_train_caret))  # 0.53
cor.test(y_f3_train_caret, predict(glmnetFit_f3, newdata = x_f3_train_caret)) # 0.521

plot(field3_train$LT50ACC, predict(netFit_f3, newdata = x_f3_train_caret), 
     ylim=range(field3_train$LT50ACC), col="white")
text(field3_train$LT50ACC, predict(netFit_f3, newdata = x_f3_train_caret),
     labels = as.character(field3_train$accession), cex = 0.8)
abline(0, 1, col = "grey")
abline(lm(predict(netFit_f3, newdata = x_f3_train_caret) ~ field3_train$LT50ACC), col = "red")
```

## try out training with first fold of training data: field 4
```{r try out training with first fold of training data field 4}
# x_f4_train_caret <- field4_train[,c(6:15)]
# y_f4_train_caret <- field4_train$LT50ACC
# x_f4_test <- field4_test[,c(6:15)]
# y_f4_test <- field4_test$LT50ACC

x_f4_train_caret <- field4_trainT[,c(6:15)]
y_f4_train_caret <- field4_trainT$LT50ACC
x_f4_test <- field4_testT[,c(6:15)]
y_f4_test <- field4_testT$LT50ACC

set.seed(1234)
lassoFit_f4T <- train(x = x_f4_train_caret, 
                     y = y_f4_train_caret,
                     method = "lasso",
                     metric = "RMSE",
                     tuneGrid = eGrid_lasso,
                     trControl = Control)

set.seed(1234)
glmnetFit_f4T <- train(x = x_f4_train_caret, 
                      y = y_f4_train_caret,
                      method = "glmnet",
                      metric = "RMSE",
                      tuneGrid = eGrid,
                      trControl = Control)

lassoFit_f4$bestTune
# 0.569
# R2: 0.3687, R: 0.606

glmnetFit_f4$bestTune
# alpha: 0.95, lambda: 0.07906043 --> nearly ridge regression
# R2: 0.3678, R: 0.606, RMSE: 1.493873

# glmnet
table(coef(netFit_f4$finalModel, s = netFit_f4$bestTune$lambda)[,1] != 0)
# 4874 predictors (+ intercept) have coefficients that are not zero, the rest was removed

# table(coef(netFit_f4$finalModel, s = netFit_f4$bestTune$fraction)[,1] != 0)

plot(lassoFit_f4, metric = "Rsquared")
plot(lassoFit_f4T, metric = "Rsquared")
plot(glmnetFit_f4, metric = "Rsquared")
plot(glmnetFit_f4T, metric = "Rsquared")

# plot(y_f4_test, predict(glmnetFit_f4, newdata = x_f4_test))
cor(y_f4_test, predict(lassoFit_f4T, newdata = x_f4_test)) #  0.537
cor(y_f4_test, predict(glmnetFit_f4T, newdata = x_f4_test)) # 0.536
cor(y_f4_train_caret, predict(lassoFit_f4T, newdata = x_f4_train_caret))  # 0.623
cor(y_f4_train_caret, predict(glmnetFit_f4T, newdata = x_f4_train_caret)) # 0.623

plot(field4_train$LT50ACC, predict(netFit_f4, newdata = x_f4_train_caret), 
     ylim=range(field4_train$LT50ACC), col="white")
text(field4_train$LT50ACC, predict(netFit_f4, newdata = x_f4_train_caret),
     labels = as.character(field4_train$accession), cex = 0.8)
abline(0, 1, col = "grey")
abline(lm(predict(netFit_f4, newdata = x_f4_train_caret) ~ field4_train$LT50ACC), col = "red")
```

```{r}
par(mfrow=c(2,2))
func_pred_plot(x = test_y,
               y = field3_lasso_ACC_pred_test_lambda,
               main_text = "lasso model field 3",
               labels = field3_test$accession)

func_pred_plot(x = test_y,
               y = predict(netFit_f3, newdata = x_f3_test),
               main_text = "caret model field 3",
               labels = field3_test$accession)

func_pred_plot(x = test_y,
               y = field4_lasso_ACC_pred_test_lambda,
               main_text = "lasso model field 4",
               labels = field4_test$accession)

func_pred_plot(x = test_y,
               y = predict(netFit_f4, newdata = x_f4_test),
               main_text = "caret model field 4",
               labels = field4_test$accession)
```



## caret with nested CV
```{r caret with nested CV}
method_val <- "glmnet"
# method_val <- "lasso"

# predicted and corresponding true values
dim(folds)
# yhats_caret = matrix(nrow = nrow(folds), ncol = ncol(folds)) # fold in column
yhats_caret <- list()
rsquared_caret <- c()

fit_caret <- list()
fitcv_caret <- list()

# iterate over the outer folds 
for(i in 1:ncol(folds)){
  
  x_f4_train_caret <- x_f4_train[[i]]
  x_f4_test_caret <- x_f4_test[[i]]
  y_f4_train_caret <- y_f4_train[[i]]
  y_f4_test_caret <- y_f4_test[[i]]
  # colnames(x_train_caret) <- seq(1:ncol(x_train_caret))
  # colnames(x_test_caret) <- seq(1:ncol(x_test_caret))
  
  fit_caret[[i]] <- train(x = x_f4_train_caret, 
                          y = y_f4_train_caret,
                          method = method_val,
                          tuneGrid = eGrid,
                          # tuneGrid = eGrid_lasso,
                          trControl = Control)

  # predict the test set observations and calculate prediction score
  # yhats_caret[,i] <- predict(fit_caret[[i]], 
  #                    newdata = x_f4_test_caret)
  
  yhats_caret[[i]] <- predict(fit_caret[[i]], 
                              newdata = x_f4_test_caret)
  
  lm_fit <- lm(yhats_caret[[i]] ~ y_f4_test[[i]])
  rsquared_caret[i] <- summary(lm_fit)$r.squared
}

plot(unlist(yhats_caret) ~ unlist(y_f4_test),
     ylab = "predicted", xlab = "observed")

lm_fit_caret <- lm(unlist(yhats_caret) ~ unlist(y_f4_test))
summary(lm_fit_caret)$r.squared


varImp(fit_caret[[1]])

# get coefficients of variables for best model (1st fold of training data)
coef(fit_caret[[1]]$finalModel, s = fit_caret[[1]]$bestTune$lambda) # is nearly equivalent to glmnet model result

table(coef(fit_caret[[1]]$finalModel, s = fit_caret[[1]]$bestTune$lambda)[,1] != 0)
table(coef(fit_caret[[2]]$finalModel, s = fit_caret[[2]]$bestTune$lambda)[,1] != 0)

sel_var <- as.integer(names(which(coef(fit_caret[[1]]$finalModel, s = fit_caret[[1]]$bestTune$lambda)[,1] != 0)))
sel_var_names <- names_keep_int_c[sel_var[-1]]
View(cor_int_all_r[sel_var_names,])
View(cor_int_all_p[sel_var_names,])

hist(log10(cor_int_c$pval_adj[sel_var_names,]))
hist(log10(cor_int_c$pval_adj))

hist(cor_int_c$coeff[sel_var_names,])
hist(cor_int_c$coeff)
```


# Save workspace
```{r save workspace}
save.image("metabolite_arafield_model.RData")
sessionInfo()
```


```{r}
# knit2html("scripts/metabolite_arafield4_models.Rmd", output = "scripts/metabolite_arafield4_models.html")
```

