---
title: "Untitled"
author: "Heike Sprenger"
date: "September 22, 2015"
output: html_document
---

# Load data
```{r load data}
# reference table for sample information
ref_gesamt_mittel <- read.table("data/gcms/ref_gesamt_mittel.txt", header=T, row.names=1, sep="\t")
head(ref_gesamt_mittel)
dim(ref_gesamt_mittel)
# 150 accessions, 12 columns

# metabolites ref table
metabolites <- read.table("data/gcms/metabolites.txt", na.strings = "NA", sep = "\t", header=T)
dim(metabolites)

# all metabolite data: field3 + field4 --> MEAN values per accession and trial
gesamt_mittel <- read.table("data/gcms/gesamt.txt", row.names = 1, na.strings = "NA", sep = "\t", header=T)
dim(gesamt_mittel)
# 117 metabolites, 150 accessions

# field 3
feld3_mittel <- read.table("data/gcms/feld3.txt", row.names = 1, na.strings = "NA", sep = "\t", header=T)
dim(feld3_mittel)
# 117 75

# field 4
feld4_mittel <- read.table("data/gcms/feld4.txt", row.names = 1, na.strings = "NA", sep = "\t", header=T)
dim(feld4_mittel)
# 117 75

# transpose data sets
gesamt_mittel.t <- t(gesamt_mittel)
feld3_mittel.t <- t(feld3_mittel)
feld4_mittel.t <- t(feld4_mittel)

# accession names
akzessionen <- colnames(gesamt_mittel)
head(akzessionen)

akzessionenfeld <- colnames(feld3_mittel)
akzessionenfeld
```


# Load data without means
For batch test, gesamt 15% NA, 50%Blank and replicates were not averaged ###
```{r load data without means}
gesamt_keinmittel <- read.table("data/gcms/gesamt_keineMittel.txt", 
                                row.names = 1, na.strings = "NA", sep = "\t", header=T) # transformed data
dim(gesamt_keinmittel)
# 117 metabolites, 306 accessions
akzessionen_keinmittel <- colnames(gesamt_keinmittel)
akzessionen_keinmittel

gesamt_keinmittel.t <- t(gesamt_keinmittel)
dim(gesamt_keinmittel.t)
# 306 117

ref_keinmittel <- read.table("data/gcms/ref_gesamtKeineMittel.txt", na.strings = "NA", header = T, sep = "\t")
dim(ref_keinmittel)
# 306 6

# number NAs
sum(is.na(gesamt_mittel))
# 257 
sum(is.na(feld3_mittel))
# 148 
sum(is.na(feld4_mittel))
# 109 
sum(is.na(gesamt_keinmittel))
# 2247
```


# Load data before log-median-ratio-transformation
```{r load data before log-median-ratio-transformation}
norm_keinmittel <- read.table("data/gcms/norm_keineMittelw_15%NA50%Blank.txt", 
                                row.names = 1, na.strings = "NA", sep = "\t", header=T) # UN-transformed data

dim(norm_keinmittel)
# 117 rows (metabolites)
# 306 columns (samples)

heatmap.x  <- as.matrix(norm_keinmittel)

range(heatmap.x, na.rm=T)
# normalized up to 100% per sample (over all metabolites)

hist(apply(heatmap.x, 2, max, na.rm=T), col = "grey", breaks=30, xlab = "maximal value")

hv <- heatmap(heatmap.x, col = heat.colors(256), scale = "column", margins = c(5,10))
hv <- heatmap(heatmap.x, col = heat.colors(256), scale = "row", margins = c(5,10))
```



# Outliers and distibution
```{r Outlier and distribution}
pdf("figures/gcms/boxplots.pdf", width=12, height=6)
boxplot(gesamt_mittel, main = "Boxplots of LOG(10)-median-transformated data",
        sub = "accessions", ylab = "LOG(10)-median-transform", xaxt = "n", col = c(rep("green",75), rep("blue",75)))
axis(at = seq(1:150), labels = akzessionen, side = 1, las = 2, cex.axis = 0.5)

boxplot(feld3_mittel, main = "Boxplots of LOG(10)-median-transformated data for field 3",
        sub = "Akzessionen", ylab = "LOG(10)-median-transform", xaxt = "n")
axis(at = seq(1:75), labels = akzessionenfeld, side = 1, las = 2, cex.axis = 0.5)

boxplot(feld4_mittel, main = "Boxplots of LOG(10)-median-transformated data for field 4",
        ylab = "LOG(10)-median-transform", xaxt = "n")
axis(at = seq(1:75), labels = akzessionenfeld, side = 1, las = 2, cex.axis = 0.5)
dev.off()

pdf("figures/gcms/quantilplots.pdf")
qqnorm(gesamt_mittel.t, main = "Normal quantil-quantil-plot of the data")
qqnorm(feld3_mittel.t, main = "Normal quantil-quantil-plot of field 3")
qqnorm(feld4_mittel.t, main = "Normal quantil-quantil-plot of field 4")
dev.off()

pdf("figures/gcms/histogramms.pdf")
hist(gesamt_mittel.t, main = "Histogramm of the data", ylab = "density", 
     xlab = "", freq = F, prob =T , ylim = c(0,2.5))
lines(density(gesamt_mittel.t, na.rm = T))             
hist(feld3_mittel.t, main = "Histogramm of field 3", ylab = "density", 
     xlab = "", freq = F, prob = T, ylim = c(0,2.5))
lines(density(feld3_mittel.t, na.rm = T))  
hist(feld4_mittel.t, main = "Histogramm of field 4", ylab = "density", 
     xlab = "", freq = F, prob = T, ylim = c(0,2.5))
lines(density(feld4_mittel.t, na.rm = T))    
dev.off()
```


# PCA and ICA
PCA with method svdImpute, see Bachelor thesis
```{r PCA}
pc_mittel <- pca(gesamt_mittel.t, method = "svdImpute", nPcs = 8, center = F)
# svdImpute calculated PCA
# Importance of component(s):
#               PC1    PC2     PC3     PC4     PC5     PC6     PC7     PC8
# R2            0.2542 0.1113 0.09437 0.06942 0.05358 0.04078 0.03322 0.02981
# Cumulative R2 0.2542 0.3654 0.45979 0.52920 0.58278 0.62356 0.65677 0.68659
##117   Variables
# 150 	Samples
# 257 	NAs ( 1.464 %)
# 8 	Calculated component(s)
# Data was NOT mean centered before running PCA 
# Data was NOT scaled before running PCA 
# Scores structure:
#   [1] 150   8
# Loadings structure:
#   [1] 117   8
```

## PCA feld3
```{r PCA feld3}
pcfeld3_mittel <- pca(feld3_mittel.t, method = "svdImpute", nPcs = 8, center = F)
# svdImpute calculated PCA
# Importance of component(s):
#                  PC1    PC2    PC3     PC4     PC5     PC6     PC7     PC8
# R2            0.1904 0.1281 0.1022 0.08714 0.05179 0.04832 0.04616 0.03614
# Cumulative R2 0.1904 0.3185 0.4207 0.50782 0.55961 0.60793 0.65409 0.69024
# 117   Variables
# 75 	Samples
# 148 	NAs ( 1.687 %)
# 8 	Calculated component(s)
# Data was NOT mean centered before running PCA 
# Data was NOT scaled before running PCA 
# Scores structure:
#   [1] 75  8
# Loadings structure:
#   [1] 117   8
```


## PCA feld4
```{r PCA feld4}
pcfeld4_mittel <- pca(feld4_mittel.t, method = "svdImpute", nPcs = 8, center = F)
# svdImpute calculated PCA
# Importance of component(s):
#               PC1    PC2     PC3     PC4     PC5     PC6     PC7     PC8
# R2            0.2368 0.1775 0.09552 0.08155 0.04644 0.04037 0.03315 0.02957
# Cumulative R2 0.2368 0.4143 0.50983 0.59137 0.63781 0.67818 0.71133 0.74090
# 117   Variables
# 75 	Samples
# 109 	NAs ( 1.242 %)
# 8 	Calculated component(s)
# Data was NOT mean centered before running PCA 
# Data was NOT scaled before running PCA 
# Scores structure:
#   [1] 75  8
# Loadings structure:
#   [1] 117   8

```


## PCA keine Mittelwerte
```{r PCA keine Mittelwerte}
pc_keineMittel <- pca(gesamt_keinmittel.t, method = "svdImpute", nPcs = 8, center = F)
# svdImpute calculated PCA
# Importance of component(s):
#                 PC1    PC2     PC3     PC4     PC5     PC6    PC7     PC8
# R2            0.2123 0.1228 0.09217 0.07026 0.06473 0.05577 0.0491 0.03423
# Cumulative R2 0.2123 0.3351 0.42725 0.49750 0.56223 0.61800 0.6671 0.70133
# 117   Variables
# 306 	Samples
# 2247 	NAs ( 6.276 %)
# 8 	Calculated component(s)
# Data was NOT mean centered before running PCA 
# Data was NOT scaled before running PCA 
# Scores structure:
#   [1] 306   8
# Loadings structure:
#   [1] 117   8

pdf("figures/gcms/pca_gesamtKeineMittel.pdf")
x = pc_keineMittel@scores[,1]
y = pc_keineMittel@scores[,2]
plot(x, y, xlab = "PC1", ylab = "PC2", main = "Scores from PC1 and PC2 of the not averaged data", col = "white")
points(x[ref_keinmittel$field=="3"], y[ref_keinmittel$field=="3"], pch=16, col = ref_keinmittel$batch[ref_keinmittel$field=="3"])
points(x[ref_keinmittel$field=="4"], y[ref_keinmittel$field=="4"], pch=17, col = ref_keinmittel$batch[ref_keinmittel$field=="4"])
legend("topright", legend = levels(ref_keinmittel$batch), 
       text.col = seq_along(ref_keinmittel$batch))
legend("bottomright", legend = c("field 3", "field 4"), pch = c(1,2))

plot(pc_keineMittel@scores[,1], pc_keineMittel@scores[,2], col = ref_keinmittel$field, 
     xlab = "PC1", ylab = "PC2", main = "Scores of the whole dataset", pch = 20)
legend("bottomright", legend = c("field 3","field 4"), fill = c("green","blue"))

plot(x, y, col = "white", 
     xlab = "PC1", ylab = "PC2", 
     main = "Scores of the whole dataset", pch = 20)

text(x, y, labels = ref_keinmittel$name_short,
     col = ref_keinmittel$field, cex = 0.7)

legend("bottomleft", legend = c("field 3","field 4"), fill = c("green","blue"))

dev.off()
```


# complete dataset (without means)
```{r complete dataset (without means)}
y_gesamt_keinmittel <- read.table("data/gcms/y_gesamtKeineMittel.txt", sep = "\t", header = T)
dim(y_gesamt_keinmittel)
# 306 accessions 117 metabolites
y_gesamt_keinmittel[1:5,1:5]

dim(gesamt_keinmittel.t)
# 306 accessions 117 metabolites
gesamt_keinmittel.t[1:5,1:5]

# y_gesamt_keinmittel equals gesamt_keinmittel.t, but the latter one has rownames

# reference table
sam_ref_keinmittel <- read.table("data/gcms/sam_ref_gesamtKeineMittel.txt", header = T, sep = "\t")
dim(sam_ref_keinmittel)
# 306 accessions 5 refs (accession feld replicate batch sequence)
class(sam_ref_keinmittel$batch)  #factor?
levels(sam_ref_keinmittel$batch) #shows all available factors

length(levels(sam_ref_keinmittel$accession))

table(sam_ref_keinmittel$accession, sam_ref_keinmittel$field)
```

# ANOVA-Normalization: apply RemoveFactors function
```{r ANOVA-normalization apply RemoveFactors function}
# gesamt_keineMittel.norm <- apply(y_gesamt_keinmittel, 2, RemoveFactors, 
#                                  sam = sam_ref_keinmittel, 
#                                  facs = c("batch","sequence","feld"), 
#                                  keep = c("feld","replicate","accession"))

gesamt_keineMittel.norm <- apply(gesamt_keinmittel.t, 2, RemoveFactors, 
                                 sam = sam_ref_keinmittel, 
                                 facs = c("batch","sequence","field","accession"), 
                                 keep = c("field","accession"))

gesamt_keineMittel.norm[1:5,1:5]
gesamt_keineMittel_anova <- gesamt_keineMittel.norm
gesamt_keineMittel_anova[1:5,1:5]
#rownames(gesamt_keineMittel_anova) <- rownames(gesamt_keinmittel.t) # add accesionnames

write.table(gesamt_keineMittel_anova, file = "output/gcms/gesamt_keineMittel_anovanorm.txt", sep = "\t")
dim(gesamt_keineMittel_anova)
# 306 117
```


## PCA of anova-normalized values
```{r PCA of anova-normalized values}
pc_keineMittel.norm <- pca(gesamt_keineMittel_anova, method = "svdImpute", nPcs = 8, center = F)
# svdImpute calculated PCA
# Importance of component(s):
#   PC1    PC2    PC3     PC4     PC5     PC6     PC7     PC8
# R2            0.1356 0.1270 0.1022 0.07908 0.06001 0.04874 0.04464 0.03956
# Cumulative R2 0.1356 0.2626 0.3649 0.44396 0.50397 0.55271 0.59735 0.63691
# 117   Variables
# 306 	Samples
# 2247 	NAs ( 6.276 %)
# 8 	Calculated component(s)
# Data was NOT mean centered before running PCA 
# Data was NOT scaled before running PCA 
# Scores structure:
#   [1] 306   8
# Loadings structure:
#   [1] 117   8
```


## PCA figure of anova-normalized values
```{r PCA figure of anova-normalized values}
pdf("figures/gcms/pca_gesamtKeineMittel_anova.pdf")
x <- pc_keineMittel.norm@scores[,1]
y <- pc_keineMittel.norm@scores[,2]
plot(x, y, xlab = "PC1", ylab = "PC2", main = "Scores from PC1 and PC2 of the not averaged data, anova normalized", col = "white")
points(x[ref_keinmittel$field=="3"], y[ref_keinmittel$field=="3"], pch = 16, col = ref_keinmittel$batch[ref_keinmittel$field=="3"])
points(x[ref_keinmittel$field=="4"], y[ref_keinmittel$field=="4"], pch = 17, col = ref_keinmittel$batch[ref_keinmittel$field=="4"])
legend("topright", legend = c("11006if","11026if","11039if","11044if","11035if","11012if","11020if"), 
       text.col = seq_along(ref_keinmittel$batch), cex = 0.8)
legend("bottomleft",legend = c("field 3","field 4"), pch = c(1,2))

plot(x, y, col = "white", 
     xlab = "PC1", ylab = "PC2", 
     main = "Scores of the whole dataset", pch = 20)

text(x, y, labels = ref_keinmittel$name_short,
     col = ref_keinmittel$field, cex = 0.7)

legend("bottomleft", legend = c("field 3","field 4"), fill = c("green","blue"))
dev.off()
```


# Mean over Replicate from anova-normalized Data
```{r Mean over Replicate from anova-normalized values}
gesamt_mean_anova <- aggregate(gesamt_keineMittel_anova, 
                               by = list(sam_ref_keinmittel$field, 
                                         sam_ref_keinmittel$accession), 
                               mean)

colnames(gesamt_mean_anova)[c(1,2)] <- c("field", "accession")
gesamt_mean_anova[1:5, 1:5]
dim(gesamt_mean_anova)
# 150 119 

# remove the first two columns
gesamt_mean_anova2 <- gesamt_mean_anova[,3:119]
gesamt_mean_anova2[1:5, 1:5]
dim(gesamt_mean_anova2)
# 150 117

interaction(gesamt_mean_anova$field, gesamt_mean_anova$accession)

#rownames(gesamt_mean_anova2) <- colnames(gesamt_mittel)
rownames(gesamt_mean_anova2) <- interaction(gesamt_mean_anova$field, gesamt_mean_anova$accession)
gesamt_mean_anova2[1:5, 1:5]

pc_gesamt_mean_anova <- pca(gesamt_mean_anova2, method = "svdImpute", nPcs = 8, center = F)

pdf("figures/gcms/pca_gesamtMittel_anova.pdf")
# plot(pc_gesamt_mean_anova@scores[,1],
#      pc_gesamt_mean_anova@scores[,2],
#      col = ref_gesamt_mittel$field, xlab = "PC1", ylab = "PC2", 
#      main = "Scores of the whole dataset, replicates were averaged after anova-norm", pch = 20)
# legend("bottomright", legend = c("field 3","field 4"), fill = c("green","blue"))

color_vector <- rep("green",150)
color_vector[gesamt_mean_anova$field=="field4"] <- "blue"

plot(pc_gesamt_mean_anova@scores[,1],
     pc_gesamt_mean_anova@scores[,2],
     col = "white", xlab = "PC1", ylab = "PC2", pch = 20,
     main = "Scores of the whole dataset, replicates were averaged after anova-norm")

text(pc_gesamt_mean_anova@scores[,1],
     pc_gesamt_mean_anova@scores[,2],
     labels = gesamt_mean_anova$accession,
     col = color_vector, cex = 0.7)

legend("bottomleft", legend = c("field 3","field 4"), fill = c("green","blue"))
dev.off()
```
