---
title: "GC-MS data of Arabidopsis field trial 3 (2008/09)"
author: "Heike Sprenger"
date: "Tuesday, September 29, 2015"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
#getwd()
#setwd("D:/work/repos/ara_field_trials/")
#setwd("~/work/repos/ara_field_trials/")
```

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(pander)
library(ggplot2)
library(pcaMethods)
library(plotrix)
library(moments)
library(fastICA)
library(pls)
library(cwhmisc)
library(gplots)
library(plyr)
library(RColorBrewer)

# set options for knitr
opts_chunk$set(fig.width=10, fig.height=5, tidy = TRUE)
opts_knit$set(root.dir = '../')

# load workspace
#load("gcms_arafield_3.RData")
```


```{r source R scripts, include=FALSE}
source("../functions/colors.R")

source("../functions/RemoveFactors_function.R")
source("../functions/func_normalize.R")
source("../functions/func_log_transform.R")

source("../functions/func_check_norm_dist.R")
source("../functions/find_one_outlier.R")
source("../functions/func_hist_outlier.R")
source("../functions/func_replace_outlier.R")

source("../functions/func_pca_plots.R")
```


# Preprocessing in Excel
* normalized to 13C6-Sorbitol and FW [max-normalized (%)] by Alex
* max-normalized means from 0 to 100% per analyte 
* to make splitless und split30 values comparable, kind of scaling
* mean over replicates
* only "best quantitative information" selected
* siloxane, alkane, all standards and contaminations were removed
* metabolites with more than 15% NAs and/or with more than 50% under blank average were removed
* accession Wassilewskija (Ws-0) was removed because it was missing in one field
* accession Blh-1 was removed, because of the amount of overfitted metabolites
* transformation log(10) ratio median (without Blanks and Multimix)); over all data and each field SEPARATELY!
* 306 individual samples (from 2 years, 75 accessions)
* 117 selected metabolites

# Load samplelist
```{r load samplelist}
samplelist <- read.table("data/gcms/11004if hincha-field-ara-cold-270_profile_combined_samplelist.txt", 
                         row.names = 1, na.strings = "NA", sep = "\t", header=T) 
dim(samplelist)
# 348 samples

table(samplelist$Field)

# select only biological samples from FIELD3
samplelist_biological_samples <- droplevels( subset(samplelist, Select == "yes" & Field == "field3"))
dim(samplelist_biological_samples)
# 150 samples

# look at batches from split and splitless measurement
table(samplelist_biological_samples$Batch_split30, 
      samplelist_biological_samples$Batch_splitless)
```


# Load accession metadata
```{r load accession metadata}
accessions_metadata <- read.table("data/gcms/accessions_metadata.txt",
                                  na.strings = "NA", sep = "\t", header = T)

colnames(accessions_metadata)[3] <- "Accession"

# compare accession names
intersect(accessions_metadata$Accession, levels(samplelist_biological_samples$Accession))
# complete overlap

samplelist_joined <- join(samplelist_biological_samples, accessions_metadata, by = "Accession")
```


# Load metabolite information
```{r load metabolite information}
metabolites_selected <- read.table("data/gcms/metabolites_selected.txt", 
                        na.strings = "NA", sep = "\t", header=T)

dim(metabolites_selected)
# 125 metabolites
```


# Load data including blanks, multimix and biological samples
**Data are not yet transformed, just normalized to sorbitol and FW**
```{r load data including blanks, multimix and biological samples}
norm_biological_samples_selected_all <- read.table("output/gcms/norm_biological_samples_selected.txt", 
                                               row.names = 1, na.strings = "NA", sep = "\t", header=T, check.names = F)
dim(norm_biological_samples_selected_all)
# 125 metabolites, 306 samples

# select samples belonging to FIELD3
norm_biological_samples_selected <- norm_biological_samples_selected_all [, which(colnames(norm_biological_samples_selected_all) %in% samplelist_biological_samples$GMO_ID) ]
dim(norm_biological_samples_selected)
# 125 metabolites, 150 samples
```


## Heatmap of all normalized samples before transformation
```{r heatmap of all normalized samples before transformation}
norm_biological_samples_selected_mat <- as.matrix(norm_biological_samples_selected)

range(norm_biological_samples_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(0,10,length=100),seq(11,30,length=100),seq(31,100,length=100))

pdf("figures/gcms/heatmap_biological_samples_field3.pdf", width = 20, height = 12)
heatmap.2(norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.4, cexCol = 0.4)
dev.off()
```


# PCA before transformation
## PCA of normalized data before transformation: biological samples
```{r PCA of normalized data before transformation: biological samples}
norm_biological_samples_selected_t <- t(norm_biological_samples_selected) 
dim(norm_biological_samples_selected_t)
# 150 samples, 125 metabolites

pca_pareto_center_norm_biological_samples_selected <- pca(norm_biological_samples_selected_t, method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_pareto_center_norm_biological_samples_selected@R2

biplot(pca_pareto_center_norm_biological_samples_selected)
```


## PCA plot of normalized biological samples
```{r PCA plot of normalized biological samples}
pdf("figures/gcms/pca_norm_biological_samples_field3.pdf")

x = pca_pareto_center_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_norm_biological_samples_selected@scores[,2]

palette(heike_palette_7)
plot(x, y, xlab = "PC1", ylab = "PC2", main = "normalized data, pareto scaling, centering", 
     col = samplelist_biological_samples$Batch_split30, pch = 19)
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)

plot(x, y, xlab = "PC1", ylab = "PC2", main = "normalized data, pareto scaling, centering", col = "white", pch = 19)
text(x, y, labels = samplelist_biological_samples$Accession, col = "black", cex = 0.7)

palette(heike_palette_7)
pairs(pca_pareto_center_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch = 19)
dev.off()
```


# Transformation: calculation of log-median-ratio
**Divide normalized value by calculate median, then take log10 of this ratio**
**Median is calculated over ALL biological samples of one metabolite (Field 3 and 4 are considered together)**
```{r transformation: calculation of log-median-ratio}
transform_norm_biological_samples_selected <- func_log_transform(norm_biological_samples_selected)
dim(transform_norm_biological_samples_selected)
# 125 150
```


## Heatmap after transformation
```{r heatmap after transformation}
transform_norm_biological_samples_selected_mat <- as.matrix(transform_norm_biological_samples_selected)

range(transform_norm_biological_samples_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(-1, -0.21, length=100), seq(-0.2, 0.2, length=100), seq(0.21, 2, length=100))

pdf("figures/gcms/heatmap_biological_samples_transformed_field3.pdf", width = 20, height = 12)
heatmap.2(transform_norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.6, cexCol = 0.6)
dev.off()
```


# PCA after transformation
## PCA of transformed data: biological samples
```{r PCA of transformed data: biological samples}
transform_norm_biological_samples_selected_t <- t(transform_norm_biological_samples_selected) 
dim(transform_norm_biological_samples_selected_t)
# 150 samples, 136 metabolites

pca_pareto_center_transform_norm_biological_samples_selected <- pca(transform_norm_biological_samples_selected_t, 
                                                                    method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_pareto_center_transform_norm_biological_samples_selected@R2
biplot(pca_pareto_center_transform_norm_biological_samples_selected)
```


## PCA plot of transformed biological samples
```{r PCA plot of transformed biological samples}
pdf("figures/gcms/pca_transform_norm_biological_samples_field3.pdf")

x = pca_pareto_center_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_transform_norm_biological_samples_selected@scores[,2]

palette(heike_palette_7)
plot(x, y, xlab = "PC1 (18.9%)", ylab = "PC2 (16.1%)", main = "normalized and transformed data, pareto scaling, centering", 
     col = samplelist_biological_samples$Batch_split30, pch = 19)
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)

plot(x, y, xlab = "PC1 (18.9%)", ylab = "PC2 (16.1%)", main = "normalized and transformed data, pareto scaling, centering", col = "white", pch=19)
text(x, y, labels = samplelist_biological_samples$Accession, col = "black", cex = 0.7)

pairs(pca_pareto_center_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch=19, main = "pareto scaling, centering")
legend(0, 1, legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7 , horiz = T)

dev.off()
```


# ANOVA normalization
## generate factors table
```{r generate factors table}
factors <- data.frame("GMO_ID" = samplelist_biological_samples$GMO_ID, 
                      "Accession" = samplelist_biological_samples$Accession,
                      "Batch" = samplelist_biological_samples$Batch_split30,
                      "Sequence" = samplelist_biological_samples$Sequence_split30)
head(factors)
```


## ANOVA normalization: use transformed data (log-median-ratio-transformation)
```{r ANOVA normalization: use transformed data}
# samples must be in rows, analytes in columns
transform_norm_biological_samples_selected_t[1:3, 1:3]

range(transform_norm_biological_samples_selected_t, na.rm=T)
hist(transform_norm_biological_samples_selected_t, breaks=30)

anova_transform_norm_biological_samples_selected_t <- func_normalize(transform_norm_biological_samples_selected_t, 
                                                           factors,
                                                           facs=c("Accession", "Sequence", "Batch"),
                                                           keep=c("Accession"))

dim(anova_transform_norm_biological_samples_selected_t)
# 150 samples, 125 metabolites 
range(anova_transform_norm_biological_samples_selected_t, na.rm=T)
hist(anova_transform_norm_biological_samples_selected_t, breaks=30)

# plot all values before and after ANOVA-normalization
plot(anova_transform_norm_biological_samples_selected_t, transform_norm_biological_samples_selected_t, col = samplelist_biological_samples$Batch_split30)
```


# PCA of ANOVA-normalized data
## PCA of ANOVA-normalized data after transformation: biological samples
```{r PCA of ANOVA-normalized data after transformation: biological samples}
dim(anova_transform_norm_biological_samples_selected_t)
# 150 samples, 125 metabolites

pca_pareto_center_anova_transform_norm_biological_samples_selected <- pca(anova_transform_norm_biological_samples_selected_t, 
                                                          method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_pareto_center_anova_transform_norm_biological_samples_selected@R2

biplot(pca_pareto_center_anova_transform_norm_biological_samples_selected)
```


## PCA plot of ANOVA-normalized data biological samples
```{r PCA plot of ANOVA-normalized data biological samples}
pdf("figures/gcms/pca_anova_transform_norm_biological_samples_field3.pdf")

x = pca_pareto_center_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_anova_transform_norm_biological_samples_selected@scores[,2]
z = pca_pareto_center_anova_transform_norm_biological_samples_selected@scores[,5]

palette(heike_palette_7)
plot(x, y, xlab = "PC1 (17.0%)", ylab = "PC2 (11.9%)", main = "normalized, transformed and ANOVA-normalized data, pareto scaling, centering", 
     col = samplelist_biological_samples$Batch_split30, pch=19)
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)

plot(x, y, xlab = "PC1 (19.5%)", ylab = "PC2 (14.5%)", main = "normalized, transformed and ANOVA-normalized data, pareto scaling, centering", col = "white", pch=19)
text(x, y, labels = samplelist_biological_samples$Accession, col = "black", cex = 0.7)

pairs(pca_pareto_center_anova_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch=19, main = "pareto scaling, centering")
legend(0, 1, legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7 , horiz = T)


# frost tolerance
rbPal <- colorRampPalette(c('blue','red'))
samplelist_joined$LT50_ACC_col <- rbPal(10)[as.numeric(cut(samplelist_joined$LT50_ACC, breaks = 10))]
samplelist_joined$LT50_ACC_col[is.na(samplelist_joined$LT50_ACC_col)] <- "grey"

plot(x, y, xlab = "PC1", ylab = "PC2", 
     main = "normalized, transformed and ANOVA-normalized data, pareto scaling, centering", 
     col = samplelist_joined$LT50_ACC_col, pch=19)
text(x, y, labels = samplelist_joined$Accession, pos = 1,
     col = "black", cex = 0.7)

plot(x, z, xlab = "PC2", ylab = "PC5", 
     main = "normalized, transformed and ANOVA-normalized data, pareto scaling, centering", 
     col = samplelist_joined$LT50_ACC_col, pch=19)
text(x, z, labels = samplelist_joined$Accession, pos = 1,
     col = "black", cex = 0.7)

pairs(pca_pareto_center_anova_transform_norm_biological_samples_selected@scores, 
      col = samplelist_joined$LT50_ACC_col, 
      pch=19, main = "no scaling, no centering")


dev.off()
```


## Heatmap after transformation AND ANOVA-tranformation
```{r heatmap after transformation AND ANOVA-tranformation}
anova_transform_norm_biological_samples_selected_mat <- as.matrix(t(anova_transform_norm_biological_samples_selected_t))

range(anova_transform_norm_biological_samples_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(-1, -0.21, length=100), seq(-0.2, 0.2, length=100), seq(0.21, 2, length=100))

pdf("figures/gcms/heatmap_biological_samples_anova_transformed_field3.pdf", width = 20, height = 12)
heatmap.2(anova_transform_norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.6, cexCol = 0.6)
dev.off()
```


## Barplot of ANOVA-normalized data biological samples
```{r barplot of ANOVA-normalized data biological samples}
pdf("figures/gcms/barplot_norm_field3.pdf", width=20, height=7)
for (i in 1:nrow(norm_biological_samples_selected)){
  barplot(as.numeric(norm_biological_samples_selected[i, 1:150]), 
          col = samplelist_biological_samples$Batch_split30, las=2,
          ylab = "ANOVA-normalized value after log-median transformation",
          names.arg = samplelist_biological_samples$Accession)
 
  mtext(metabolites_selected$Name_Changed[i], side = 3, line = 2, cex = 1.3)
  }
dev.off()

pdf("figures/gcms/barplot_transform_norm_field3.pdf", width=20, height=7)
for (i in 1:nrow(transform_norm_biological_samples_selected)){
  barplot(as.numeric(transform_norm_biological_samples_selected[i, 1:150]), 
          col = samplelist_biological_samples$Batch_split30, las=2,
          ylab = "ANOVA-normalized value after log-median transformation",
          names.arg = samplelist_biological_samples$Accession)
  
  mtext(metabolites_selected$Name_Changed[i], side = 3, line = 2, cex = 1.3)
  }
dev.off()


pdf("figures/gcms/barplot_anova_transform_norm_field3.pdf", width=20, height=7)
for (i in 1:nrow(anova_transform_norm_biological_samples_selected_mat)){
  barplot(as.numeric(anova_transform_norm_biological_samples_selected_mat[i, 1:150]), 
          col = samplelist_biological_samples$Batch_split30, las=2,
          ylab = "ANOVA-normalized value after log-median transformation",
          names.arg = samplelist_biological_samples$Accession)
  
  mtext(metabolites_selected$Name_Changed[i], side = 3, line = 2, cex = 1.3)
  }
dev.off()
```


# Calculate mean over replicates
```{r calculate mean over replicates}
anova_transform_norm_biological_samples_selected_t[1:3, 1:3]
sum(is.na(anova_transform_norm_biological_samples_selected_t))
# 782
# 4.2%

mean_anova_transform_norm_biological_samples_selected <- aggregate(anova_transform_norm_biological_samples_selected_t, 
                                                                   by = list(samplelist_biological_samples$Accession), 
                                                                   mean)

colnames(mean_anova_transform_norm_biological_samples_selected)[1] <- c("Accession")
mean_anova_transform_norm_biological_samples_selected[1:5, 1:5]
dim(mean_anova_transform_norm_biological_samples_selected)
# 75 126
sum(is.na(mean_anova_transform_norm_biological_samples_selected))
# 654
# 7.0%

samplelist_joined_mean <- join(mean_anova_transform_norm_biological_samples_selected[,c(1,2)], accessions_metadata, by = "Accession")
samplelist_joined_mean <- samplelist_joined_mean[,-2]

apply(mean_anova_transform_norm_biological_samples_selected[, -1], 2, function(x) sum(is.na(x)) )

# compare percentage of NAs before and after mean calculation 
# some metabolites have too many NAs after mean calculation!
plot(apply(anova_transform_norm_biological_samples_selected_t, 2, function(x) sum(is.na(x)) )/150*100,
     apply(mean_anova_transform_norm_biological_samples_selected[,-1], 2, function(x) sum(is.na(x)) )/75*100, 
     pch=19, xlim=c(0,50), ylim=c(0,50), xlab="percentage NA BEFORE mean calculation", ylab="percentage NA AFTER mean calculation")
abline(0,1)
grid(10,10)
```


## Heatmap after mean calculation
```{r heatmap after mean calculation}
mean_anova_transform_norm_biological_samples_selected_mat <- as.matrix(t(mean_anova_transform_norm_biological_samples_selected[,-1]))
colnames(mean_anova_transform_norm_biological_samples_selected_mat) <- mean_anova_transform_norm_biological_samples_selected$Accession

# save mean table
write.table(mean_anova_transform_norm_biological_samples_selected_mat, 
            "output/gcms/mean_anova_transform_norm_biological_samples_selected_field3.txt", sep="\t")

range(mean_anova_transform_norm_biological_samples_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(-1, -0.21, length=100), seq(-0.2, 0.2, length=100), seq(0.21, 2, length=100))

pdf("figures/gcms/heatmap_biological_samples_mean_anova_transformed_field3.pdf", width = 20, height = 12)
heatmap.2(mean_anova_transform_norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.6, cexCol = 1)
dev.off()
```


# PCA of mean data
## PCA of mean data (ANOVA-normalized data after transformation)
```{r PCA of mean data}
dim(mean_anova_transform_norm_biological_samples_selected[,-1]) # remove first column!
# 75 samples, 125 metabolites

pca_pareto_center_mean_anova_transform_norm_biological_samples_selected <- pca(mean_anova_transform_norm_biological_samples_selected[,-1], 
                                                          method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@R2

biplot(pca_pareto_center_mean_anova_transform_norm_biological_samples_selected)
```


## PCA scores plot of mean data (ANOVA-normalized data after transformation)
```{r PCA scores plot of mean data}
pdf("figures/gcms/pca_mean_anova_transform_norm_biological_samples_field3.pdf")

x = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,2]

plot(x, y, xlab = "PC1 (19.3%)", ylab = "PC2 (13.8%)", 
     main = "normalized, transformed and ANOVA-normalized MEAN data, pareto scaling, centering", col = "white", pch=19)
text(x, y, labels = mean_anova_transform_norm_biological_samples_selected$Accession, 
     col = "black", cex = 0.7)


dev.off()
```


## PCA scores plot of mean data for frost tolerance, longitude, latitude
```{r PCA scores plot of mean data for frost tolerance longitude latitude, fig.width=6, fig.height=6}
x = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,2]

rbPal <- colorRampPalette(c('blue','red'))
display.brewer.pal(11,"RdYlGn")

# frost tolerance: LT50_ACC
samplelist_joined_mean$LT50_ACC_col <- rbPal(10)[as.numeric(cut(samplelist_joined_mean$LT50_ACC, breaks = 10))]
samplelist_joined_mean$LT50_ACC_col[is.na(samplelist_joined_mean$LT50_ACC_col)] <- "grey"

# longitude
samplelist_joined_mean$Longitude_col <- brewer.pal(11, "RdYlGn")[as.numeric(cut(samplelist_joined_mean$Longitude, breaks = 11))]
samplelist_joined_mean$Longitude_col[is.na(samplelist_joined_mean$Longitude_col)] <- "grey"

# latitude
samplelist_joined_mean$Latitude_col <- brewer.pal(11, "RdYlGn")[as.numeric(cut(samplelist_joined_mean$Latitude, breaks = 11))]
samplelist_joined_mean$Latitude_col[is.na(samplelist_joined_mean$Latitude_col)] <- "grey"

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = samplelist_joined_mean$LT50_ACC_col, pch = 19, main = "LT50_ACC gradient", cex = 1.5)
text(x, y, labels = samplelist_joined_mean$Accession, col = "black", cex = 0.7)

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = samplelist_joined_mean$Longitude_col, pch = 19, main = "Longitude gradient", cex = 1.5)
text(x, y, labels = samplelist_joined_mean$Accession,  col = "black", cex = 0.7)

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = samplelist_joined_mean$Latitude_col, pch = 19, main = "Latitude gradient", cex = 1.5)
text(x, y, labels = samplelist_joined_mean$Accession, col = "black", cex = 0.7)

pairs(pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores, 
      col = samplelist_joined_mean$LT50_ACC_col, 
      pch=19, main = "no scaling, no centering")
```


## PCA loadings plot of mean data (ANOVA-normalized data after transformation)
```{r PCA loadings plot of mean data}
pdf("figures/gcms/pca_loadings_mean_anova_transform_norm_biological_samples_field3.pdf")

x = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@loadings[,1]
y = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@loadings[,2]

palette("default")

plot(x, y, xlab = "PC1", ylab = "PC2", col = "white", pch=19,
     main = "normalized, transformed and ANOVA-normalized MEAN data, pareto scaling, centering")
text(x, y, labels = names(x), col = "black", cex = 0.7)

plot(x, y, xlab = "PC1", ylab = "PC2", col = "white", pch=19,
     main = "normalized, transformed and ANOVA-normalized MEAN data, pareto scaling, centering")
text(x, y, labels = metabolites_selected$Name_Changed, cex = 0.7, col = as.integer(metabolites_selected$Class))

dev.off()

```

# Save workspace
```{r save workspace}
save.image("gcms_arafield_3.RData")
```

