---
title: "GC-MS data of Arabidopsis field trial 3 and 4 (2008/09 and 2009/10)"
author: "Sandra Schaedel"
date: "Tuesday, September 15, 2015"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
#getwd()
#setwd("D:/work/repos/ara_field_trials/")
#setwd("~/work/repos/ara_field_trials/")
```

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(pander)
library(ggplot2)
library(pcaMethods)
library(plotrix)
library(moments)
library(fastICA)
library(pls)
library(cwhmisc)
library(gplots)
library(plyr)
library(RColorBrewer)
library(corrplot)

# set options for knitr
opts_chunk$set(fig.width=10, fig.height=5, tidy = TRUE)
opts_knit$set(root.dir = '../')

# load workspace
#load("gcms_arafields.RData")
```


```{r source R scripts, include=FALSE}
source("../functions/colors.R")

source("../functions/RemoveFactors_function.R")
source("../functions/func_normalize.R")
source("../functions/func_log_transform.R")

source("../functions/func_check_norm_dist.R")
source("../functions/find_one_outlier.R")
source("../functions/func_hist_outlier.R")
source("../functions/func_replace_outlier.R")

source("../functions/func_pca_plots.R")

source("../functions/func_cor_mtest.R")
```


# Preprocessing in Excel
* normalized to 13C6-Sorbitol and FW [max-normalized (%)] by Alex
* max-normalized means from 0 to 100% per analyte 
* to make splitless und split30 values comparable, kind of scaling
* mean over replicates
* only "best quantitative information" selected
* siloxane, alkane, all standards and contaminations were removed
* metabolites with more than 15% NAs and/or with more than 50% under blank average were removed
* accession Wassilewskija (Ws-0) was removed because it was missing in one field
* accession Blh-1 was removed, because of the amount of overfitted metabolites
* transformation log(10) ratio median (without Blanks and Multimix)); over all data and each field SEPARATELY!
* 306 individual samples (from 2 years, 75 accessions)
* 117 selected metabolites

# Load samplelist
```{r load samplelist}
samplelist <- read.table("data/gcms/11004if hincha-field-ara-cold-270_profile_combined_samplelist.txt", 
                         row.names = 1, na.strings = "NA", sep = "\t", header=T) 
dim(samplelist)
# 348 samples

table(samplelist$Field)

# select only biological samples
samplelist_biological_samples <- droplevels( samplelist[samplelist$Select=="yes",] )
dim(samplelist_biological_samples)
# 306 samples

# look at batches from split and splitless measurement
table(samplelist_biological_samples$Batch_split30, 
      samplelist_biological_samples$Batch_splitless)

# samples per field measured in each batch
table(samplelist_biological_samples$Batch_split30, 
      samplelist_biological_samples$Field)
```


# Load accession metadata
```{r load accession metadata}
accessions_metadata <- read.table("data/gcms/accessions_metadata.txt",
                                  na.strings = "NA", sep = "\t", header = T)

colnames(accessions_metadata)[3] <- "Accession"

# compare accession names
intersect(accessions_metadata$Accession, levels(samplelist_biological_samples$Accession))
# complete overlap

samplelist_joined <- join(samplelist_biological_samples, accessions_metadata, by = "Accession")
```


# Load metabolite information
```{r load metabolite information}
metabolites <- read.table("data/gcms/11004if hincha-field-ara-cold-270_profile_combined_metabolites.txt", 
                         row.names = 1, na.strings = "NA", sep = "\t", header=T)
dim(metabolites)
# 166 18

# contains also contaminations, internal reagents and duplicate metabolite (split/splitless)
metabolites_complete <- read.table("data/gcms/11004if hincha-field-ara-cold-270_profile_combined_metabolites_complete.txt", 
                         row.names = 1, na.strings = "NA", sep = "\t", header=T)
dim(metabolites_complete)
# 221 18
```


# Load data including blanks, multimix and biological samples
**Data are not yet transformed, just normalized to sorbitol and FW**
```{r load data including blanks, multimix and biological samples}
norm_part <- read.table("data/gcms/11004if hincha-field-ara-cold-270_profile_combined_part.txt", 
                                row.names = 1, na.strings = "NA", sep = "\t", header=T, check.names = F)
dim(norm_part)
# 166 349

# change colnames to make them unique
head(colnames(norm_part), n=20)
for (i in 2:9) { colnames(norm_part)[i] <- paste("multimix", i-1, sep = "") }
for (i in 10:43) { colnames(norm_part)[i] <- paste("blank", i-9, sep = "") }

norm_biological_samples <- droplevels( norm_part[,44:349])
dim(norm_biological_samples)
# 166 metabolites, 306 samples
```


## corrplot of norm_complete: contains also contaminations, internal reagents and duplicate metabolite (split/splitless)
```{r load norm_complete}

norm_complete <- read.table("data/gcms/11004if hincha-field-ara-cold-270_profile_combined_complete.txt", 
                            row.names = 1, na.strings = "NA", sep = "\t", header=T, check.names = F)
dim(norm_complete)
# 221 metabolites, 349 columns (348 samples)

# calculate NAs and remove analytes with many NAs
ncol_val <- ncol(norm_complete)
percentage_na_complete <- apply(norm_complete, 1, function(x) {sum(is.na(x)) / ncol_val *100 })
index_remove_metabolites_complete <- which(percentage_na_complete > 45) # 50 removed
norm_complete_selected <- norm_complete[-index_remove_metabolites_complete, ]
dim(norm_complete_selected)
# 171 metabolites, 349 columns

metabolites_complete_selected <- metabolites_complete[-index_remove_metabolites_complete,]
dim(metabolites_complete_selected)

norm_complete_selected_t <- t(norm_complete_selected[,-1])
dim(norm_complete_selected_t)
# 348 samples, 171 metabolites (columns)
#colnames(norm_complete_selected_t) <- norm_complete_selected[,1]
colnames(norm_complete_selected_t) <- metabolites_complete_selected$Name

# remove blanks and multimix
norm_complete_t_samples <- norm_complete_selected_t[43:348,]
dim(norm_complete_t_samples)
# 306 biological samples, 171 metabolites
```


```{r corrplot of norm_complete}
norm_complete_cor <- cor(norm_complete_t_samples, method = "pearson", use = "complete")

# corrplot(norm_complete_cor, method="circle", order = "hclust", tl.col = "black", tl.cex = 0.5)

range(norm_complete_cor, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(-1, -0.31, length=100), 
                seq(-0.3, 0.3, length=100), 
                seq(0.31, 1, length=100))

# colors for different metabolite classes
cols_class <- rep("green", 171)
cols_class[metabolites_complete_selected$Class=="Reagent (Contamination)"] <- "blue"
cols_class[metabolites_complete_selected$Class=="Internal Standard (Retention Time)"] <- "red"
cols_class[metabolites_complete_selected$Class=="Internal Standard (Quantification)"] <- "red"

#table(metabolites_complete_selected$Class, cols_class)

#norm_complete_samples_cor <- func_cor_mtest(norm_complete_t_samples)
```

## corrplot heatmap
```{r corrplot heatmap}
pdf("figures/gcms/corr_heatmap.pdf", width = 25, height = 20)
heatmap.2(norm_complete_cor, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.7, cexCol = 0.7,
          ColSideColors = cols_class)
dev.off()
```


## Calculate count and percentage of NAs per metabolite
```{r calculate count and percentage of NAs per metabolite}
ncol_val <- ncol(norm_biological_samples)

count_na <- apply(norm_biological_samples, 1, function(x) sum(is.na(x)) )
percentage_na <- apply(norm_biological_samples, 1, function(x) {sum(is.na(x)) / ncol_val *100 })

length(which(percentage_na > 25))
# 125 metabolites remain, 41 removed
length(which(percentage_na > 45))
# 136 metabolites remain, 30 removed

# remove metabolites with more than 25% NAs
index_remove_metabolites <- which(percentage_na > 25)

norm_biological_samples_selected <- norm_biological_samples[-index_remove_metabolites, ]
dim(norm_biological_samples_selected)
# 125 metabolites, 306 samples
write.table(norm_biological_samples_selected, "output/gcms/norm_biological_samples_selected.txt", sep = "\t")

norm_selected <- norm_part[-index_remove_metabolites, ]
dim(norm_selected)
# 125 metabolites, 349 samples

# table of selected metabolites
metabolites_selected <- droplevels( metabolites[-index_remove_metabolites, ] )
dim(metabolites_selected)
write.table(metabolites_selected, "data/gcms/metabolites_selected.txt", sep = "\t", row.names = F)

# percentage of NAs in selected dataset
sum(is.na(norm_biological_samples_selected)) / (ncol(norm_biological_samples_selected) * nrow(norm_biological_samples_selected)) * 100
# 3.8%
```


## Heatmap of all normalized samples before transformation
```{r heatmap of all normalized samples before transformation}
norm_selected_mat <- as.matrix(norm_selected[,-1])
norm_biological_samples_selected_mat <- as.matrix(norm_biological_samples_selected)

range(norm_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(0,10,length=100),seq(11,30,length=100),seq(31,100,length=100))

pdf("figures/gcms/heatmap_all_samples.pdf", width = 20, height = 12)
heatmap.2(norm_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.4, cexCol = 0.4)
dev.off()

# colors for two different field trials (columns)
cols_field <- rep("green", 306)
cols_field[samplelist_biological_samples$Field=="field4"] <- "blue"

pdf("figures/gcms/heatmap_biological_samples.pdf", width = 20, height = 12)
heatmap.2(norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.4, cexCol = 0.4,
          ColSideColors = cols_field)
dev.off()
```


## Barplot of normalized samples before transformation
```{r barplot of normalized samples before transformation}
pdf("figures/gcms/barplot_norm_before_transformation.pdf", width=20, height=7)

for (i in 1:nrow(norm_selected)){
  barplot(as.numeric(norm_selected[i,2:349]), 
          col = c(rep("orange",8), rep("green", 34), rep("blue",306)),
          ylab = "max-normalized value")
  
  mtext(metabolites_selected$Name_Changed[i], side = 3, line = 2, cex = 1.3)
  mtext(paste("percentage NA:", round(percentage_na[i], digits = 1), "; select:", metabolites$Select[i]),
        side = 3, line = 1, cex = 1)
  
  legend("topright", 
         legend = c("multimix","blank","sample"), 
         fill = c("orange","green","blue"))
}
dev.off()
```


## Boxplot of normalized samples before transformation: look at influence of batch effect (and field)
```{r boxplot of normalized samples before transformation}
pdf("figures/gcms/boxplot_norm_before_transformation.pdf", width=7, height=6)
par(mar=c(8,4,4,2))
for (i in 1:nrow(norm_biological_samples_selected)){
  boxplot(as.numeric(norm_biological_samples_selected[i,]) ~ 
            samplelist_biological_samples$Batch_split30 * 
            samplelist_biological_samples$Field,
          col = heike_palette_7, las = 2,
          main = metabolites_selected$Name_Changed[i])
}
dev.off()
```


## Analyze blanks
```{r analyze blanks}
norm_blanks <- norm_selected[,10:43]

norm_blanks_rowNA <- apply(norm_blanks, 1, mean, na.rm=T)
# if mean is NaN, then no value for this analyte over all blanks
norm_blanks_rowNA_NaN <- is.nan(norm_blanks_rowNA)
table(norm_blanks_rowNA_NaN)

# remove NaNs and then replace NAs by 0 (just for heatmap)
norm_blanks_part <- norm_blanks[!norm_blanks_rowNA_NaN,]
norm_blanks_part[is.na(norm_blanks_part)] <- 0

norm_blanks_mat  <- as.matrix(norm_blanks_part)

pdf("figures/gcms/heatmap_blanks.pdf", width = 20, height = 10)
heatmap.2(norm_blanks_mat, na.rm = T, trace = "none",
          symbreaks = min(norm_blanks_mat, na.rm=TRUE), scale = "row")
dev.off()
```


# PCA before transformation
## PCA of normalized data before transformation: ALL samples
```{r PCA of normalized data before transformation: ALL samples}
norm_selected_t <- t(norm_selected[,-1]) # without column "select"" (if samples are selected)
dim(norm_selected_t)
# 348 samples, 125 metabolites

pca_norm_selected <- pca(norm_selected_t, method = "svdImpute", nPcs = 5, center = F, scale = "none")
pca_pareto_norm_selected <- pca(norm_selected_t, method = "svdImpute", nPcs = 5, center = F, scale = "pareto")
pca_pareto_center_norm_selected <- pca(norm_selected_t, method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

biplot(pca_norm_selected)
biplot(pca_pareto_norm_selected)
biplot(pca_pareto_center_norm_selected)
```


## PCA of normalized data before transformation: biological samples
```{r PCA of normalized data before transformation: biological samples}
norm_biological_samples_selected_t <- t(norm_biological_samples_selected) 
dim(norm_biological_samples_selected_t)
# 306 samples, 125 metabolites

pca_norm_biological_samples_selected <- pca(norm_biological_samples_selected_t, method = "svdImpute", nPcs = 5, center = F, scale = "none")
pca_pareto_norm_biological_samples_selected <- pca(norm_biological_samples_selected_t, method = "svdImpute", nPcs = 5, center = F, scale = "pareto")
pca_pareto_center_norm_biological_samples_selected <- pca(norm_biological_samples_selected_t, method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_norm_biological_samples_selected@R2
pca_pareto_center_norm_biological_samples_selected@R2

biplot(pca_norm_biological_samples_selected)
biplot(pca_pareto_norm_biological_samples_selected)
biplot(pca_pareto_center_norm_biological_samples_selected)
```


## PCA plot of normalized biological samples
```{r PCA plot of normalized biological samples}
pdf("figures/gcms/pca_norm_biological_samples.pdf")
x = pca_norm_biological_samples_selected@scores[,1]
y = pca_norm_biological_samples_selected@scores[,2]

palette(heike_palette_7)
plot(x, y, xlab = "PC1", ylab = "PC2", main = "normalized data, no scaling, no centering", 
     col = samplelist_biological_samples$Batch_split30, pch = c(16,17)[as.numeric(samplelist_biological_samples$Field)])
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)
legend("bottomright", legend = c("field 3", "field 4"), pch = c(16,17), cex = 0.7)

palette(c("green","blue"))
plot(x, y, xlab = "PC1", ylab = "PC2", main = "ormalized data, no scaling, no centering", col = "white", pch = 19)
text(x, y, labels = samplelist_biological_samples$Accession, col = as.numeric(samplelist_biological_samples$Field), cex = 0.7)
legend("topright", legend = c("field 3","field 4"), fill = c("green","blue"))

x = pca_pareto_center_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_norm_biological_samples_selected@scores[,2]

palette(heike_palette_7)
plot(x, y, xlab = "PC1", ylab = "PC2", main = "normalized data, pareto scaling, centering", 
     col = samplelist_biological_samples$Batch_split30, pch = c(16,17)[as.numeric(samplelist_biological_samples$Field)])
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)
legend("bottomright", legend = c("field 3", "field 4"), pch = c(16,17), cex = 0.7)

palette(c("green","blue"))
plot(x, y, xlab = "PC1", ylab = "PC2", main = "normalized data, pareto scaling, centering", col = "white", pch = 19)
text(x, y, labels = samplelist_biological_samples$Accession, col = as.numeric(samplelist_biological_samples$Field), cex = 0.7)
legend("topright", legend = c("field 3","field 4"), fill = c("green","blue"))

pairs(pca_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Field, pch = 19)
pairs(pca_pareto_center_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Field, pch = 19)

palette(heike_palette_7)
pairs(pca_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch = 19)
pairs(pca_pareto_center_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch = 19)
dev.off()
```


# Transformation: calculation of log-median-ratio
**Divide normalized value by calculate median, then take log10 of this ratio**
**Median is calculated over ALL biological samples of one metabolite (Field 3 and 4 are considered together)**
```{r transformation: calculation of log-median-ratio}
transform_norm_biological_samples_selected <- func_log_transform(norm_biological_samples_selected)
dim(transform_norm_biological_samples_selected)
# 125 306
```


## Boxplot of normalized samples after transformation: look at influence of batch effect (and field)
```{r boxplot of normalized samples after transformation}
pdf("figures/gcms/boxplot_norm_after_transformation.pdf", width=7, height=6)
par(mar=c(8,4,4,2))
for (i in 1:nrow(transform_norm_biological_samples_selected)){
  boxplot(as.numeric(transform_norm_biological_samples_selected[i,]) ~ 
            samplelist_biological_samples$Batch_split30 * 
            samplelist_biological_samples$Field,
          col = heike_palette_7, las = 2,
          main = metabolites_selected$Name_Changed[i])
}
dev.off()
```

## Heatmap after transformation
```{r heatmap after transformation}
transform_norm_biological_samples_selected_mat <- as.matrix(transform_norm_biological_samples_selected)

range(transform_norm_biological_samples_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(-1, -0.21, length=100), seq(-0.2, 0.2, length=100), seq(0.21, 2, length=100))

# colors for two different field trials (columns)
cols_field <- rep("green", 306)
cols_field[samplelist_biological_samples$Field=="field4"] <- "blue"

pdf("figures/gcms/heatmap_biological_samples_transformed.pdf", width = 20, height = 12)
heatmap.2(transform_norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.6, cexCol = 0.6,
          ColSideColors = cols_field)
dev.off()
```


# PCA after transformation
## PCA of transformed data: biological samples
```{r PCA of transformed data: biological samples}
transform_norm_biological_samples_selected_t <- t(transform_norm_biological_samples_selected) 
dim(transform_norm_biological_samples_selected_t)
# 306 samples, 136 metabolites

pca_transform_norm_biological_samples_selected <- pca(transform_norm_biological_samples_selected_t, 
                                                      method = "svdImpute", nPcs = 5, center = F, scale = "none")
pca_pareto_transform_norm_biological_samples_selected <- pca(transform_norm_biological_samples_selected_t, 
                                                             method = "svdImpute", nPcs = 5, center = F, scale = "pareto")
pca_pareto_center_transform_norm_biological_samples_selected <- pca(transform_norm_biological_samples_selected_t, 
                                                                    method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_transform_norm_biological_samples_selected@R2
pca_pareto_center_transform_norm_biological_samples_selected@R2

biplot(pca_transform_norm_biological_samples_selected)
biplot(pca_pareto_transform_norm_biological_samples_selected)
biplot(pca_pareto_center_transform_norm_biological_samples_selected)
```


## PCA plot of transformed biological samples
```{r PCA plot of transformed biological samples}
pdf("figures/gcms/pca_transform_norm_biological_samples.pdf")
x = pca_transform_norm_biological_samples_selected@scores[,1]
y = pca_transform_norm_biological_samples_selected@scores[,2]

palette(heike_palette_7)
plot(x, y, xlab = "PC1 (25.2%)", ylab = "PC2 (13.5%)", main = "normalized and transformed data, no scaling, no centering", 
     col = samplelist_biological_samples$Batch_split30, pch = c(16,17)[as.numeric(samplelist_biological_samples$Field)])
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)
legend("bottomright", legend = c("field 3", "field 4"), pch = c(16,17), cex = 0.7)

palette(c("green","blue"))
plot(x, y, xlab = "PC1 (25.2%)", ylab = "PC2 (13.5%)", main = "normalized and transformed data, no scaling, no centering", col = "white", pch=19)
text(x, y, labels = samplelist_biological_samples$Accession, col = as.numeric(samplelist_biological_samples$Field), cex = 0.7)
legend("topright", legend = c("field 3","field 4"), fill = c("green","blue"))

x = pca_pareto_center_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_transform_norm_biological_samples_selected@scores[,2]
z = pca_pareto_center_transform_norm_biological_samples_selected@scores[,3]

palette(heike_palette_7)
plot(x, y, xlab = "PC1 (18.9%)", ylab = "PC2 (16.1%)", main = "normalized and transformed data, pareto scaling, centering", 
     col = samplelist_biological_samples$Batch_split30, pch = c(16,17)[as.numeric(samplelist_biological_samples$Field)])
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)
legend("bottomright", legend = c("field 3", "field 4"), pch = c(16,17), cex = 0.7)

plot(x, z, xlab = "PC1 (18.9%)", ylab = "PC3 (8.1%)", main = "normalized and transformed data, pareto scaling, centering", 
     col = samplelist_biological_samples$Batch_split30, pch = c(16,17)[as.numeric(samplelist_biological_samples$Field)])
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)
legend("bottomright", legend = c("field 3", "field 4"), pch = c(16,17), cex = 0.7)

palette(c("green","blue"))
plot(x, y, xlab = "PC1 (18.9%)", ylab = "PC2 (16.1%)", main = "normalized and transformed data, pareto scaling, centering", col = "white", pch=19)
text(x, y, labels = samplelist_biological_samples$Accession, col = as.numeric(samplelist_biological_samples$Field), cex = 0.7)
legend("topright", legend = c("field 3","field 4"), fill = c("green","blue"))

pairs(pca_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Field, pch=19, main = "no scaling, no centering")
pairs(pca_pareto_center_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Field, pch=19, main = "pareto scaling, centering")
par(xpd=TRUE)
legend(0, 1, legend = c("field 3","field 4"), fill = c("green","blue"), cex = 0.7 , horiz = T)

palette(heike_palette_7)
pairs(pca_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch=19, main = "no scaling, no centering")
pairs(pca_pareto_center_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch=19, main = "pareto scaling, centering")
legend(0, 1, legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7 , horiz = T)

dev.off()
```


# ANOVA normalization
## generate factors table
```{r generate factors table}
factors <- data.frame("GMO_ID" = samplelist_biological_samples$GMO_ID, 
                      "Accession" = samplelist_biological_samples$Accession,
                      "Field" = samplelist_biological_samples$Field,
                      "Batch" = samplelist_biological_samples$Batch_split30,
                      "Sequence" = samplelist_biological_samples$Sequence_split30)
head(factors)
```


## ANOVA normalization: use normalized data
```{r ANOVA normalization: use normalized data, warning=FALSE}
# samples must be in rows, analytes in columns
norm_biological_samples_selected_t[1:3, 1:3]

# try out log10-tranformation (to avoid negative values)
anova_input <- log10(norm_biological_samples_selected_t)
hist(anova_input, breaks=30)
hist(norm_biological_samples_selected_t, breaks=30)

anova_norm_biological_samples_selected_t <- func_normalize(anova_input, 
                                                           factors,
                                                           facs=c("Accession", "Field", "Sequence", "Batch"),
                                                           keep=c("Accession", "Field"))

dim(anova_norm_biological_samples_selected_t)
# 306 samples, 136 metabolites --> needs to be transposed for log-transformation
range(anova_norm_biological_samples_selected_t, na.rm=T)
hist(anova_norm_biological_samples_selected_t, breaks=30)

# ANOVA-normalization produces negative values!

norm_biological_samples_selected[1:3,1:3]
t(anova_norm_biological_samples_selected_t)[1:3,1:3]

# log-transformation
transform_anova_norm_biological_samples_selected <- func_log_transform(t(anova_norm_biological_samples_selected_t))
dim(transform_anova_norm_biological_samples_selected)
# 125 306

# WARNINGS: calculation log-median-ratio is not working for negative values --> NaNs!!!

transform_norm_biological_samples_selected[1:3,1:3]
transform_anova_norm_biological_samples_selected[1:3, 1:3]

sum(is.na(transform_norm_biological_samples_selected)) # 1440
sum(is.na(transform_anova_norm_biological_samples_selected)) # 1821
```

**Thus: Can't use Sorbitol-normalized data for ANOVA-normalization!**

## ANOVA normalization: use transformed data (log-median-ratio-transformation)
```{r ANOVA normalization: use transformed data}
# samples must be in rows, analytes in columns
transform_norm_biological_samples_selected_t[1:3, 1:3]

range(transform_norm_biological_samples_selected_t, na.rm=T)
hist(transform_norm_biological_samples_selected_t, breaks=30)

anova_transform_norm_biological_samples_selected_t <- func_normalize(transform_norm_biological_samples_selected_t, 
                                                           factors,
                                                           facs=c("Accession", "Field", "Sequence", "Batch"),
                                                           keep=c("Accession", "Field"))

# don't include Field as factor
anova_transform_norm_biological_samples_selected_t2 <- func_normalize(transform_norm_biological_samples_selected_t, 
                                                           factors,
                                                           facs=c("Accession", "Sequence", "Batch"),
                                                           keep=c("Accession"))

dim(anova_transform_norm_biological_samples_selected_t)
# 306 samples, 125 metabolites 
range(anova_transform_norm_biological_samples_selected_t, na.rm=T)
hist(anova_transform_norm_biological_samples_selected_t, breaks=30)

# plot all values before and after ANOVA-normalization
plot(anova_transform_norm_biological_samples_selected_t, transform_norm_biological_samples_selected_t)

# with/without Field as factor
plot(anova_transform_norm_biological_samples_selected_t, anova_transform_norm_biological_samples_selected_t2)
```


# PCA of ANOVA-normalized data
## PCA of ANOVA-normalized data after transformation: biological samples
```{r PCA of ANOVA-normalized data after transformation: biological samples}
dim(anova_transform_norm_biological_samples_selected_t)
# 306 samples, 125 metabolites

pca_anova_transform_norm_biological_samples_selected <- pca(anova_transform_norm_biological_samples_selected_t, 
                                                            method = "svdImpute", nPcs = 5, center = F, scale = "none")
pca_pareto_anova_transform_norm_biological_samples_selected <- pca(anova_transform_norm_biological_samples_selected_t, 
                                                   method = "svdImpute", nPcs = 5, center = F, scale = "pareto")
pca_pareto_center_anova_transform_norm_biological_samples_selected <- pca(anova_transform_norm_biological_samples_selected_t, 
                                                          method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_anova_transform_norm_biological_samples_selected@R2
pca_pareto_center_anova_transform_norm_biological_samples_selected@R2
```


## PCA plot of ANOVA-normalized data biological samples
```{r PCA plot of ANOVA-normalized data biological samples}
pdf("figures/gcms/pca_anova_transform_norm_biological_samples.pdf")
x = pca_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_anova_transform_norm_biological_samples_selected@scores[,2]

palette(heike_palette_7)
plot(x, y, xlab = "PC1 (24.8%)", ylab = "PC2 (11.7%)", main = "normalized, transformed and ANOVA-normalized data, no scaling, no centering", 
     col = samplelist_biological_samples$Batch_split30, pch = c(16,17)[as.numeric(samplelist_biological_samples$Field)])
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)
legend("bottomright", legend = c("field 3", "field 4"), pch = c(16,17), cex = 0.7)

palette(c("green","blue"))
plot(x, y, xlab = "PC1 (24.8%)", ylab = "PC2 (11.7%)", main = "normalized, transformed and ANOVA-normalized data, no scaling, no centering", col = "white", pch=19)
text(x, y, labels = samplelist_biological_samples$Accession, col = as.numeric(samplelist_biological_samples$Field), cex = 0.7)
legend("topright", legend = c("field 3","field 4"), fill = c("green","blue"))

x = pca_pareto_center_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_anova_transform_norm_biological_samples_selected@scores[,2]

palette(heike_palette_7)
plot(x, y, xlab = "PC1 (19.5%)", ylab = "PC2 (14.5%)", main = "normalized, transformed and ANOVA-normalized data, pareto scaling, centering", 
     col = samplelist_biological_samples$Batch_split30, pch = c(16,17)[as.numeric(samplelist_biological_samples$Field)])
legend("topright", legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7)
legend("bottomright", legend = c("field 3", "field 4"), pch = c(16,17), cex = 0.7)

palette(c("green","blue"))
plot(x, y, xlab = "PC1 (19.5%)", ylab = "PC2 (14.5%)", main = "normalized, transformed and ANOVA-normalized data, pareto scaling, centering", col = "white", pch=19)
text(x, y, labels = samplelist_biological_samples$Accession, col = as.numeric(samplelist_biological_samples$Field), cex = 0.7)
legend("topright", legend = c("field 3","field 4"), fill = c("green","blue"))

pairs(pca_anova_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Field, pch=19, main = "no scaling, no centering")
pairs(pca_pareto_center_anova_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Field, pch=19, main = "pareto scaling, centering")
par(xpd=TRUE)
legend(0, 1, legend = c("field 3","field 4"), fill = c("green","blue"), cex = 0.7 , horiz = T)

palette(heike_palette_7)
pairs(pca_anova_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch=19, main = "no scaling, no centering")
pairs(pca_pareto_center_anova_transform_norm_biological_samples_selected@scores, col = samplelist_biological_samples$Batch_split30, pch=19, main = "pareto scaling, centering")
legend(0, 1, legend = levels(samplelist_biological_samples$Batch_split30), fill = heike_palette_7, cex = 0.7 , horiz = T)


# frost tolerance
rbPal <- colorRampPalette(c('blue','red'))
samplelist_joined$LT50_ACC_col <- rbPal(10)[as.numeric(cut(samplelist_joined$LT50_ACC, breaks = 10))]
samplelist_joined$LT50_ACC_col[is.na(samplelist_joined$LT50_ACC_col)] <- "grey"

plot(x, y, xlab = "PC1", ylab = "PC2", 
     main = "normalized, transformed and ANOVA-normalized data, pareto scaling, centering", 
     col = samplelist_joined$LT50_ACC_col, pch=19)
text(x, y, labels = samplelist_joined$Accession, pos = 1,
     col = "black", cex = 0.7)

pairs(pca_pareto_center_anova_transform_norm_biological_samples_selected@scores, 
      col = samplelist_joined$LT50_ACC_col, 
      pch=19, main = "no scaling, no centering")


dev.off()
```


## Heatmap after transformation AND ANOVA-tranformation
```{r heatmap after transformation AND ANOVA-tranformation}
anova_transform_norm_biological_samples_selected_mat <- as.matrix(t(anova_transform_norm_biological_samples_selected_t))

range(anova_transform_norm_biological_samples_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(-1, -0.21, length=100), seq(-0.2, 0.2, length=100), seq(0.21, 2, length=100))

# colors for two different field trials (columns)
cols_field <- rep("green", 306)
cols_field[samplelist_biological_samples$Field=="field4"] <- "blue"

pdf("figures/gcms/heatmap_biological_samples_anova_transformed.pdf", width = 20, height = 12)
heatmap.2(anova_transform_norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.6, cexCol = 0.6,
          ColSideColors = cols_field)
dev.off()
```


## Boxplot of ANOVA-normalized data: look at influence of batch effect (and field)
```{r boxplot of ANOVA-normalized data}
pdf("figures/gcms/boxplot_norm_after_transformation_and_anova.pdf", width=7, height=6)
par(mar=c(8,4,4,2))
for (i in 1:nrow(anova_transform_norm_biological_samples_selected_mat)){
  boxplot(as.numeric(anova_transform_norm_biological_samples_selected_mat[i,]) ~ 
            samplelist_biological_samples$Batch_split30 * 
            samplelist_biological_samples$Field,
          col = heike_palette_7, las = 2,
          main = metabolites_selected$Name_Changed[i])
}
dev.off()

pdf("figures/gcms/boxplot_norm_after_transformation_and_anova2.pdf", width=4, height=5)
par(mar=c(3,4,4,2))
for (i in 1:nrow(anova_transform_norm_biological_samples_selected_mat)){
  boxplot(as.numeric(anova_transform_norm_biological_samples_selected_mat[i,]) ~ 
            samplelist_biological_samples$Field,
          col = c("green", "blue"), 
          main = metabolites_selected$Name_Changed[i])
}
dev.off()
```


# Calculate mean over replicates
```{r calculate mean over replicates}
anova_transform_norm_biological_samples_selected_t[1:3, 1:3]
sum(is.na(anova_transform_norm_biological_samples_selected_t))
# 1440
# 3.8%

mean_anova_transform_norm_biological_samples_selected <- aggregate(anova_transform_norm_biological_samples_selected_t, 
                                                                   by = list(samplelist_biological_samples$Accession,
                                                                             samplelist_biological_samples$Field), 
                                                                   mean)

colnames(mean_anova_transform_norm_biological_samples_selected)[c(1,2)] <- c("Accession", "Field")
mean_anova_transform_norm_biological_samples_selected[1:5, 1:5]
dim(mean_anova_transform_norm_biological_samples_selected)
# 150 127
sum(is.na(mean_anova_transform_norm_biological_samples_selected))
# 1221
# 6.5%

samplelist_joined_mean <- join(mean_anova_transform_norm_biological_samples_selected[,c(1,2)], accessions_metadata, by = "Accession")

apply(mean_anova_transform_norm_biological_samples_selected[, -c(1,2)], 2, function(x) sum(is.na(x)) )

# compare percentage of NAs before and after mean calculation 
# some metabolites have too many NAs after mean calculation!
plot(apply(anova_transform_norm_biological_samples_selected_t, 2, function(x) sum(is.na(x)) )/306*100,
     apply(mean_anova_transform_norm_biological_samples_selected[,-c(1,2)], 2, function(x) sum(is.na(x)) )/150*100, 
     pch=19, xlim=c(0,50), ylim=c(0,50), xlab="percentage NA BEFORE mean calculation", ylab="percentage NA AFTER mean calculation")
abline(0,1)
grid(10,10)
```


## Heatmap after mean calculation
```{r heatmap after mean calculation}
mean_anova_transform_norm_biological_samples_selected_mat <- as.matrix(t(mean_anova_transform_norm_biological_samples_selected[,-c(1,2)]))
colnames(mean_anova_transform_norm_biological_samples_selected_mat) <- interaction(mean_anova_transform_norm_biological_samples_selected$Field,
                                                                                   mean_anova_transform_norm_biological_samples_selected$Accession)

# save mean table
write.table(mean_anova_transform_norm_biological_samples_selected_mat, 
            "output/gcms/mean_anova_transform_norm_biological_samples_selected.txt", sep="\t")

range(mean_anova_transform_norm_biological_samples_selected_mat, na.rm=T)
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 299)
col_breaks <- c(seq(-1, -0.21, length=100), seq(-0.2, 0.2, length=100), seq(0.21, 2, length=100))

# colors for two different field trials (columns)
cols_field_mean <- rep("green", 150)
cols_field_mean[mean_anova_transform_norm_biological_samples_selected$Field=="field4"] <- "blue"

pdf("figures/gcms/heatmap_biological_samples_mean_anova_transformed.pdf", width = 20, height = 12)
heatmap.2(mean_anova_transform_norm_biological_samples_selected_mat, 
          na.rm = T, key = T,
          scale = "none", trace = "none", 
          col = my_palette, breaks = col_breaks, 
          na.color = "grey",
          cexRow = 0.6, cexCol = 0.6,
          ColSideColors = cols_field_mean)
dev.off()
```


# PCA of mean data
## PCA of mean data (ANOVA-normalized data after transformation)
```{r PCA of mean data}
dim(mean_anova_transform_norm_biological_samples_selected[,-c(1,2)]) # remove first two columns!
# 150 samples, 125 metabolites

pca_mean_anova_transform_norm_biological_samples_selected <- pca(mean_anova_transform_norm_biological_samples_selected[,-c(1,2)], 
                                                            method = "svdImpute", nPcs = 5, center = F, scale = "none")
pca_pareto_mean_anova_transform_norm_biological_samples_selected <- pca(mean_anova_transform_norm_biological_samples_selected[,-c(1,2)], 
                                                   method = "svdImpute", nPcs = 5, center = F, scale = "pareto")
pca_pareto_center_mean_anova_transform_norm_biological_samples_selected <- pca(mean_anova_transform_norm_biological_samples_selected[,-c(1,2)], 
                                                          method = "svdImpute", nPcs = 5, center = T, scale = "pareto")

pca_mean_anova_transform_norm_biological_samples_selected@R2
pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@R2
```


## PCA scores plot of mean data (ANOVA-normalized data after transformation)
```{r PCA scores plot of mean data}
pdf("figures/gcms/pca_mean_anova_transform_norm_biological_samples.pdf")
x = pca_mean_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_mean_anova_transform_norm_biological_samples_selected@scores[,2]

palette(c("green","blue"))
plot(x, y, xlab = "PC1 (30.7%)", ylab = "PC2 (11.7%)", 
     main = "normalized, transformed and ANOVA-normalized MEAN data, no scaling, no centering", col = "white", pch=19)
text(x, y, labels = mean_anova_transform_norm_biological_samples_selected$Accession, 
     col = as.numeric(mean_anova_transform_norm_biological_samples_selected$Field), cex = 0.7)
legend("bottomleft", legend = c("field 3","field 4"), fill = c("green","blue"))

x = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,2]

palette(c("green","blue"))
plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", 
     main = "normalized, transformed and ANOVA-normalized MEAN data, pareto scaling, centering", col = "white", pch=19)
text(x, y, labels = mean_anova_transform_norm_biological_samples_selected$Accession, 
     col = as.numeric(mean_anova_transform_norm_biological_samples_selected$Field), cex = 0.7)
legend("topleft", legend = c("field 3","field 4"), fill = c("green","blue"))

# pairs plot
pairs(pca_mean_anova_transform_norm_biological_samples_selected@scores, 
      col = mean_anova_transform_norm_biological_samples_selected$Field, 
      pch=19, main = "no scaling, no centering")
pairs(pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores, 
      col = mean_anova_transform_norm_biological_samples_selected$Field,
      pch=19, main = "pareto scaling, centering")
par(xpd=TRUE)
legend(0, 1, legend = c("field 3","field 4"), fill = c("green","blue"), cex = 0.7 , horiz = T)

dev.off()
```

## PCA scores plot of mean data for frost tolerance, longitude, latitude
```{r PCA scores plot of mean data for frost tolerance, longitude, latitude}
x = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,1]
y = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores[,2]

rbPal <- colorRampPalette(c('blue','red'))
display.brewer.pal(11,"RdYlGn")

# frost tolerance: LT50_ACC
samplelist_joined_mean$LT50_ACC_col <- rbPal(10)[as.numeric(cut(samplelist_joined_mean$LT50_ACC, breaks = 10))]
samplelist_joined_mean$LT50_ACC_col[is.na(samplelist_joined_mean$LT50_ACC_col)] <- "grey"

# longitude
samplelist_joined_mean$Longitude_col <- brewer.pal(11, "RdYlGn")[as.numeric(cut(samplelist_joined_mean$Longitude, breaks = 11))]
samplelist_joined_mean$Longitude_col[is.na(samplelist_joined_mean$Longitude_col)] <- "grey"

# latitude
samplelist_joined_mean$Latitude_col <- brewer.pal(11, "RdYlGn")[as.numeric(cut(samplelist_joined_mean$Latitude, breaks = 11))]
samplelist_joined_mean$Latitude_col[is.na(samplelist_joined_mean$Latitude_col)] <- "grey"

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = samplelist_joined_mean$LT50_ACC_col, pch = 19, main = "LT50_ACC gradient", cex = 1.5)
text(x, y, labels = samplelist_joined_mean$Accession, col = "black", cex = 0.7)

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = samplelist_joined_mean$Longitude_col, pch = 19, main = "Longitude gradient", cex = 1.5)
text(x, y, labels = samplelist_joined_mean$Accession,  col = "black", cex = 0.7)

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = samplelist_joined_mean$Latitude_col, pch = 19, main = "Latitude gradient", cex = 1.5)
text(x, y, labels = samplelist_joined_mean$Accession, col = "black", cex = 0.7)

pairs(pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@scores, 
      col = samplelist_joined_mean$LT50_ACC_col, 
      pch=19, main = "no scaling, no centering")
```


## PCA loadings plot of mean data (ANOVA-normalized data after transformation)
```{r PCA loadings plot of mean data}
pdf("figures/gcms/pca_loadings_mean_anova_transform_norm_biological_samples.pdf")

x = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@loadings[,1]
y = pca_pareto_center_mean_anova_transform_norm_biological_samples_selected@loadings[,2]

palette("default")

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = "white", pch=19,
     main = "normalized, transformed and ANOVA-normalized MEAN data, pareto scaling, centering")
text(x, y, labels = names(x), col = "black", cex = 0.7)

plot(x, y, xlab = "PC1 (23.7%)", ylab = "PC2 (14.3%)", col = "white", pch=19,
     main = "normalized, transformed and ANOVA-normalized MEAN data, pareto scaling, centering")
text(x, y, labels = metabolites_selected$Name_Changed, cex = 0.7, col = as.integer(metabolites_selected$Class))

dev.off()

table(metabolites_selected$Class)

biplot(pca_pareto_center_mean_anova_transform_norm_biological_samples_selected)
```

# Save workspace
```{r save workspace}
save.image("gcms_arafields.RData")
```

