---
title: "Plots of metabolite data Arabidopsis field trial 4 (2009/10)"
author: "Heike Sprenger"
date: "Tuesday, May 07, 2015"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
#getwd()
#setwd("D:/work/repos/ara_field_trials/")
#setwd("~/work/repos/ara_field_trials/")
```


# Load workspace, packages and scripts
```{r load workspace}
# load packages
library(knitr)
library(pander)
library(ggplot2)
library(lattice)

trellis.device() # zuruecksetzen auf default
show.settings() # Farbeinstellungen usw. anzeigen

# png default size: 480x480 pixel

# set options for knitr
opts_chunk$set(fig.width=10, fig.height=5)
opts_knit$set(root.dir = '../')

# load workspace
#load("metabolite_arafield4.RData")
```


# Source functions
```{r source functions}
source("../libpurzel/func_bw_plot.R")
source("../libpurzel/func_boxplot.R")
source("../libpurzel/func_cv.R")
```


# Load data
Use changed metabolite data, from ``metabolite_arafield4_overview4.Rmd``
```{r load data}
metabolite_arafield4 <- read.table("output/metabolite_arafield4.txt", header=TRUE, sep="\t")

lt50 <- read.table("data/LT50.txt", header=TRUE, sep="\t")
pander(head(lt50))

yCapacity4full <- metabolite_arafield4$LT50ACC - metabolite_arafield4$LT50NAC 
yCapacity4 <- lt50$LT50.ACC - lt50$LT50.NAC
```


# Boxplots (from lattice package)
## Boxplot of metabolites and LT50 for all samples
```{r boxplots from lattice package}
pdf("figures/Metabolite_AraField4_bwplots.pdf")   

func_bw_plot(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot(metabolite_arafield4$malate_positive, title = "Malate")
func_bw_plot(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot(metabolite_arafield4$anthocyaninT, title = "Anthocyanins", 
             xlab_text = expression("absolute Absorption A"[530]*"/g FG"))
func_bw_plot(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants", 
             xlab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants", 
             xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot(yCapacity4, title = "Acclimation capacity", 
             xlab_text = expression("decrease of LT"[50]*" in °C by acclimation"))

dev.off()
```


## Boxplots of metabolites and LT50 per BBCH category
```{r boxplots of metabolites and LT50 per BBCH category}
pdf("figures/Metabolite_AraField4_bwplots_per_bbch.pdf")   
cat1_value <- metabolite_arafield4$bbch_categories_en

func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield4$malate_positive, title = "Malate")
func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield4$anthocyaninT, title = "Anthocyanins",
                  ylab_text = expression("absolute Absorption A"[530]*"/g FG"))
func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(yCapacity4, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))
dev.off()
```


## Boxplots of metabolites and LT50 per flowering state
**here: is equal to BBCH category because there are no plants in vegetative stage**
```{r boxplots of metabolites and LT50 per flowering state}
pdf("figures/Metabolite_AraField4_bwplots_per_flowering_state.pdf")  

trellis.par.set(box.dot = new.dot, 
                box.rectangle = new.rectangle, 
                box.umbrella = new.umbrella, 
                plot.symbol = new.symbol, 
                strip.background = new.strip.background, 
                strip.shingle = new.strip.shingle,
                axis.text = new.axis.text, 
                par.main.text = new.par.main.text, 
                par.xlab.text = new.par.xlab.text, 
                par.ylab.text = new.par.ylab.text, 
                par.strip.text = new.par.strip.text)

cat1_value <- metabolite_arafield4$flowering_state

func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield4$malate_positive, title = "Malate")
func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield4$anthocyaninT, title = "Anthocyanins",
                  ylab_text = expression("absolute Absorption A"[530]*"/g FG"))
func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(yCapacity4, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))

dev.off()
```


## Boxplots of metabolites and LT50 per vegetative state
**doesn't make sense here: there are no plants in vegetative stage!**
```{r boxplots of metabolites and LT50 per vegetative state}
# pdf("figures/Metabolite_AraField4_bwplots_per_vegetative_state.pdf")  
# cat1_value <- metabolite_arafield4$vegetative_state
# 
# func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
# func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
# func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
# func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
# func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
# func_bw_plot_cat1(metabolite_arafield4$malate_positive, title = "Malate")
# func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
# func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
# func_bw_plot_cat1(metabolite_arafield4$anthocyaninT, title = "Anthocyanins",
#                   ylab_text = expression("absolute Absorption A"[530]*"/g FG"))
# func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
#                   ylab_text = expression("LT"[50]*"NAC in °C"))
# func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
#                   xlab_text = expression("LT"[50]*"ACC in °C"))
# func_bw_plot_cat1(yCapacity4, title = "Acclimation capacity", 
#                   ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))

# dev.off()
```


## Boxplots of metabolites and LT50 per anthocyanin score
```{r boxplots of metabolites and LT50 per anthocyanin score}
pdf("figures/Metabolite_AraField4_bwplots_per_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield4$antho_value_en

func_bw_plot_cat1(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat1(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat1(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat1(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat1(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat1(metabolite_arafield4$malate_positive, title = "Malate")
func_bw_plot_cat1(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat1(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat1(metabolite_arafield4$anthocyaninT, title = "Anthocyanins",
                  ylab_text = expression("absolute Absorption A"[530]*"/g FG"))
func_bw_plot_cat1(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat1(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat1(yCapacity4, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))

dev.off()
```


## Boxplots of metabolites and LT50 per BBCH categories and anthocyanin score
**Attention: Chunck is not evaluated because it resulted in error messages during knitting HTML document, but not during normal code running!**
```{r boxplots of metabolites and LT50 per BBCH categories and anthocyanin score, eval=FALSE}
pdf("figures/Metabolite_AraField4_bwplots_per_bbch_and_anthocyanin_score.pdf")  
cat1_value <- metabolite_arafield4$antho_value_en
cat2_value <- metabolite_arafield4$bbch_categories_en

func_bw_plot_cat2(metabolite_arafield4$glucoseT, title = "Glucose")
func_bw_plot_cat2(metabolite_arafield4$fructoseT, title = "Fructose")
func_bw_plot_cat2(metabolite_arafield4$sucroseT, title = "Sucrose")
func_bw_plot_cat2(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_bw_plot_cat2(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_bw_plot_cat2(metabolite_arafield4$malate_positive, title = "Malate")
func_bw_plot_cat2(metabolite_arafield4$fumarateT, title = "Fumarate")
func_bw_plot_cat2(metabolite_arafield4$prolineT, title = "Proline")
func_bw_plot_cat2(metabolite_arafield4$anthocyaninT, title = "Anthocyanins",
                  ylab_text = expression("absolute Absorption A"[530]*"/g FG"))
func_bw_plot_cat2(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants",
                  ylab_text = expression("LT"[50]*"NAC in °C"))
func_bw_plot_cat2(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants",
                  xlab_text = expression("LT"[50]*"ACC in °C"))
func_bw_plot_cat2(yCapacity4, title = "Acclimation capacity", 
                  ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))

dev.off()
```

# Boxplots (from graphics package)
## Boxplots to get overview of measurements
```{r boxplots from graphics package}
pdf("figures/Metabolite_AraField4_Boxplots.pdf")
par(mar=c(5.1, 5.1, 4.1, 2.1))

func_boxplot_simple(metabolite_arafield4$glucoseT, title = "Glucose")
func_boxplot_simple(metabolite_arafield4$fructoseT, title = "Fructose")
func_boxplot_simple(metabolite_arafield4$sucroseT, title = "Sucrose")
func_boxplot_simple(metabolite_arafield4$raffinoseT, title = "Raffinose")
func_boxplot_simple(metabolite_arafield4$amino_acidsT, title = "Amino acids")
func_boxplot_simple(metabolite_arafield4$malate_positive, title = "Malate")
func_boxplot_simple(metabolite_arafield4$fumarateT, title = "Fumarate")
func_boxplot_simple(metabolite_arafield4$prolineT, title = "Proline")
func_boxplot_simple(metabolite_arafield4$anthocyaninT, title = "Anthocyanins", 
             ylab_text = expression("absolute Absorption A"[530]*"/g FG"))
func_boxplot_simple(metabolite_arafield4$LT50NAC, title = "Frost tolerance of non-acclimated plants", 
             ylab_text = expression("LT"[50]*"NAC in °C"))
func_boxplot_simple(metabolite_arafield4$LT50ACC, title = "Frost tolerance of acclimated plants", 
             ylab_text = expression("LT"[50]*"ACC in °C"))
func_boxplot_simple(yCapacity4, title = "Acclimation capacity", 
             ylab_text = expression("decrease of LT"[50]*" in °C by acclimation"))

dev.off()
```


# Boxplots to get overview of coefficient of variation (CV)
```{r boxplots to get overview of coefficient of variation}
# calculate CV per metabolite and accession
metabolite_arafield4.x <- subset(metabolite_arafield4, 
                                 select= c(accession, glucose_positive, fructose_positive, sucrose_positive, 
                                           raffinose_positive, amino_acids_positive, malate_positive, 
                                           fumarate_positive, proline_positive, anthocyanin_positive))
head(metabolite_arafield4.x)
metabolite_arafield4_CV <- aggregate(. ~ accession, metabolite_arafield4.x, func_CV)
head(metabolite_arafield4_CV)

ylab_text_cv <- "Coefficient of variation (CV)"

pdf("figures/Metabolite_AraField4_CV_Boxplots.pdf")
par(mar=c(9, 5, 4, 2))
# all metabolites
boxplot(metabolite_arafield4_CV[,-1], main = "all metabolites", ylab = ylab_text_cv, las=2)
# single metabolites
par(mar=c(5, 5, 4, 2))
func_boxplot_simple(metabolite_arafield4_CV$glucose_positive, title = "Glucose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$fructose_positive, title = "Fructose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$sucrose_positive, title = "Sucrose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$raffinose_positive, title = "Raffinose", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$amino_acids_positive, title = "Amino acids", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$malate_positive, title = "Malate", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$fumarate_positive, title = "Fumarate", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$proline_positive, title = "Proline", ylab_text = ylab_text_cv)
func_boxplot_simple(metabolite_arafield4_CV$anthocyanin_positive, title = "Anthocyanins", ylab_text = ylab_text_cv)
dev.off()
```


# Pairs plot
```{r pairs plot}
pdf("figures/Metabolite_AraField4_Pairs.pdf", width=15, height=15)
pairs(metabolite_arafield4[,c(16:17,21,24,31:39)], 
      panel = panel.smooth, labels = c(expression("LT"[50]*"NAC"), 
                                       expression("LT"[50]*"ACC"),"BBCH",
                                       "Antho- \n Score","Glc","Frc","Suc",
                                       "Raf","AA","Mal","Fum","Pro","Antho- \n cyane"), 
      cex.labels = 2, font.labels = 2, cex.main = 2, cex.axis = 1.5, 
      main = "complete dataset field trial 4", gap = 0.1)
dev.off()
```


# Save workspace
```{r save workspace}
save.image("metabolite_arafield4.RData")
```

